# Running the Model

## Inport libraries

First, download the package from github.

```{r}
library(devtools)
# install_github("josimms/CASSIA")
library(CASSIA)
```

## Run the model

The model has automatic settings which produces the results as in the 2015 paper in Hyytiälä. The parameters and input weather for Hyytiälä are automatically included in the package.

```{r}
out = CASSIA_cpp(data_format, site = "Hyde") # TODO: make this work
```

The model has a list datasets for the output. These are for the growth, sugar model and preles (photosynthesis model) respectively.

The original model or basic settings mainly has results in the Growth database.

```{r}
head(out$Growth)
```

The other settings of the model an be accessed by the function argument switches. These settings can be found in the help function for the CASSIA model. 

The following example means that the trees growth, so that the model uses the previous year's growth as the initial conditions for the next year.

```{r}
out2 = CASSIA_cpp(data_format, site = "Hyde", trees_grow = TRUE) # TODO: make this work

# To ccheck the possible options
?CASSIA
```

To change the parameters they must be in a table form as the preloaded ratios_p, parameters_p, common_p. sperling_p and repo_p. The easiest way of doing this could be to redefine the parameters from the default objects.

The meaning of the parameters can be found in the CASSIA instruction booklet.

```{r}
parameters_updated <- parameters_p
parameters_updated[c("root.lifetime	"),c("Hyde")] <- 2

out3 = CASSIA_cpp(data_format, site = "Hyde", parameters = parameters_updated) # TODO: make this work
```

Finally, the weather data should be considered. Due to the linking with different photosynthesis models the weather needed in the model currently takes two forms. One for the basic model which includes the photosynthesis values as an input and the other the environmental variables needed for the PRELES model.

Due to the current definition of the model the weather data should be in the same format as the default Hyyyiälä data even if the PRELES model is not used to calculate the photosynthesis input. However the values in these columns can just be NA values.

```{r}
head(data_format)
```
