## Inport libraries

```{r}
library(devtools)
# install_github("josimms/CASSIA", force = TRUE)

library(CASSIA)
library(data.table)
library(ggplot2)
library(purrr)
library(dplyr)
library(tidyr)
library(viridis)
library(gridExtra)
```

## Useful Functions for later!

```{r}
# Function that changes the input values inputted into the code!
run_sensitivity <- function(param_name, range) {
  map(range, function(i) {
    parameters_p_G_new <- parameters_p_G
    if (param_name %in% rownames(parameters_p_G_new)) {
      parameters_p_G_new$HF_China[rownames(parameters_p_G_new) == param_name] <- i
      result <- CASSIA_cpp(weather, site = "HF_China",
                           ratios = ratios_p_G,
                           parameters = parameters_p_G_new,
                           common = common_p_G,
                           repo = repo_p_G,
                           photosynthesis_as_input = TRUE,
                           LD_estim = FALSE)
      list(param_value = i, result = result)
    } else {
    print("Parameter not in the row names, is it spelt correctly? Is there an underscore rather than a dot?")
    }
  })
}

# Function that extracts the rows that we want to plot
prepare_plot_data <- function(sensitivity_results, param_name, range) {
  result_list <- list()
  
  for (i in seq_along(sensitivity_results)) {
    result <- sensitivity_results[[i]]$result
    param_value <- range[i]
    
    if (is.null(result$Growth) || is.null(result$Growth$ring_width)) {
      print("Warning: Growth or ring_width is NULL")
      next
    }
    
    data <- tibble(
      param_name = param_name,
      param_value = param_value,
      days = seq_along(result$Growth$ring_width),
      ring_width = result$Growth$ring_width
    )
    
    result_list[[i]] <- data
  }
  
  final_result <- bind_rows(result_list)
  return(final_result)
}
```

## Weather and parameters imported

```{r}
wd = "/home/josimms/Documents/CASSIA_yiheng/DATA/"

site = "HF_China"
HF_weather <- read.csv(paste0(wd, "zzs_weather.csv"),header = T,sep=',')
weather <- HF_weather
weather_extra <- read.csv(paste0(wd, "zzs_Weather_extra.csv"),header = T,sep=',')

weather <- cbind(weather, weather_extra)

dates <- seq(as.Date(weather$date[1]), as.Date(weather$date[nrow(weather)]), by = "day")

parameters_p_G <- read.csv(paste0(wd, "parameters_p_G.csv"), header = T, sep=',', row.names = 1)
common_p_G <- read.csv(paste0(wd, "common_p_g.csv"), header = T, sep=',', row.names = 1)
repo_p_G <- read.csv(paste0(wd, "repo_p_g.csv"), header = T, sep=',', row.names = 1)
ratios_p_G <-read.csv(paste0(wd, "ratios_p_g.csv"), header = T, sep=',', row.names = 1)
```

## Weather plotted

```{r}
plot_weather_variables(HF_weather[,-1], dates)
```

## CASSIA with the original parameters, but the chinese weather data

```{r}
CASSIA_out_original_values <- CASSIA_cpp(weather, site = "HF_China",
                                         ratios = ratios_p_G,
                                         parameters = parameters_p_G,
                                         common = common_p_G,
                                         repo = repo_p_G,
                                         photosynthesis_as_input = TRUE,
                                         LD_estim = FALSE)
```

## Sensitivity analysis

So although you have changed the weather parameters the results will not be ready as the parameters have not been changed.
If you look at the 2015 paper then you can see the temperature controls for the seasonal behaviour
First let's get the ring width to work, then it's possible to look at the other areas of growth.

First we look at the diameter growth.
g is controlled by the air temperature and common parameters a and b.
sD0.Trad and sDc control the seasonality of the fD multiplier.
And finally the LD multiplier is controlled by the LD0 and diameter_growth_coefficient parameters.
LD_estim being true alters the summer response of the LD multiplier. But this is currently set up for boreal weather conditions. (TODO: make this dependent on temperature). I would advise you to have the LD_estim parameter as FALSE for now, but we can look into this later!

SEASONAL PARAMETER CHECK

(functions found after library!)

```{r}
#ratio.diameter_growth_coefficient
#parameters.LD0

# Run sensitivity analyses
sensitivity_sDc <- run_sensitivity("sDc", seq(0, 10, length.out = 20))
sensitivity_sD0_Trad <- run_sensitivity("sD0.Trad", seq(-7, 0, length.out = 20))
sensitivity_dgs <- run_sensitivity("diameter_start_day", seq(50, 200, length.out = 20))
sensitivity_uggla <- run_sensitivity("Uggla", seq(1, 10, length.out = 20))

# Prepare plot data
plot_data <- bind_rows(
  prepare_plot_data(sensitivity_sDc, "sDc", seq(4, 6, length.out = 20)),
  prepare_plot_data(sensitivity_sD0_Trad, "sD0.Trad", seq(-7, 0, length.out = 20)),
  prepare_plot_data(sensitivity_dgs, "diameter_start_day", seq(50, 100, length.out = 20)),
  prepare_plot_data(sensitivity_uggla, "Uggla", seq(1, 10, length.out = 20))
)

par(mfrow = c(2, 2))
for (param in unique(plot_data$param_name)) {
  to_plot = plot_data[plot_data$param_name == param,]
  cols = rainbow(20)
  plot(NULL,
       xlim = c(min(to_plot$days),
                max(to_plot$days)),
       ylim = c(min(to_plot$ring_width), 
                max(to_plot$ring_width)), main = param,
       xlab = "Days",
       ylab = "Ring Width")
  count = 1
  for (value in unique(to_plot$param_value)) {
    lines = to_plot[to_plot$param_value == value,]
    points(lines$days, lines$ring_width, col = cols[count], pch = "-")
    count = count + 1
  }
}

```

RING WIDTH AND DIAMETER PARAMETERS

The diameter growth, itself depends on:
tau_Ee, tau_We, tau_El, tau_Wl
And ring width also depends on
Uggla, cell_d_ew

```{r}
## Look at the late and early wood parameters
sensitivity_tau_Ee <- run_sensitivity("tau.Ee", seq(1, 10, length.out = 20))
sensitivity_tau_We <- run_sensitivity("tau.We", seq(1, 10, length.out = 20))
sensitivity_tau_El <- run_sensitivity("tau.El", seq(1, 10, length.out = 20))
sensitivity_tau_Wl <- run_sensitivity("tau.Wl", seq(1, 10, length.out = 20))
sensitivity_cell_d_ew <- run_sensitivity("cell.d.ew", seq(0.00001, 0.001, length.out = 20))
sensitivity_cell_d_ew <- run_sensitivity("cell.d.lw", seq(0.00001, 0.001, length.out = 20))
sensitivity_cell_l_ew <- run_sensitivity("cell.l.ew", seq(0.00001, 0.001, length.out = 20))
sensitivity_cell_l_ew <- run_sensitivity("cell.l.lw", seq(0.00001, 0.001, length.out = 20))

plot_data <- bind_rows(
  prepare_plot_data(sensitivity_tau_Ee, "tau_Ee", seq(1, 10, length.out = 20)),
  prepare_plot_data(sensitivity_tau_We, "tau_We", seq(1, 10, length.out = 20)),
  prepare_plot_data(sensitivity_tau_El, "tau_El", seq(1, 10, length.out = 20)),
  prepare_plot_data(sensitivity_tau_Wl, "tau_Wl", seq(1, 10, length.out = 20)),
  prepare_plot_data(sensitivity_cell_d_ew, "cell_d_ew", seq(0.00001, 0.001, length.out = 20)),
  prepare_plot_data(sensitivity_cell_d_ew, "cell_d_lw", seq(0.00001, 0.001, length.out = 20)),
  prepare_plot_data(sensitivity_cell_d_ew, "cell_l_ew", seq(0.00001, 0.001, length.out = 20)),
  prepare_plot_data(sensitivity_cell_d_ew, "cell_l_lw", seq(0.00001, 0.001, length.out = 20))
)

par(mfrow = c(2, 2))
for (param in unique(plot_data$param_name)) {
  to_plot = plot_data[plot_data$param_name == param,]
  cols = rainbow(20)
  plot(NULL,
       xlim = c(min(to_plot$days),
                max(to_plot$days)),
       ylim = c(min(to_plot$ring_width), 
                max(to_plot$ring_width)), main = param,
       xlab = "Days",
       ylab = "Ring Width")
  count = 1
  for (value in unique(to_plot$param_value)) {
    lines = to_plot[to_plot$param_value == value,]
    points(lines$days, lines$ring_width, col = cols[count], pch = "-")
    count = count + 1
  }
}
```

## Final parameters for the HF site

```{r}

## Final parameters when you have them!
parameters_p_G_final <- parameters_p_G

CASSIA_out <- CASSIA_cpp(weather, site = "HF_China",
                         ratios = ratios_p_G,
                         parameters = parameters_p_G,
                         common = common_p_G,
                         repo = repo_p_G,
                         photosynthesis_as_input = TRUE)

```

## Monthly aggregation

```{r}
convert_yday_to_month <- function(year, day_of_year) {
  # Create a date object
  date <- as.Date(paste(year, day_of_year), format="%Y %j")

  # Extract year and month
  year_month <- format(date, "%Y-%m")

  return(year_month)
}

CASSIA_out$Growth$YM <- convert_yday_to_month(CASSIA_out$Growth$year, CASSIA_out$Growth$day)
daily <- data.table::as.data.table(CASSIA_out$Growth)
monthly_mean <- daily[, lapply(.SD, mean, na.rm = T), by = YM, .SDcols = -c("year", "day")]
monthly_sum <- daily[, lapply(.SD, sum, na.rm = T), by = YM, .SDcols = -c("year", "day")]
```

## Testing the data

```{r}

# Generated random data here, just to plot the graphs
your_data <- 0.01* abs(rnorm(nrow(monthly_sum))) # TODO: import your data here!

# Plot your data and CASSIA data
par(mfrow = c(2, 1), mar=c(5.1,6,4.1,2.1))
plot(as.Date(paste0(monthly_mean$YM, "-01")), monthly_mean$roots_growth_potential,
     main = "Wall Mean Growth Monthly Values", type = "l", cex = 0.5,
     xlab = "Year", ylab = "Mean Monthly\n Roots Growth Potential (kg C)")
points(as.Date(paste0(monthly_mean$YM, "-01")), your_data, col = "blue")
legend("topright", c("Your Data", "CASSIA"), lty = c(0, 1), pch = c("o", ""), col = c("blue", "black"), bty = "n")
plot(as.Date(paste0(monthly_sum$YM, "-01")), monthly_sum$roots_growth_potential,
     main = "Wall Sum Growth Monthly Values", type = "l", cex = 0.5,
     xlab = "Year", ylab = "Sum Monthly\n Roots Growth Potential (kg C)")

# If good, then should be straight line
# R^2, RMSE etc.
par(mfrow = c(2, 1))
plot(monthly_mean$roots_growth_potential, your_data,
     main = "Modelled against Measured", xlab = "Modelled Data", ylab = "Wang YiHeng Data")
plot(as.Date(paste0(monthly_mean$YM, "-01")), monthly_mean$roots_growth_potential - your_data,
     main = "Residuals: Modelled - Measured", xlab = "Year", ylab = "Difference between\n modelled and measured")
```
