---
title: "Component Models: Repola and PRELES"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Component Models: Repola and PRELES}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
```

> **More detail:** The CASSIA Instruction Booklet (`CASSIA_Instruction_Booklet.pdf`,
> available in the package root or on GitHub) contains the full mathematical
> descriptions of both the Repola allometry equations and the PRELES
> photosynthesis model as implemented in CASSIA.

CASSIA integrates two external sub-models — Repola allometry and PRELES
photosynthesis — that can also be useful on their own. This vignette shows
how to use each independently and explains how they connect to `CASSIA_cpp`.

---

## Repola biomass allometry

The `repola()` function implements the Repola (2009) mixed-effects allometry
model for Scots pine, relating stem diameter and tree height to needle mass,
sapwood mass, and fine root mass. CASSIA calls it internally during
initialisation to set starting organ masses for each simulation year.

### Standalone use

The function needs diameter at breast height (`D0`, m), tree height (`h0`, m),
a year index (`n.year`), a needle mass history vector, a site string, and the
`parameters_p` and `repo_p` matrices.

```{r}
# A 15 cm diameter, 12 m tall Scots pine at Hyytiälä
out_repola <- repola(
  D0        = 0.15,
  h0        = 12,
  n.year    = 1,
  needle_mas = rep(NA, 10),   # history vector; only index n.year is written
  ste       = "Hyde",
  params    = parameters_p,
  reps      = repo_p
)
names(out_repola)
```

```{r}
# Total needle carbon (kg C tree⁻¹)
out_repola$m.N.tot

# Fine root carbon (= 50 % of needle carbon by default)
out_repola$m.R.tot

# Needle mass per mm of the youngest cohort
out_repola$m.N
```

### Scanning a diameter range

```{r, fig.width=6, fig.height=4}
D0_seq <- seq(0.05, 0.30, by = 0.01)

needle_mass <- sapply(D0_seq, function(d) {
  repola(D0 = d, h0 = 10 + d * 30, n.year = 1,
         needle_mas = rep(NA, 10),
         ste = "Hyde", params = parameters_p, reps = repo_p)$m.N.tot
})

plot(D0_seq * 100, needle_mass,
     type = "l", lwd = 2,
     xlab = "DBH (cm)", ylab = "Needle mass (kg C tree⁻¹)",
     main = "Repola allometry: needle mass vs diameter")
```

### How it connects to CASSIA_cpp

`CASSIA_cpp` calls `repola` during setup to compute `m.N.tot` and `m.R.tot`
before the C++ loop begins. You can supply a custom needle mass history via the
`needle_mass` argument to `CASSIA_cpp` if your tree has measured values.

---

## PRELES photosynthesis

PRELES (Prentice *et al.* 2014) is a process-based GPP model driven by PAR,
VPD, precipitation, temperature, CO₂ and fAPAR. CASSIA uses the C++
implementation internally, which is faster than the R version.

### Using PRELES via CASSIA_cpp

Pass `preles = TRUE` together with `photosynthesis_as_input = FALSE` — only
one photosynthesis source may be active at a time. The built-in `data_format`
dataset has the required PRELES columns (`PAR`, `VPD`, `CO2`, `fAPAR`) for
Hyytiälä 2005 onward:

```{r}
names(data_format)
```

```{r, fig.width=7, fig.height=4}
# Take the first year from data_format
weather_preles        <- data_format[1:365, ]
weather_preles$dates  <- as.Date(weather_preles$Date)

out_preles <- CASSIA_cpp(weather = weather_preles, site = "Hyde",
                         photosynthesis_as_input = FALSE,
                         preles = TRUE)

par(mfrow = c(1, 2))
plot(weather_preles$dates, out_preles$Preles$GPP,
     type = "l", main = "PRELES GPP",
     xlab = "Date", ylab = "g C m⁻² day⁻¹")
plot(weather_preles$dates, out_preles$Preles$ET,
     type = "l", col = "blue", main = "PRELES ET",
     xlab = "Date", ylab = "mm day⁻¹")
```

The `out$Preles` data frame columns are:

```{r}
names(out_preles$Preles)
```

### Photosynthesis options

CASSIA supports three ways to supply GPP — exactly one must be active:

| Toggle | Source of GPP |
|--------|----------------|
| `photosynthesis_as_input = TRUE` (default) | Column `P` in weather data — measured or pre-computed GPP |
| `preles = TRUE` | C++ PRELES, driven by PAR, VPD, CO2, fAPAR in the weather data |
| `phydro = TRUE` | p-hydro hydraulic optimality model (also needs PPFD, PPFD\_max, SWP) |

When using `preles = TRUE`, always set `photosynthesis_as_input = FALSE`.

### PRELES_GPP (R wrapper)

The exported function `PRELES_GPP` wraps the R package `Rprebasso` and is
retained for compatibility with the pure-R version of CASSIA. It is not
needed for normal use of `CASSIA_cpp` and requires `Rprebasso` to be
installed separately. Use `preles = TRUE` in `CASSIA_cpp` for the C++
implementation instead.
