---
title: "Exploring Built-in Datasets and Parameters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring Built-in Datasets and Parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
```

> **More detail:** The CASSIA Instruction Booklet (`CASSIA_Instruction_Booklet.pdf`,
> available in the package root or on GitHub) documents every parameter's
> meaning, units, and derivation.

CASSIA ships with pre-calibrated parameter sets and reference datasets for
Hyytiälä (Hyde), Lettosuo, and Väriö/Flakaliden. This vignette shows how to
explore and modify them.

---

## Weather data: weather_original

`weather_original` is a 1096-row data frame covering Hyytiälä SMEAR II
daily measurements for 2015–2017.

```{r}
dim(weather_original)
names(weather_original)
```

Add a `dates` column before passing to `CASSIA_cpp`:

```{r}
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))
head(weather_original[, c("dates", "T", "P", "TSA", "TSB", "MB", "Rain")], 3)
```

Column descriptions: `T` = air temperature (°C), `P` = GPP input
(g C m⁻² day⁻¹), `TSA`/`TSB` = soil temperature A/B horizon (°C),
`MB` = soil moisture B horizon (m³ m⁻³), `Rain` = precipitation (mm day⁻¹).

---

## Site-specific parameters: parameters_p

`parameters_p` is a 75 × 3 matrix — one column per site.

```{r}
dim(parameters_p)
colnames(parameters_p)
```

Browse selected rows by category:

```{r}
# Growth parameters
parameters_p[c("bud.day", "bud.T", "leaf.out", "gD"), ]
```

```{r}
# Respiration parameters
parameters_p[c("Q10.N", "Rm.NR", "Q10.S", "Rm.S", "Q10.R", "Rm.R"), ]
```

To modify a parameter for a sensitivity run, copy the matrix and edit:

```{r}
params_modified <- parameters_p
params_modified["root.lifetime", "Hyde"] <- 2.0   # originally 1.0
```

Then pass it to `CASSIA_cpp`:

```{r, eval=FALSE}
out_mod <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                      parameters = params_modified)
```

---

## Sugar model parameters: sperling_p

`sperling_p` is a 55 × 3 matrix covering sugar kinetics for three sites.
The third site is Flakaliden\_c (a Swedish boreal site).

```{r}
dim(sperling_p)
colnames(sperling_p)
```

```{r}
# Transfer rate parameters (tau = relaxation time, days)
sperling_p[c("tau.s", "tau.t", "SCb", "SCs"), ]
```

---

## Allocation ratios: ratios_p

`ratios_p` holds allocation and morphology ratios for Hyde and Lettosuo.

```{r}
ratios_p
```

---

## Site-independent parameters: common_p

`common_p` is a named numeric vector of constants that do not vary between
sites: soil physics, gas constants, and growth respiration coefficients.

```{r}
length(common_p)
head(common_p, 20)
```

---

## Repola allometry parameters: repo_p

`repo_p` holds the fixed-effect parameters from the Repola (2009) biomass
allometry model, used internally to initialise organ masses.

```{r}
repo_p
```

---

## initialize_parameters

`initialize_parameters()` bundles all the adjustments CASSIA applies to the
raw parameter matrices before a run: it sets lower-bound sugar concentrations,
initialises transfer rates, and assembles the PRELES parameter vector.

```{r}
params <- initialize_parameters()
names(params)
```

The returned list contains:

| Name | Contents |
|------|----------|
| `needle_mass_in` | Initial needle mass (kg C tree⁻¹) |
| `N_parameters` | Nitrogen uptake parameters |
| `pPREL` | 30-element PRELES parameter vector |
| `parameters_test` | Modified `parameters_p` with adjusted bounds |
| `sperling_test` | Modified `sperling_p` with adjusted transfer rates |
| `Throughfall` | Throughfall fraction |
| `parameters_R` | Soil/nitrogen sub-model parameters (length 58) |

You can inspect the PRELES vector directly:

```{r}
params$pPREL
```

---

## Reference datasets

| Dataset | Contents |
|---------|----------|
| `Hyde_daily_original` | Daily Hyytiälä observations 2015–2018 (growth, respiration, GPP) |
| `Hyde_yearly_original` | Annual summaries of the same observations |
| `GPP_ref` | Reference daily GPP (365 values) used when `LN_estim = TRUE` |
| `oyewole_2015_calibration_data` | Nitrogen uptake calibration data |

```{r}
names(Hyde_daily_original)
head(GPP_ref)
```

---

## Calibration data: load_data

`load_data()` loads the full set of Hyytiälä calibration files (sugar
measurements, trenching fluxes, SMEAR observations, core drilling data).
These files are not bundled in the package; they are stored in a separate
calibration directory.

```{r, eval=FALSE}
loaded <- load_data(data_directory = "~/Documents/CASSIA_Calibration/Processed_Data/")
names(loaded)
# $yu_data            — sugar/starch concentration measurements
# $original_data      — original CASSIA R-version outputs for comparison
# $smearII_data       — SmearII ecological measurements
# $core_drilling_list — list of core drilling ring-width files
# ...
```

Contact the maintainer (joanna.x.simms@helsinki.fi) for access to the
calibration data directory.
