---
title: "ERA5 Weather Data for CASSIA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ERA5 Weather Data for CASSIA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
```

> **More detail:** The CASSIA Instruction Booklet (`CASSIA_Instruction_Booklet.pdf`,
> available in the package root or on GitHub) describes the weather input
> requirements for each sub-model and explains the ERA5 bias-correction
> approach used in `ERAS_reading_nc`.

ERA5 is the ECMWF global reanalysis product and is useful when SMEAR station
data are not available — for example, at new sites or for multi-decadal
hindcasts. This vignette explains the ERA5 workflow in CASSIA and demonstrates
`calculate_VPD`, the key unit-conversion function.

---

## Overview of the ERA5 pipeline

```
ERA5 .nc files
    └─► ERAS_reading_nc()   # reads, aggregates, bias-corrects hourly ERA5 to daily
            └─► ERA5_CASSIA_preles.csv   →  CASSIA_cpp(preles = TRUE)
            └─► ERA5_CASSIA_phydro.csv   →  CASSIA_cpp(phydro = TRUE)
```

The package ships with pre-processed ERA5 datasets for Hyytiälä (1960–2022)
so you can run CASSIA without downloading any ERA5 files.

---

## calculate_VPD

ERA5 provides dew-point temperature (`d2m`) and air temperature (`t2m`) rather
than VPD directly. `calculate_VPD` converts these to relative humidity and VPD
using the Magnus formula.

```{r}
# VPD for a warm, moderately humid summer day
result <- calculate_VPD(dew_point_temp = 12, air_temp = 25)
result$RH    # relative humidity (0–1)
result$VPD   # vapour pressure deficit (hPa)
```

The output VPD is in **hPa**. PRELES expects VPD in kPa, so divide by 10 when
preparing PRELES inputs:

```{r}
vpd_kPa <- result$VPD / 10
vpd_kPa
```

### Applying to a weather data frame

```{r, fig.width=6, fig.height=3}
# Simulate dew-point / temperature pairs over a year
T_air  <- 10 * sin(seq(0, pi, length.out = 365)) - 5   # -5 to +5 °C, boreal
T_dew  <- T_air - 5                                     # 5 °C depression

vpd_hPa <- mapply(function(dp, ta) calculate_VPD(dp, ta)$VPD, T_dew, T_air)

plot(vpd_hPa, type = "l",
     main = "Computed VPD over a synthetic year",
     xlab = "Day of year", ylab = "VPD (hPa)")
```

---

## Pre-processed ERA5 datasets in the package

`ERAS_reading_nc` produces two CASSIA-ready files. Both are included as package
datasets so that most users do not need to re-process the raw ERA5 NetCDF files.

### ERA5_CASSIA_preles — for preles = TRUE

```{r}
head(ERA5_CASSIA_preles)
```

Column | Description | Unit
-------|-------------|-----
`date` | Date string (YYYY-MM-DD) | —
`T` | Air temperature | °C
`VPD` | Vapour pressure deficit | kPa
`PAR` | Photosynthetically active radiation | mol m⁻² day⁻¹
`Rain` | Precipitation | mm day⁻¹
`CO2` | Atmospheric CO₂ | ppm
`fAPAR` | Fraction of absorbed PAR | 0–1
`TSA` | Soil temperature, A horizon | °C
`TSB` | Soil temperature, B horizon | °C
`MB` | Soil moisture, B horizon | m³ m⁻³

Run CASSIA with PRELES photosynthesis on this dataset:

```{r, eval=FALSE}
df <- as.data.frame(ERA5_CASSIA_preles)
df$dates <- as.Date(df$date)

out <- CASSIA_cpp(weather = df, site = "Hyde", preles = TRUE)
```

### ERA5_CASSIA_phydro — for phydro = TRUE

```{r}
head(ERA5_CASSIA_phydro)
```

The phydro dataset includes additional columns needed for the hydraulic
optimality sub-model (`PPFD`, `PPFD_max`, `SWP`):

```{r, eval=FALSE}
df_phydro <- as.data.frame(ERA5_CASSIA_phydro)
df_phydro$dates <- as.Date(df_phydro$date)

out_ph <- CASSIA_cpp(weather = df_phydro, site = "Hyde", phydro = TRUE)
```

---

## ERAS_reading_nc: processing your own ERA5 files

`ERAS_reading_nc` downloads and processes hourly ERA5 NetCDF files for a
target location, aggregates to daily resolution, and bias-corrects against
SMEAR II Hyytiälä observations. It requires:

1. ERA5 NetCDF files in two sets — accumulated (`*_era5_accum.nc`: precipitation,
   radiation) and instantaneous (`*_era5_instant.nc`: temperature, soil, moisture)
2. Raw Hyytiälä SMEAR data files for the bias correction reference

The function is designed for the Hyytiälä coordinates (lat 61.85°N, lon 24.29°E)
and saves its output to disk. Adapt it for a new site by changing
`lon_target`/`lat_target` and supplying appropriate reference data.

```{r, eval=FALSE}
ERAS_reading_nc(
  path_nc   = "/path/to/era5/netcdf/files",
  path_test = "/path/to/output/directory",
  raw.directory = "/path/to/smear/raw/data"
)
```

ERA5 data can be downloaded from the
[Copernicus Climate Data Store](https://cds.climate.copernicus.eu) using the
`ecmwfr` R package. The variables needed are: `2m_temperature` (t2m),
`2m_dewpoint_temperature` (d2m), `total_precipitation` (tp),
`surface_solar_radiation_downwards` (ssrd), `soil_temperature_level_1` (stl1),
`soil_temperature_level_2` (stl2), `volumetric_soil_water_layer_1` (swvl1),
and `volumetric_soil_water_layer_2` (swvl2).
