---
title: "Model Investigation: Ring Width Formation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ring Width Process}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

This vignette walks through the xylogenesis sub-model that underlies ring width
formation in CASSIA. Rather than running the full `CASSIA_cpp()` model, it uses
a transparent R implementation of the same mathematics so that each state
variable can be inspected step by step.

Ring width forms through a sequential cell pipeline: temperature drives a pool
of potential cambial cells into enlargement, then into wall thickening, and
finally into mature cells. The total width of early-wood and late-wood cells at
the end of the season is the potential ring width. Understanding how each stage
responds to parameters and climate helps with calibration and with interpreting
model behaviour.

The weather data here is synthetic — a smooth seasonal temperature cycle with a
corresponding photosynthesis signal. This makes it easy to see how the model
responds to climate signals without the noise of real data. Substitute your own
dataframe to explore site-specific behaviour.

## Parameters and synthetic weather

```{r}
set.seed(42)
n_days <- 365

weather <- data.frame(
  day    = 0:(n_days - 1),
  TAir   = 10 + 10 * sin(2 * pi * (1:n_days) / 365),
  PF     = pmax(0, 5 + 3 * sin(2 * pi * (1:n_days - 80) / 365)),
  GPP_ref = 6
)

parameters <- list(
  sD0_Trad           = -3.72,
  diameter_start_day = 79,
  sDc                = 5.08,
  LD0                = 1.29,
  tau_GPP            = 5,
  tau_Ee             = 10.7,
  tau_We             = 25.3,
  tau_El             = 8.8,
  tau_Wl             = 35.1,
  cell_d_ew          = 35.7e-6,
  cell_d_lw          = 24.2e-6,
  h0                 = 17.9
)

common <- list(
  Uggla              = 1.95,
  osmotic_sugar_conc = 2e6,
  gas_const          = 8.314,
  abs_zero           = 273.15,
  M_C                = 12,
  a                  = 0.185,
  b                  = 18.4
)

ratio <- list(
  diameter_growth_coefficient = 1,
  form_factor                 = 0.5
)

init_state <- list(
  sD              = 0,
  GD              = 0,
  g               = 0,
  fD              = 0,
  n_E_pot         = 0,
  n_W_pot         = 0,
  n_M_pot         = 0,
  tau_W           = parameters$tau_We,
  tau_E           = parameters$tau_Ee,
  ew_cells_pot_max = 0,
  pot_mm_max      = 0
)
```

## R implementation of the growth process

```{r}
grow_one_day <- function(day, weather, state, parameters, common, ratio,
                         LD_estim = TRUE, sD_estim_T_count = FALSE) {

  TAir    <- weather$TAir[day + 1]
  PF      <- weather$PF[day + 1]
  GPP_ref <- weather$GPP_ref[day + 1]

  ## Temperature-driven growth function (g)
  if (TAir > 0) {
    g <- 1 / (1 + exp(-common$a * (TAir - common$b)))
    if (g < 0) g <- 0
  } else {
    g <- 0
  }
  g_sD_T <- g

  ## Diameter state (sD)
  if (day == 0) {
    sD <- parameters$sD0_Trad
  } else if (day <= parameters$diameter_start_day) {
    sD <- state$sD
  } else {
    sD <- state$sD + g_sD_T
  }

  ## Potential diameter growth shape function (fD)
  if (sD > 0 && sD < parameters$sDc^2) {
    fD <- (parameters$sDc * sqrt(sD) - sD) / (parameters$sDc^2 / 4)
  } else {
    fD <- 0
  }

  ## Growth driver (GD)
  LD <- parameters$LD0 * ratio$diameter_growth_coefficient
  GD <- g_sD_T * fD * LD

  ## Early/late wood switch (Uggla threshold)
  if (sD < parameters$sDc^2 / common$Uggla) {
    tau_E <- parameters$tau_Ee
    tau_W <- parameters$tau_We
  } else {
    tau_E <- parameters$tau_El
    tau_W <- parameters$tau_Wl
  }

  ## Cell pools
  if (day < 1) {
    n_E_pot <- n_W_pot <- n_M_pot <- 0
  } else {
    n_E_pot <- state$n_E_pot + GD - state$n_E_pot / tau_E
    n_W_pot <- state$n_W_pot + state$n_E_pot / tau_E -
               state$n_W_pot / state$tau_W
    n_M_pot <- state$n_M_pot + state$n_W_pot / state$tau_W
  }

  ## Carbon cost of cell enlargement
  CE <- function(cell_d) {
    common$osmotic_sugar_conc * pi * (cell_d / 2)^2 * 0.00266 *
      342.296 / (common$gas_const * (TAir + common$abs_zero) * 1000) *
      12 * common$M_C / 342.296 / tau_E
  }
  CE_ew <- CE(parameters$cell_d_ew)
  CE_lw <- CE(parameters$cell_d_lw)

  n_rows <- ratio$form_factor * parameters$h0 / 0.00266 * pi * 0.175 /
            parameters$cell_d_ew

  carbon_enlargement_pot <- if (sD < parameters$sDc^2 / common$Uggla) {
    CE_ew * n_E_pot
  } else {
    CE_lw * n_E_pot
  }

  ## Wall formation carbon cost
  CW <- 1.8e-11
  wall_pot_growth <- n_rows * CW * n_W_pot

  ## Cell counts and ring width
  cells_pot <- n_W_pot + n_M_pot

  if (sD < parameters$sDc^2 / common$Uggla) {
    ew_cells_pot     <- cells_pot
    ew_cells_pot_max <- max(ew_cells_pot, state$ew_cells_pot_max)
    lw_cells_pot     <- 0
  } else {
    ew_cells_pot     <- 0
    ew_cells_pot_max <- state$ew_cells_pot_max
    lw_cells_pot     <- cells_pot - ew_cells_pot - ew_cells_pot_max
  }

  ew_width_pot <- ew_cells_pot * parameters$cell_d_ew * 1000
  lw_width_pot <- lw_cells_pot * parameters$cell_d_lw * 1000

  if (sD < parameters$sDc^2 / common$Uggla) {
    pot_mm     <- ew_width_pot
    pot_mm_max <- max(pot_mm, state$pot_mm_max)
  } else {
    pot_mm_max <- state$pot_mm_max
    pot_mm     <- lw_width_pot + pot_mm_max
  }

  list(
    sD = sD, GD = GD, g = g, fD = fD,
    n_E_pot = n_E_pot, n_W_pot = n_W_pot, n_M_pot = n_M_pot,
    tau_E = tau_E, tau_W = tau_W,
    ew_cells_pot_max = ew_cells_pot_max,
    pot_mm_max = pot_mm_max, pot_mm = pot_mm
  )
}

run_growth <- function(weather, parameters, common, ratio) {
  n_days   <- nrow(weather)
  states   <- vector("list", n_days)
  states[[1]] <- init_state
  for (d in 1:(n_days - 1)) {
    states[[d + 1]] <- grow_one_day(
      day = d, weather = weather, state = states[[d]],
      parameters = parameters, common = common, ratio = ratio
    )
  }
  states
}
```

## Run and collect output

```{r}
states <- run_growth(weather = weather, parameters = parameters,
                     common = common, ratio = ratio)

output <- data.frame(
  day     = weather$day,
  g       = vapply(states, function(x) x$g,       numeric(1)),
  sD      = vapply(states, function(x) x$sD,      numeric(1)),
  fD      = vapply(states, function(x) x$fD,      numeric(1)),
  GD      = vapply(states, function(x) x$GD,      numeric(1)),
  n_E_pot = vapply(states, function(x) x$n_E_pot, numeric(1)),
  n_W_pot = vapply(states, function(x) x$n_W_pot, numeric(1)),
  n_M_pot = vapply(states, function(x) x$n_M_pot, numeric(1)),
  tau_W   = vapply(states, function(x) x$tau_W,   numeric(1)),
  tau_E   = vapply(states, function(x) x$tau_E,   numeric(1)),
  pot_mm  = vapply(states, function(x) x$pot_mm,  numeric(1)),
  TAir    = weather$TAir
)
```

---

## Step 1: Temperature-driven growth response (g)

`g` is a logistic function of air temperature, controlled by the shape
parameters `a` and `b` in `common`. It is zero when temperature is at or below
zero, and approaches 1 at high temperatures. This represents the physiological
readiness of the cambium to divide — no growth is possible when the tree is too
cold. The `b` parameter sets the inflection temperature (default 5 °C), and `a`
controls the steepness of the response.

```{r}
plot(output$day, output$g, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "g",
     main = "Temperature-driven growth response (g)")
title(sub = "Controlled by common parameters: a (steepness), b (inflection temperature)")
mtext("g > 0 whenever growth is temperature-permissive", side = 3, line = 0.5, cex = 0.9)
```

---

## Step 2: Cumulative growth state (sD)

`sD` accumulates `g` each day, but only after `diameter_start_day`. This
represents the cumulative thermal stimulus driving cambial activity — analogous
to growing-degree days. The starting value `sD0_Trad` can be negative,
reflecting a "debt" of thermal stimulus that must be overcome before growth
begins. Growth only occurs while `sD` is between 0 and `sDc²` (shown by the
horizontal line); outside that range the potential growth function `fD` returns
zero.

```{r}
plot(output$day, output$sD, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "sD",
     main = "Cumulative thermal growth state (sD)")
abline(h = parameters$sDc^2, lty = 2, col = "red")
title(sub = "Controlled by: diameter_start_day, sD0_Trad")
mtext("Diameter growth is active when sD is between 0 and the dashed line", side = 3, line = 0.5, cex = 0.9)
```

---

## Step 3: Potential growth shape function (fD)

`fD` is a bell-shaped function of `sD`, peaking in the middle of the growing
season and returning to zero at the ends. It is parameterised by `sDc`, which
controls both the width and height of the bell. A larger `sDc` gives a broader,
taller growth pulse (more total growth over the season). `fD` is a multiplier
on the growth driver `GD` — it ensures that the growth rate is highest at the
peak of the season and tapers off naturally.

```{r}
plot(output$day, output$fD, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "fD",
     main = "Potential growth shape function (fD)")
title(sub = "Controlled by: sDc")
mtext("fD is the seasonal modulator on GD", side = 3, line = 0.5, cex = 0.9)
```

---

## Step 4: Daily growth driver (GD)

`GD` is the product of the temperature response `g`, the seasonal shape `fD`,
and a scaling factor `LD` (itself derived from `LD0` and the
`diameter_growth_coefficient`). It represents the daily input of new potential
cells into the enlargement pool. `GD` has a single peak in early summer when
both temperature and the seasonal bell function are near their maxima.

```{r}
plot(output$day, output$GD, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "GD",
     main = "Daily growth driver (GD)")
title(sub = "Controlled by: diameter_growth_coefficient, LD0")
mtext("GD is the daily input to the enlargement cell pool (n_E_pot)", side = 3, line = 0.5, cex = 0.9)
```

---

## Step 5: Cell transit times and the early/late wood switch

Cells move through three pools — enlargement (n_E_pot), wall thickening
(n_W_pot), and maturation (n_M_pot) — each with a residence time `tau_E` and
`tau_W` respectively. The Uggla threshold (`sD < sDc² / Uggla`) determines
whether early-wood or late-wood cell dimensions and transit times apply. When
`sD` crosses this threshold, both `tau_E` and `tau_W` switch from their
early-wood values (`tau_Ee`, `tau_We`) to their late-wood values (`tau_El`,
`tau_Wl`). Late-wood cells are smaller and have shorter transit times, producing
the denser, narrower cells characteristic of late-season growth.

```{r, fig.width=8, fig.height=4}
par(mfrow = c(1, 2))
plot(output$day, output$tau_E, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "tau_E",
     main = "Enlargement transit time (tau_E)")
title(sub = "tau_Ee → tau_El at the Uggla threshold")
mtext("Controls rate of flow from n_E_pot to n_W_pot", side = 3, line = 0.5, cex = 0.9)

plot(output$day, output$tau_W, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "tau_W",
     main = "Wall-thickening transit time (tau_W)")
title(sub = "tau_We → tau_Wl at the Uggla threshold")
mtext("Controls rate of flow from n_W_pot to n_M_pot", side = 3, line = 0.5, cex = 0.9)
par(mfrow = c(1, 1))
```

---

## Step 6: Cell pools

The three cell pools track how many cells are currently in each stage of
development. `n_E_pot` (enlargement) is fed by `GD` and drained into `n_W_pot`
(wall thickening) at rate `1/tau_E`. `n_W_pot` is drained into `n_M_pot`
(maturation) at rate `1/tau_W`. This creates a lagged, smoothed response to the
growth driver — cells produced at the peak of summer continue moving through the
pipeline for weeks afterwards, which is why ring width measurements often show
later peaks than temperature alone would predict.

```{r, fig.width=10, fig.height=4}
par(mfrow = c(1, 3))
plot(output$day, output$n_E_pot, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "n_E_pot",
     main = "Enlargement pool (n_E_pot)")
title(sub = "Driven by GD, drained by tau_E")
mtext("Carbon cost: cell enlargement", side = 3, line = 0.5, cex = 0.9)

plot(output$day, output$n_W_pot, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "n_W_pot",
     main = "Wall-thickening pool (n_W_pot)")
title(sub = "Fed from n_E_pot, drained by tau_W")
mtext("Carbon cost: wall formation", side = 3, line = 0.5, cex = 0.9)

plot(output$day, output$n_M_pot, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "n_M_pot",
     main = "Mature cell pool (n_M_pot)")
title(sub = "Fed from n_W_pot")
mtext("Contributes to total cell count (cells_pot)", side = 3, line = 0.5, cex = 0.9)
par(mfrow = c(1, 1))
```

---

## Step 7: Potential ring width (pot_mm)

The final output is `pot_mm`, the cumulative potential ring width in millimetres.
During the early-wood phase, it grows as early-wood cells (diameter `cell_d_ew`)
accumulate. Once the Uggla threshold is crossed, late-wood cells (smaller
diameter `cell_d_lw`) begin forming, and `pot_mm` continues to increase but more
slowly. The value at the end of the year is the model's prediction of annual ring
width under these conditions.

Note that this is the *potential* ring width — it assumes no carbon limitation.
In the full `CASSIA_cpp()` model, actual ring width is additionally constrained
by available carbon, linking it to the sugar and photosynthesis sub-models.

```{r}
par(mfrow = c(1, 1))
plot(output$day, output$pot_mm, type = "l", lwd = 2,
     xlab = "Day of year", ylab = "Potential ring width (mm)",
     main = "Potential ring width over the growing season")
title(sub = "Controlled by sDc, Uggla, cell_d_ew, cell_d_lw")
mtext("Early wood (fast increase) transitions to late wood (slower increase) at the Uggla threshold",
      side = 3, line = 0.5, cex = 0.9)
```

---

## Summary of the cell pipeline

```
Temperature (TAir)
      │
      ▼
   g (logistic response)
      │
      ▼
   sD (cumulative thermal state)  ──── diameter_start_day, sD0_Trad
      │
      ▼
   fD (seasonal bell function)    ──── sDc
      │
      ▼
   GD (daily growth driver)       ──── LD0, diameter_growth_coefficient
      │
      ▼
  n_E_pot (enlargement)           ──── tau_Ee / tau_El (Uggla threshold)
      │
      ▼
  n_W_pot (wall thickening)       ──── tau_We / tau_Wl
      │
      ▼
  n_M_pot (mature)
      │
      ▼
  pot_mm (ring width)             ──── cell_d_ew, cell_d_lw, Uggla
```

To understand how a parameter affects ring width, trace it forward through this
pipeline. For example, `sDc` affects `fD`, which affects `GD`, which drives all
three cell pools, and finally `pot_mm`. `tau_Ee` only affects the lag between
`n_E_pot` and `n_W_pot` — it shifts the timing of wall growth without changing
the total cells produced.

See `vignette("Sensitivity_Analysis")` for how to quantify these effects
systematically.
