---
title: "Running the CASSIA Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the CASSIA Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup

Install the package from GitHub if you haven't already, then load it.

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("josimms/CASSIA")
```

```{r}
library(CASSIA)
```

## Run the model

The package includes built-in weather data for Hyytiälä (2015–2017) so you can
run a basic simulation straight away. The model produces results comparable to
Schiestl-Aalto et al. (2015).

```{r}
# Built-in Hyytiälä weather data (2015-2017)
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

# Run the model
out <- CASSIA_cpp(weather = weather_original, site = "Hyde")
```

The output is a list with three dataframes:

- `out$Growth` — daily carbon fluxes and organ growth
- `out$Sugar`  — daily sugar and starch pool values (when sugar model is active)
- `out$Preles` — daily photosynthesis outputs (GPP, ET, soil water)

```{r}
head(out$Growth)
```

## Changing model settings

Settings are controlled by toggle arguments. For example, `LN_estim = TRUE`
(the default) scales needle elongation by the previous year's GPP. To turn it
off:

```{r}
out2 <- CASSIA_cpp(weather = weather_original, site = "Hyde", LN_estim = FALSE)
```

See `?CASSIA_cpp` for the full list of toggles and what they do.

## Changing parameters

Parameters are stored in dataframes (`ratios_p`, `parameters_p`, `common_p`,
`sperling_p`, `repo_p`). The easiest way to modify them is to copy a default
dataframe and change the value you want:

```{r}
parameters_updated <- parameters_p
parameters_updated["root.lifetime", "Hyde"] <- 2

out3 <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                   parameters = parameters_updated)
```

The meaning of each parameter is explained in the CASSIA instruction booklet
(`man/CASSIA+_Instruction_And_Plans_Booklet.pdf`).

## Weather data format

The required columns depend on which photosynthesis sub-model is active.
Columns not needed by the active sub-model may be `NA`.

**Always required:**

| Column | Description | Units |
|--------|-------------|-------|
| `dates` | Date (class `Date`) | YYYY-MM-DD |
| `T` | Air temperature | °C |
| `TSA` | Soil temperature, A horizon | °C |
| `TSB` | Soil temperature, B horizon | °C |
| `MB` | Soil moisture, B horizon | m³ m⁻³ |
| `Rain` | Precipitation | mm day⁻¹ |

**Additional columns by sub-model:**

| Sub-model toggle | Extra columns needed | Units |
|-----------------|----------------------|-------|
| `photosynthesis_as_input = TRUE` *(default)* | `P` — GPP per tree | g C m⁻² day⁻¹ |
| `PRELES_GPP = TRUE` | `PAR`, `VPD`, `CO2`, `fAPAR` | mol m⁻² day⁻¹, kPa, ppm, 0–1 |

The built-in `weather_original` includes all columns used by any sub-model:

```{r}
names(weather_original)
```

See `?CASSIA_cpp` for the full parameter list and `vignette("Weather")` for
guidance on preparing your own weather data.
