---
title: "Running the CASSIA Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the CASSIA Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup

Install the package from GitHub if you haven't already, then load it.

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("josimms/CASSIA")
```

```{r}
library(CASSIA)
```

## Run the model

The package includes built-in weather data for Hyytiälä (2015–2017) so you can
run a basic simulation straight away. The model produces results comparable to
Schiestl-Aalto et al. (2015).

```{r}
# Built-in Hyytiälä weather data (2015-2017)
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

# Run the model
out <- CASSIA_cpp(weather = weather_original, site = "Hyde")
```

The output is a list with three dataframes:

- `out$Growth` — daily carbon fluxes and organ growth
- `out$Sugar`  — daily sugar and starch pool values (when sugar model is active)
- `out$Preles` — daily photosynthesis outputs (GPP, ET, soil water)

```{r}
head(out$Growth)
```

## Changing model settings

Settings are controlled by toggle arguments. For example, `LN_estim = TRUE`
(the default) scales needle elongation by the previous year's GPP. To turn it
off:

```{r}
out2 <- CASSIA_cpp(weather = weather_original, site = "Hyde", LN_estim = FALSE)
```

See `?CASSIA_cpp` for the full list of toggles and what they do.

## Changing parameters

Parameters are stored in dataframes (`ratios_p`, `parameters_p`, `common_p`,
`sperling_p`, `repo_p`). The easiest way to modify them is to copy a default
dataframe and change the value you want:

```{r}
parameters_updated <- parameters_p
parameters_updated["root.lifetime", "Hyde"] <- 2

out3 <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                   parameters = parameters_updated)
```

The meaning of each parameter is explained in the CASSIA instruction booklet
(`CASSIA_Instruction_Booklet.pdf` in the package root, or on GitHub).

## Weather data format

The weather dataframe must have these columns (column `X` is the day-of-year):

```{r}
names(weather_original)
```

If you are using your own weather data, match these column names exactly. The
units are: T (°C), P (g C m-2 day-1), TSA/TSB (°C), MB (m3 m-3), Rain
(mm day-1). The `dates` column must be class `Date`.

The model will give an informative error if the weather data is wrongly named
or contains NAs. See `vignette("Sensitivity_Analysis")` for guidance on fitting
the model to a new site.

## Experimental features (this branch)

This branch (`adding_externals`) includes features not yet in the stable
release. All are off by default and enabled by toggles:

| Toggle | What it enables |
|--------|----------------|
| `sperling_model = TRUE` | Enzyme-driven sugar model with organ-level concentration pools |
| `phydro = TRUE` | GPP calculated via the p-hydro hydraulic model |
| `myco_model = TRUE` | Experimental mycorrhizal carbon/nitrogen exchange model |
| `xylogenesis = TRUE` | Xylogenesis sub-model with early/late-wood cell stages |
| `nitrogen_change = TRUE` | Nitrogen addition treatment |

The sugar model also exposes additional numeric parameters — `ratio_sugar`,
`tau.myco`, `tau.t.needles`, `tau.t.phloem`, `tau.t.roots`,
`tau.t.xylem.sh`, `tau.t.xylem.st` — which control transfer rates between
organ pools. Their defaults are chosen for Hyytiälä; see `?CASSIA_cpp` for
the full list and meanings.

```{r}
out_sugar <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                        sperling_model = TRUE)
head(out_sugar$Sugar)
```
