---
title: "Running the CASSIA Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the CASSIA Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup

Install the package from GitHub if you haven't already, then load it.

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("josimms/CASSIA")
```

```{r}
library(CASSIA)
```

## Run the model

The package includes built-in weather data for Hyytiälä (2015–2017) so you can
run a basic simulation straight away. The model produces results comparable to
Schiestl-Aalto et al. (2015).

```{r}
# Built-in Hyytiälä weather data (2015-2017)
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

# Run the model
out <- CASSIA_cpp(weather = weather_original, site = "Hyde")
```

The output is a list with three dataframes:

- `out$Growth` — daily carbon fluxes and organ growth
- `out$Sugar`  — daily sugar and starch pool values (when sugar model is active)
- `out$Preles` — daily photosynthesis outputs (GPP, ET, soil water)

```{r}
head(out$Growth)
```

## Changing model settings

Settings are controlled by toggle arguments. For example, `LN_estim = TRUE`
(the default) scales needle elongation by the previous year's GPP. To turn it
off:

```{r}
out2 <- CASSIA_cpp(weather = weather_original, site = "Hyde", LN_estim = FALSE)
```

See `?CASSIA_cpp` for the full list of toggles and what they do.

## Changing parameters

Parameters are stored in dataframes (`ratios_p`, `parameters_p`, `common_p`,
`sperling_p`, `repo_p`). The easiest way to modify them is to copy a default
dataframe and change the value you want:

```{r}
parameters_updated <- parameters_p
parameters_updated["root.lifetime", "Hyde"] <- 2

out3 <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                   parameters = parameters_updated)
```

The meaning of each parameter is explained in the CASSIA instruction booklet
(`CASSIA_Instruction_Booklet.pdf` in the package root, or on GitHub).

## Weather data format

The required columns depend on which photosynthesis sub-model is active.
Columns not needed by the active sub-model do not need to be present.

**Always required:**

| Column | Description | Units |
|--------|-------------|-------|
| `dates` | Date (class `Date`) | YYYY-MM-DD |
| `T` | Air temperature | °C |
| `TSA` | Soil temperature, A horizon | °C |
| `TSB` | Soil temperature, B horizon | °C |
| `MB` | Soil moisture, B horizon | m³ m⁻³ |
| `Rain` | Precipitation | mm day⁻¹ |

**Additional columns by sub-model:**

| Sub-model toggle | Extra columns needed | Units |
|-----------------|----------------------|-------|
| `photosynthesis_as_input = TRUE` *(default)* | `P` — GPP per tree | g C m⁻² day⁻¹ |
| `preles = TRUE` | `PAR`, `VPD`, `CO2`, `fAPAR` | mol m⁻² day⁻¹, kPa, ppm, 0–1 |
| `phydro = TRUE` | `PPFD`, `PPFD_max`, `VPD`, `CO2`, `Nitrogen`, `PA`, `SWP` | µmol m⁻² s⁻¹, µmol m⁻² s⁻¹, kPa, ppm, g m⁻², Pa, MPa |

The built-in `weather_original` covers the default and PRELES sub-models.
The phydro sub-model requires additional columns (`PPFD`, `PPFD_max`, `PA`,
`SWP`) that are not included. Current columns:

```{r}
names(weather_original)
```

See `?CASSIA_cpp` for the full parameter list and `vignette("Sensitivity_Analysis")`
for guidance on fitting the model to a new site.

## Optional sub-models

The following features are off by default and enabled by toggles:

| Toggle | What it enables |
|--------|----------------|
| `organ_level_sugar = TRUE` | Organ-level sugar model: per-organ concentration pools with separate sugar/starch dynamics |
| `phydro = TRUE` | GPP calculated via the p-hydro hydraulic model |
| `myco_model = TRUE` | Experimental mycorrhizal carbon/nitrogen exchange model |
| `xylogenesis = TRUE` | Xylogenesis sub-model with early/late-wood cell stages |
| `nitrogen_change = TRUE` | Nitrogen addition treatment |

The sugar model also exposes additional numeric parameters — `ratio_sugar`,
`tau.myco`, `tau.t.needles`, `tau.t.phloem`, `tau.t.roots`,
`tau.t.xylem.sh`, `tau.t.xylem.st` — which control transfer rates between
organ pools. Their defaults are chosen for Hyytiälä; see `?CASSIA_cpp` for
the full list and meanings.

```{r}
out_sugar <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                        organ_level_sugar = TRUE)
head(out_sugar$Sugar)
```
