---
title: "Sensitivity Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sensitivity Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
library(dplyr)
```

## Why run a sensitivity analysis?

CASSIA has many parameters, most of which come pre-calibrated for Hyytiälä
(`"Hyde"`). If you want to use the model at a new site, you first need to know
which parameters actually matter for the output you care about. A sensitivity
analysis answers this by running the model repeatedly, varying one parameter at
a time across a plausible range, and observing how much the output changes.

Parameters that produce large output changes are the ones worth calibrating
carefully with observations. Parameters that have little effect can be left at
their default values.

## Setup weather data

The built-in Hyytiälä weather data covers 2015–2017 and is a reasonable
starting point for a sensitivity analysis even if your site is different — the
goal at this stage is to rank parameter importance, not to match observations.

```{r}
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))
```

## The sensitivity function

The function below runs the model once for each value in `range`, changing only
the named parameter while keeping all others at their defaults. It returns a
list of results, one per run.

```{r}
run_sensitivity <- function(param_name, range, weather) {
  lapply(range, function(i) {
    parameters_p_new <- parameters_p
    if (param_name %in% rownames(parameters_p_new)) {
      parameters_p_new["Hyde"][rownames(parameters_p_new) == param_name, ] <- i
      result <- CASSIA_cpp(weather, site = "Hyde",
                           parameters = parameters_p_new)
      list(param_value = i, result = result)
    } else {
      message("Parameter '", param_name,
              "' not found. Check spelling — underscore vs dot matters.")
      NULL
    }
  })
}
```

## Choosing which parameters to test

Start with the parameters that control the timing and magnitude of diameter
growth, as these have the most direct link to ring width observations:

| Parameter | What it controls | Typical range |
|-----------|-----------------|---------------|
| `sDc` | Peak potential diameter growth rate | 0 – 10 |
| `sD0.Trad` | Starting value of the temperature-driven growth state | −7 – 0 |
| `diameter_start_day` | Day of year when diameter growth begins accumulating | 50 – 200 |

`sDc` controls the overall magnitude of the growing season growth pulse.
`sD0.Trad` shifts when that pulse begins relative to the spring temperature
signal. `diameter_start_day` sets the earliest possible start of growth
regardless of temperature.

```{r, eval=FALSE}
sensitivity_sDc      <- run_sensitivity("sDc",           seq(0, 10, length.out = 10), weather_original)
sensitivity_sD0_Trad <- run_sensitivity("sD0.Trad",      seq(-7, 0,  length.out = 10), weather_original)
sensitivity_dgs      <- run_sensitivity("diameter_start_day", seq(50, 200, length.out = 10), weather_original)
```

## Plotting the results

Extract the annual peak ring width across all runs. Here we use
`ring_width` (cumulative ring width within the year, mm), taking the maximum
per year as the final annual value. You could substitute any column from
`out$Growth`.

```{r, eval=FALSE}
extract_yearly <- function(sensitivity_list, param_name, output_col, fun = max) {
  bind_rows(lapply(sensitivity_list, function(x) {
    if (is.null(x)) return(NULL)
    daily <- x$result$Growth
    data.frame(
      param_name  = param_name,
      param_value = x$param_value,
      year        = unique(daily$year),
      value       = tapply(daily[[output_col]], daily$year, fun)
    )
  }))
}

plot_data <- bind_rows(
  extract_yearly(sensitivity_sDc,      "sDc",                "ring_width"),
  extract_yearly(sensitivity_sD0_Trad, "sD0.Trad",           "ring_width"),
  extract_yearly(sensitivity_dgs,      "diameter_start_day", "ring_width")
)

library(ggplot2)
ggplot(plot_data, aes(x = param_value, y = value, colour = factor(year))) +
  geom_line() +
  facet_wrap(~param_name, scales = "free_x") +
  labs(y = "Annual ring width (mm)", colour = "Year")
```

## Interpreting the results

A steep slope in a sensitivity plot means a small change in that parameter
produces a large change in output — this parameter is important to get right.
A flat line means the output is insensitive to that parameter and it can be
left at its default.

Also check whether the sensitivity is consistent across years (coloured lines
parallel) or interacts with inter-annual climate variability (lines crossing).
Year-dependent sensitivity usually means the parameter interacts with a
temperature or drought signal and should be treated with extra care when
calibrating.

## Next steps: calibration

Once you know which parameters matter, calibrate them by comparing modelled
outputs against observations. Useful observation types for the diameter growth
parameters above are:

- **Dendrometer data**: daily or weekly diameter measurements — compare directly with cumulative growth from `out$Growth`
- **Ring width measurements**: annual ring widths — compare with annual peak `ring_width`
- **Onset of cambial activity**: phenological observations of when growth begins — compare with the day when `ring_width` first exceeds zero

See `vignette("ring_width_process")` for a step-by-step walkthrough of how
CASSIA models ring width formation internally.
