---
title: "Visualising CASSIA Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising CASSIA Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
```

## Run the model

The examples below use the built-in Hyytiälä weather data (2015–2017) and the
matching observation dataset `Hyde_daily_original`.

```{r}
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

out <- CASSIA_cpp(weather = weather_original, site = "Hyde")
```

`CASSIA_cpp` returns a named list:

| Element | Contents |
|---------|----------|
| `out$Growth` | Daily carbon fluxes and organ growth (one row per day) |
| `out$Sugar` | Daily sugar and starch pools (when `organ_level_sugar = TRUE`) |
| `out$Preles` | Daily photosynthesis output: GPP, ET, soil water |
| `out$Culm_Growth` | Cumulative growth totals |

---

## plot_weather_variables

`plot_weather_variables` produces a multi-panel overview of the weather drivers.
It accepts any data frame that has a `date` (or `Date`) column.

```{r, fig.width=7, fig.height=5}
# Add date column if not already present
weather_original$date <- weather_original$dates
plot_weather_variables(weather_original)
```

---

## Inspecting output manually

The Growth and Preles dataframes can be plotted with base R. The columns most
commonly compared to observations are:

```{r, fig.width=7, fig.height=4}
dates <- weather_original$dates

par(mfrow = c(1, 3))
plot(dates, out$Growth$needle_growth,
     type = "l", main = "Needle growth",
     xlab = "Date", ylab = "kg C tree⁻¹ day⁻¹")
plot(dates, out$Growth$root_growth,
     type = "l", main = "Root growth",
     xlab = "Date", ylab = "kg C tree⁻¹ day⁻¹")
plot(dates, out$Preles$GPP,
     type = "l", main = "GPP (PRELES)",
     xlab = "Date", ylab = "g C m⁻² day⁻¹")
```

---

## plot_comparison

`plot_comparison` overlays model output on observations and adds a residual
panel for each variable. Because it compares row-by-row, the model output and
observation data must span the same dates.

`Hyde_daily_original` covers 2015–2018 (four years); the run above covers
2015–2017, so we subset to match:

```{r, fig.width=7, fig.height=7}
# Subset observations to the same three years
hyde_obs <- Hyde_daily_original[Hyde_daily_original$year %in% 2015:2017, ]

# Model column names → observation column names
variables_new      <- c("bud_growth", "needle_growth", "root_growth")
variables_original <- c("bud",        "needle_daily",  "root_daily")

plot_comparison(out, variables_new, hyde_obs, variables_original)
```

The three-column layout for each variable shows: time series overlay,
model-vs-observed scatter, and residuals. A point on the 1:1 line in the
scatter panel (red) means perfect agreement.

---

## plot_sugar_starch_comparison

The sugar/starch comparison overlays measured concentration data on the modelled
organ pools. Run the model with the sugar sub-model active first:

```{r}
out_sugar <- CASSIA_cpp(weather = weather_original, site = "Hyde",
                        organ_level_sugar = TRUE)
names(out_sugar$Sugar)
```

The function also requires Hyytiälä sugar measurements (`original_data`) and Yu
et al. needle data (`yu_data`), which are loaded via `load_data()` from the
calibration data directory. If you have that directory set up:

```{r, eval=FALSE}
loaded <- load_data()   # see ?load_data for the required directory

dates <- weather_original$dates
plot_sugar_starch_comparison(out_sugar, dates,
                             original_data = loaded$original_data,
                             yu_data       = loaded$yu_data)
```

Without the calibration files you can still plot the modelled pools directly:

```{r, fig.width=7, fig.height=4}
sugar_total <- rowSums(out_sugar$Sugar[, c("sugar_needles", "sugar_phloem",
                                            "sugar_roots", "sugar_xylem_sh",
                                            "sugar_xylem_st")])
starch_total <- rowSums(out_sugar$Sugar[, c("starch_needles", "starch_phloem",
                                             "starch_roots", "starch_xylem_sh",
                                             "starch_xylem_st")])

par(mfrow = c(1, 2))
plot(dates, sugar_total, type = "l", col = "blue",
     main = "Total sugar (all organs)", xlab = "Date", ylab = "kg C tree⁻¹")
plot(dates, starch_total, type = "l", col = "brown",
     main = "Total starch (all organs)", xlab = "Date", ylab = "kg C tree⁻¹")
```
