---
title: "Preparing Weather Data for CASSIA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preparing Weather Data for CASSIA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(CASSIA)
```

## What CASSIA needs

Which columns are required depends on the photosynthesis sub-model selected
via the toggle arguments to `CASSIA_cpp()`. Columns not used by the active
sub-model may be `NA`.

**Always required:**

| Column | Description | Units | Notes |
|--------|-------------|-------|-------|
| `dates` | Calendar date | `Date` class | Created from year + `X` |
| `X` | Day of year | integer (1–366) | |
| `T` | Air temperature | °C | Typically at 2–4 m height |
| `TSA` | Soil temperature, A horizon | °C | ~5 cm depth |
| `TSB` | Soil temperature, B horizon | °C | ~10 cm depth |
| `MB` | Volumetric soil moisture, B horizon | m³ m⁻³ | |
| `Rain` | Precipitation | mm day⁻¹ | Daily total |

**Additional columns by photosynthesis sub-model:**

| Toggle | Extra columns required | Units |
|--------|------------------------|-------|
| `photosynthesis_as_input = TRUE` *(default)* | `P` — GPP per tree | g C m⁻² day⁻¹ |
| `preles = TRUE` | `PAR`, `VPD`, `CO2`, `fAPAR` | mol m⁻² day⁻¹, kPa, ppm, 0–1 |
| `phydro = TRUE` | `PPFD`, `PPFD_max`, `VPD`, `CO2`, `Nitrogen`, `PA`, `SWP` | µmol m⁻² s⁻¹, µmol m⁻² s⁻¹, kPa, ppm, g m⁻², Pa, MPa |

The built-in `weather_original` includes all columns for all sub-models.
The model will give an informative error if column names are wrong, values
are outside expected bounds, or required columns contain `NA`s.

## Option 1: Use the built-in Hyytiälä data

The fastest way to get started is to use `weather_original`, which contains
pre-processed daily weather for Hyytiälä (SMEAR II station) covering 2015–2017.
It just needs a `dates` column added:

```{r}
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

str(weather_original)
```

All the columns CASSIA needs are already present. See `vignette("Running_The_Model")`
for how to use this to run your first simulation.

---

## Option 2: Local weather station data

If you have your own station data, the general pipeline is:

1. **Download** raw sub-daily records (typically hourly or half-hourly)
2. **Aggregate** to daily means (or sums for precipitation)
3. **Rename** columns to match CASSIA's expected names
4. **Gap-fill** any missing values
5. **Validate** with plots before running the model

### Downloading from the SMEAR portal (Hyytiälä users)

The SMEAR II station data is available via a REST API. The URL pattern below
downloads daily records for a single variable:

```{r, eval=FALSE}
# Replace HYY_META.T168 with the variable you need, and adjust the date range
url <- paste0(
  "https://smear-backend.rahtiapp.fi/search/timeseries/csv",
  "?tablevariable=HYY_META.T168",
  "&from=2015-01-01T00:00:00.000",
  "&to=2017-12-31T23:59:59.999",
  "&quality=ANY&aggregation=NONE&interval=1"
)
raw <- read.csv(url)
```

Key SMEAR variables and their CASSIA equivalents:

| SMEAR variable | CASSIA column | Aggregation |
|---------------|---------------|-------------|
| `HYY_META.T168` | `T` | daily mean |
| `HYY_META.tsoil_5` | `TSA` | daily mean |
| `HYY_META.tsoil_10` | `TSB` | daily mean |
| `HYY_META.wpsoil_B` | `MB` | daily mean |
| `HYY_META.Precip` | `Rain` | daily sum |
| `HYY_EDDY233.GPP` | `P` | daily mean |

### Aggregating to daily values

```{r, eval=FALSE}
library(data.table)

# fread() returns a data.table directly, which supports the .() aggregation syntax
raw <- fread(url)
raw$date <- as.Date(paste(raw$Year, raw$Month, raw$Day, sep = "-"))

daily <- raw[, .(
  T    = mean(HYY_META.T168,    na.rm = TRUE),
  TSA  = mean(HYY_META.tsoil_5, na.rm = TRUE),
  TSB  = mean(HYY_META.tsoil_10, na.rm = TRUE),
  MB   = mean(HYY_META.wpsoil_B, na.rm = TRUE),
  Rain = sum(HYY_META.Precip,   na.rm = TRUE)
), by = date]

daily$X     <- as.integer(format(daily$date, "%j"))
daily$dates <- daily$date
```

### Gap-filling missing values

CASSIA will error on `NA`s. A simple approach for isolated gaps is to fill
with the long-term mean for that day of year (mean-bias imputation):

```{r, eval=FALSE}
# Calculate the day-of-year mean across all years
doy_mean <- tapply(daily$T, daily$X, mean, na.rm = TRUE)

# Fill gaps
daily$T[is.na(daily$T)] <- doy_mean[daily$X[is.na(daily$T)]]
```

For longer gaps (> 1–2 weeks), consider using ERA5 reanalysis data to fill
in the missing period and then bias-correct as described in Option 3 below.

---

## Option 3: ERA5 reanalysis data

ERA5 is useful when you have no local station data, or when you need long
historical records. It is globally available at 0.25° resolution from the
[Copernicus Climate Data Store](https://cds.climate.copernicus.eu/).

The ERA5 pipeline in `R/ERAS_weather_processing.R` covers this in full.
The key steps are:

### 1. Download ERA5 NetCDF files

Download the following variables at your site coordinates using the CDS API
or web interface: `t2m` (2 m temperature), `ssrd` (surface solar radiation
downwards), `swvl2` (volumetric soil water layer 2), `d2m` (2 m dew point).

### 2. Extract and convert units

```{r, eval=FALSE}
library(ncdf4)
library(data.table)

nc <- nc_open("download.nc")

# Extract at your site (adjust lon/lat indices)
lon_index <- which.min(abs(ncvar_get(nc, "longitude") - YOUR_LON))
lat_index <- which.min(abs(ncvar_get(nc, "latitude")  - YOUR_LAT))

daily <- data.table(
  T    = ncvar_get(nc, "t2m",  start = c(lon_index, lat_index, 1), count = c(1,1,-1)) - 273.15,
  MB   = ncvar_get(nc, "swvl2", start = c(lon_index, lat_index, 1), count = c(1,1,-1)),
  PAR  = bigleaf::Rg.to.PPFD(
           ncvar_get(nc, "ssrd", start = c(lon_index, lat_index, 1), count = c(1,1,-1)) / 3600)
)
nc_close(nc)
```

ERA5 does not provide soil temperature at the CASSIA depths. A practical
substitute is to smooth the 2 m air temperature with a lag to approximate
the thermal damping at depth:

```{r, eval=FALSE}
# Simple lagged smoother — adjust lag and width to match your soil depth
daily$TSA <- stats::filter(daily$T, rep(1/7,  7),  sides = 1)  # ~5 cm depth
daily$TSB <- stats::filter(daily$T, rep(1/14, 14), sides = 1)  # ~10 cm depth
```

### 3. Bias correct against local observations

ERA5 has systematic biases at fine spatial scales. Correct by calculating
the monthly mean error between ERA5 and any local observations you have,
then subtracting it from the ERA5 record:

```{r, eval=FALSE}
library(dplyr)

# 'obs' is a dataframe of local observations; 'era' is the ERA5 data
# Both must have Month and T columns
obs_monthly <- obs  %>% group_by(Month) %>% summarise(T_obs  = mean(T, na.rm = TRUE))
era_monthly  <- era %>% group_by(Month) %>% summarise(T_era  = mean(T, na.rm = TRUE))

bias <- left_join(obs_monthly, era_monthly, by = "Month") %>%
  mutate(T_bias = T_obs - T_era)

# Apply correction
era <- left_join(era, bias[, c("Month", "T_bias")], by = "Month")
era$T <- era$T + era$T_bias
```

Repeat for each variable (`T`, `MB`, and precipitation if available).
If no local observations exist, skip bias correction but note the uncertainty
this introduces, particularly for soil moisture.

---

## Checking your data before running CASSIA

Before passing any weather dataframe to `CASSIA_cpp()`, plot each column
to check for outliers, implausible values, and gaps:

```{r, fig.width=7, fig.height=9}
# Using the built-in data as an example
par(mfrow = c(3, 2))
plot(weather_original$dates, weather_original$T,
     type = "l", xlab = "", ylab = "Temperature (°C)", main = "Air temperature")
plot(weather_original$dates, weather_original$TSA,
     type = "l", xlab = "", ylab = "°C", main = "Soil temperature A")
plot(weather_original$dates, weather_original$TSB,
     type = "l", xlab = "", ylab = "°C", main = "Soil temperature B")
plot(weather_original$dates, weather_original$MB,
     type = "l", xlab = "", ylab = "m³ m⁻³", main = "Soil moisture B")
plot(weather_original$dates, weather_original$Rain,
     type = "l", xlab = "", ylab = "mm day⁻¹", main = "Precipitation")
plot(weather_original$dates, weather_original$P,
     type = "l", xlab = "", ylab = "g C m⁻² day⁻¹", main = "Photosynthesis (P)")
par(mfrow = c(1, 1))
```

Also check for missing values — CASSIA will error on any `NA`:

```{r}
colSums(is.na(weather_original[, c("T", "TSA", "TSB", "MB", "Rain", "P")]))
```

Expected units and approximate ranges for a boreal site:

| Column | Min | Max | Notes |
|--------|-----|-----|-------|
| `T` | −30 | 30 | °C; seasonal cycle clear |
| `TSA` | −5 | 25 | °C; smaller amplitude than air temperature |
| `TSB` | −2 | 20 | °C; more damped than TSA |
| `MB` | 0 | 0.6 | m³ m⁻³; never negative |
| `Rain` | 0 | 60 | mm day⁻¹; all values ≥ 0 |
| `P` | 0 | 15 | g C m⁻² day⁻¹; near-zero in winter |

Values persistently outside these ranges usually indicate a unit conversion
error or a sensor fault in the raw data.
