---
title: "The parameters in CASSIA"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parameters_pCROB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FLASE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CASSIA)
dim(parameters_p)
colnames(parameters_p)
```

| Parameter Name   | Meaning   | Reference |
|:----------------:|:---------:|:---------:|
|Q10.N            | Change in needle respiration at 10'C difference | Schiestl-Aalto 2015 |
|Rm.N             | Maintanaince respiration needles | Schiestl-Aalto 2015 |
|Q10.S            | Change in woody respiration at 10'C difference | Schiestl-Aalto 2015 |
|Rm.S             | Maintainance respiration ... | Schiestl-Aalto 2015 |
|Q10.R            | Change in root respiration at 10'C difference | Schiestl-Aalto 2015 |
|Rm.R             | Maintainance respiraton roots | Schiestl-Aalto 2015 |
|sR0              |The ontogenetic development stage of root biomass fraction initial position|Schiestl-Aalto 2015|
|sRc              |The ontogenetic development stage of root biomass fraction last position|Schiestl-Aalto 2015|
|growth.myco      |Percentage of photosynthates go to mycorrhiza|Schiestl-Aalto 2019|
|root.lifetime    |Lettosuo root lifetime|Lettosuo parameterisation|
|HH0              |Length growth initial value mm|Schiestl-Aalto 2015|
|sH0              | The ontogenetic development stage of height biomass fraction initial poisition |Schiestl-Aalto 2015|
|LH0              |Daily maximum height growth|Schiestl-Aalto 2015|
|sHc              |The ontogenetic development stage of height biomass fraction last poisition |Schiestl-Aalto 2015|
|sN0              |The ontogenetic development stage of needle biomass fraction first position |Schiestl-Aalto 2015|
|LN0              |Daily maximum needle growth|Schiestl-Aalto 2015|
|sNc              |The ontogenetic development stage of needle biomass fraction last position |Schiestl-Aalto 2015|
|HN0              |Length growth initial value mm|Schiestl-Aalto 2015|
|sD0.Trad         |The ontogenetic development stage of diameter biomass fraction initial position (Trad refers to the old method, sD.estim.T.count == F) |Schiestl-Aalto 2015|
|LD0              |Daily maximum secondary growth|Schiestl-Aalto 2015|
|sDc              |The ontogenetic development stage of diameter biomass fraction last position|Schiestl-Aalto 2015|
|tau.Ee           |Duration of tracheid enlargement earlywood|Schiestl-Aalto 2015|
|tau.El           |Duration of tracheid enlargement latewood|Schiestl-Aalto 2015|
|tau.We           |Duration of secondary cell wall enlargement early wood|Schiestl-Aalto 2015|
|tau.Wl           |Duration of secondary cell wall enlargement early wood|Schiestl-Aalto 2015|
|tau.GPP          |Dapening factor for GPP|Schiestl-Aalto 2015|
|Uggla            ||Xylogenesis|
|sB0              |The ontogenetic development stage of bud biomass fraction initial position|Schiestl-Aalto 2015|
|sBc              |The ontogenetic development stage of bud biomass fraction last position|Schiestl-Aalto 2015|
|cell.d.ew        |Early wood cell diameter|Schiestl-Aalto 2015|
|cell.d.lw        |Late wood cell diameter|Schiestl-Aalto 2015|
|cell.l.ew        |Early wood cell length|Schiestl-Aalto 2015|
|cell.l.lw        |Late  wood cell length|Schiestl-Aalto 2015|
|cell.wall.density.ew  ||Xylogenesis|
|cell.wall.density.lw  ||Xylogenesis|
|wall.thickness.ew     ||Xylogenesis|
|wall.thickness.lw     ||Xylogenesis|
|cell.volumn.growth.per.day.ew  ||Xylogensis|
|cell.volumn.growth.per.day.lw  ||Xylogensis|
|density_tree                   |Tree density|Schiestl-Aalto 2015|
|carbon_share                   |Mean carbon for entire tree|Schiestl-Aalto 2015|
|D0                             |Mean stem diameter at 1.3 m|Schiestl-Aalto 2015|
|h0                             |Dominant hieght|Schiestl-Aalto 2015|
|n_age                          |Mean needle age|Schiestl-Aalto 2015|
|n_lenght                       |Mean needle length|Schiestl-Aalto 2015|
|h_increment                    |Height increment|Schiestl-Aalto 2015|
|SLA                            |Surface Leaf Area|Not currently working|

```{r setup}
library(CASSIA)
dim(ratios_p)
colnames(ratios_p)
```

| Parameter Name   | Meaning   | Reference |
|:----------------:|:---------:|:---------:|
| form_factor      |  | |
| needle_fineroot_ratio | | |
| spawood.share | | |
| height_growth_coefficient 
| diameter_growth_coefficient 
| height_growth_coefficient 
| height_growth_coefficient 
| diameter_growth_coefficient 
| diameter_growth_coefficient 

```{r setup}
library(CASSIA)
dim(sperling_p)
colnames(sperling_p)
```

| Parameter Name   | Meaning   | Reference |
|:----------------:|:---------:|:---------:|
| starch0 | Initial starch value | Schiestl-Aalto 2019|
| sugar0 | Initial sugar value | Schiestl-Aalto 2019 |
| starch.needles0 | Initial starch value needles ||
| starch.phloem0 | Initial starch value phloem ||
| starch.xylem.sh0 | Initial starch value shoot xylem ||
| starch.xylem.st0 | Initial starch value stem xylem ||
| starch.roots0 | Initial starch value roots ||
| suagr.needles0 | Initial sugar value needles ||
| suagr.phloem0 | Initial sugar value phloem ||
| suagr.xylem.sh0 | Initial sugar value shoot xylem ||
| suagr.xylem.st0 | Initial sugar value stem xylem ||
| suagr.roots0 | Initial sugar value roots ||
| Wala | | Schiestl-Aalto 2019 | # TODO: understand this!
| carbon.sugar | The amount of carbon that is in sugar | Schiestl-Aalto 2019 |
| carbon.starch | The amount of carbon that is in starch | Schiestl-Aalto 2019 |
| tau.t | Dampenning parameters for starch transfer if values are too low and emergency starch is needed in sperling for every step in original | Schiestl-Aalto 2019|
| tau.s | Dampenning parameters for sugar transder for every step in the original model, not in sperling model| Schiestl-Aalto 2019 |
| starch00 | Initial starch value used when the model doesn't change starch initial condition on a yearly basis | Schiestl-Aalto 2019 |
| sugar00 | Initial sugar value used when the model doesn't change suagr initial condition on a yearly basis | Schiestl-Aalto 2019 |
| Q10s | Change in starch synthesis at 10'C difference | Sperling 2019 |
| Q10d | Change in starch degradation at 10'C difference | Sperling 2019 |
| SCb | The level of sugar the phloem needs to cross to trigger bud burst | Sperling 2019 |
| sugar.level | Simple sugar model equilibrium point | Schiestl-Aalto 2019 |
| Ad0 | Initial starch degradation frequency factor | Sperling, 2019 |
| lamda | Natural decay rate of proteins | Sperling, 2019 |
| delta | Additional expression | Sperling, 2019 |
| k_np | Concentration difference between sugar in needles and phloem |  |
| k_pr | Concentration difference between sugar in phloem and roots ||
| p_pxsh | Concentration difference between sugar in phloem and shoot xylem ||
| p_pxst | Concentration difference between sugar in phloem and stem xylem ||
| myco.thesh | The percentage of sugar that goes to mycorrhiza | |








