---
title: "Photosynthesis Sub-model Comparison: Original vs PRELES vs p-hydro (Boreal)"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 5)
```

## Overview

This notebook compares CASSIA output when driven by three different sources
of photosynthesis on the same boreal (ERA5) weather data:

| Sub-model | Toggle | Approach |
|-----------|--------|----------|
| **Photosynthesis (Original)** | `photosynthesis_as_input = TRUE` | GPP taken directly from the `P` column of the weather data, which was pre-computed by the SPP (Scots Pine Process) model. This is the standard baseline used throughout CASSIA development. |
| **PRELES** | `preles = TRUE` | Empirical light-use efficiency; driven by PAR, VPD, CO₂, fAPAR |
| **p-hydro** | `phydro = TRUE` | Hydraulic optimality; driven by PPFD, VPD, SWP, atmospheric pressure |

The key questions are:

1. How well do PRELES and p-hydro reproduce the SPP-generated baseline GPP?
2. Does any GPP difference between models propagate into growth and ring width?
3. Where do PRELES and p-hydro diverge from the original — early season,
   late season, drought events?

The comparison plots use synthetic output that reflects the expected character
of each model. Replace with real runs once the ERA5 input file is available.

## Setup

```{r setup-params}
devtools::load_all(quiet = TRUE)
library(ggplot2)
library(patchwork)

parameters_all <- initialize_parameters()

organ_cols <- c(
  diameter = "#8B4513",
  needle   = "#228B22",
  root     = "#D2691E",
  height   = "#4169E1"
)
model_cols <- c(
  "Original" = "#009E73",
  "PRELES"   = "#E69F00",
  "phydro"   = "#0072B2"
)

# Set TRUE only if the ERA5 data files are present on this machine.
# All weather-loading and model-run chunks are gated on this flag, so the
# notebook is safe to run interactively without the data files.
DATA_AVAILABLE <- file.exists("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")
```

---

## Weather data

PRELES and p-hydro each require different weather columns, so two separate
ERA5 files are used. The Original (SPP) run uses the built-in `weather_original`
package data (2015–2017). SMEAR eddy covariance observations provide a
validation reference.

### Original (SPP) weather

The built-in package dataset. No external file needed.

```{r weather-original, eval=DATA_AVAILABLE}
data("weather_original")
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))
```

### ERA5 — PRELES run

Contains `PAR`, `VPD`, `CO2`, `fAPAR`, and the standard core columns.

```{r weather-preles, eval=DATA_AVAILABLE}
weather_preles <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")
names(weather_preles)[names(weather_preles) == "date"] <- "dates"
weather_preles$dates <- as.Date(weather_preles$dates)
```

### ERA5 — p-hydro run

Contains `PPFD`, `PPFD_max`, `PA`, `SWP` in addition to the core columns.

```{r weather-phydro, eval=DATA_AVAILABLE}
weather_phydro <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_phydro.csv")
names(weather_phydro)[names(weather_phydro) == "date"] <- "dates"
weather_phydro$dates    <- as.Date(weather_phydro$dates)
weather_phydro$Nitrogen <- 1        # placeholder; fit if N deposition data available
weather_phydro$PA       <- 101325   # sea-level atmospheric pressure (Pa)
```

### SMEAR eddy covariance observations

Measured ecosystem GPP from the SMEAR II tower at Hyytiälä. This is used
as an observational reference when comparing model GPP. Note that it
represents ecosystem-level flux (all species), whereas the model output is
for a single pine tree scaled to ground area; the difference is primarily
due to stand composition and leaf area.

```{r smear-obs, eval=DATA_AVAILABLE}
smear <- data.table::fread("~/Documents/CASSIA/data/daily_dataframe.csv")
smear$Date <- as.Date(smear$Date)
```

### Weather comparison: ERA5 vs SMEAR station

ERA5 reanalysis is used for long-term runs but carries systematic biases.
The four panels below show how ERA5 compares to the observed SMEAR
meteorological record for the overlapping 2015–2018 period.

```{r weather-comparison, eval=DATA_AVAILABLE}
smear_meteo <- data.table::as.data.table(
  readxl::read_xlsx(
    "~/Documents/CASSIA_Calibration/Raw_Data/hyy_all_2008_2019.xlsx",
    sheet = 1, skip = 1
  )
)
names(smear_meteo) <- names(
  readxl::read_xlsx(
    "~/Documents/CASSIA_Calibration/Raw_Data/hyy_all_2008_2019.xlsx",
    sheet = 1
  )
)
smear_meteo$Date <- as.Date(smear_meteo$Date)

weather_preles_plot <- weather_preles[
  as.integer(format(weather_preles$date, "%Y")) %in% 2015:2018, ]

p_par  <- ggplot() +
  geom_line(aes(x = weather_preles_plot$date, y = weather_preles_plot$PAR,
                colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$PAR,
                 colour = "SMEAR"), size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "PAR (mol m⁻² day⁻¹)", colour = NULL) + theme_bw()

p_t    <- ggplot() +
  geom_line(aes(x = weather_preles_plot$date, y = weather_preles_plot$T,
                colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$TA_F,
                 colour = "SMEAR"), size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "Temperature (°C)", colour = NULL) + theme_bw()

p_rain <- ggplot() +
  geom_line(aes(x = weather_preles_plot$date, y = weather_preles_plot$Rain,
                colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$P_F,
                 colour = "SMEAR"), size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "Rain (mm day⁻¹)", colour = NULL) + theme_bw()

p_vpd  <- ggplot() +
  geom_line(aes(x = weather_preles_plot$date, y = weather_preles_plot$VPD,
                colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$VPD_F...8,
                 colour = "SMEAR"), size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "VPD (kPa)", colour = NULL) + theme_bw()

(p_par | p_t) / (p_rain | p_vpd) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "ERA5 vs SMEAR station weather (2015–2018)",
                  subtitle = "ERA5 is used for model runs; SMEAR is the observed reference")
```

---

## Model runs

### Photosynthesis (Original)

GPP comes directly from the `P` column of `weather_original`, which was
generated by the SPP model. This is the standard input used throughout
CASSIA development and acts as the baseline for this comparison.

```{r original-run, eval=DATA_AVAILABLE}
CASSIA_original <- CASSIA_cpp(
  weather                 = weather_original,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = TRUE,
  CASSIA_graphs           = FALSE
)
```

### PRELES

`fAPAR_Tian = TRUE` enables the dynamic fAPAR scaling derived from the
Tian et al. LAI product, which improves GPP seasonality relative to using
a fixed fAPAR constant.

```{r preles-run, eval=DATA_AVAILABLE}
CASSIA_preles <- CASSIA_cpp(
  weather                 = weather_preles,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  CASSIA_graphs           = FALSE
)
```

### p-hydro

The `phydro_param` vector contains calibrated hydraulic parameters. The
p-hydro run uses a separate weather file that provides the required
`PPFD`, `PPFD_max`, `PA`, and `SWP` columns. Note: `ecoevolutionary = TRUE`
is not currently compiled into the package (`CASSIA_eeo` export is disabled
in `src/ecoevolutionary.cpp`); use standard `phydro = TRUE` only.

```{r phydro-run, eval=DATA_AVAILABLE}
phydro_parameters_in <- c(
  0.1008, 0.180496537959982, 5, 0.026263945805926, 0.011, 50,
  0.5, -0.857817410110663, 4.1311874912949e17, 1, 2.45e-2, 2.0, 1.1, 0.1,
  15, 10, 5, 0, 1,
  exp(-0.5 * 1.8), exp(-0.5 * 3.5), exp(-0.5 * 5.5)
)

CASSIA_phydro <- CASSIA_cpp(
  weather                 = weather_phydro,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  phydro_param            = phydro_parameters_in,
  photosynthesis_as_input = FALSE,
  phydro                  = TRUE,
  CASSIA_graphs           = FALSE
)
```

---

## Synthetic example output

Synthetic boreal output for 2015–2017. The series reflect the expected
character of each source:

- **SMEAR (observed)**: ecosystem eddy covariance GPP from the Hyytiälä tower.
  Higher than single-tree model output because it integrates all species
  and stand-level canopy area; used here only as a shape reference.
- **Original (SPP)**: smooth, process-based seasonal curve calibrated to
  Hyytiälä; treated as the model baseline.
- **PRELES**: similar shape to Original but independently driven by radiation
  and LUE; slight offset due to different calibration.
- **p-hydro**: responds additionally to VPD and soil water potential, causing
  a 35% mid-summer depression in the dry year 2016.

Replace `CASSIA_original`, `CASSIA_preles`, `CASSIA_phydro`, and `smear` with
real model output and observations once the ERA5 files are available.

```{r synthetic-data}
set.seed(42)
n    <- 1096
year <- rep(2015:2017, times = c(365, 366, 365))
doy  <- c(1:365, 1:366, 1:365)

# ── SMEAR EC observations: ecosystem GPP, higher than single-tree model ───────
# Eddy covariance integrates all species and stand LAI; used as shape reference
gpp_smear <- pmax(0,
  12.5 * exp(-((doy - 195)^2) / (2 * 65^2)) + rnorm(n, 0, 0.60)
)
smear <- data.frame(
  Date     = as.Date(doy - 1, origin = paste0(year, "-01-01")),
  GPP_mean = gpp_smear
)

# ── Original (SPP): process-based reference ───────────────────────────────────
gpp_original <- pmax(0,
  10.2 * exp(-((doy - 192)^2) / (2 * 62^2)) + rnorm(n, 0, 0.20)
)

# ── PRELES: light-use efficiency, similar shape with its own calibration ──────
gpp_preles <- pmax(0,
  9.5 * exp(-((doy - 190)^2) / (2 * 60^2)) + rnorm(n, 0, 0.25)
)

# ── p-hydro: hydraulic optimality, drought-sensitive ─────────────────────────
drought_penalty <- ifelse(
  year == 2016 & doy >= 170 & doy <= 220,
  0.65,   # 35% suppression during mid-summer dry spell in 2016
  1.0
)
gpp_phydro <- pmax(0,
  8.8 * exp(-((doy - 188)^2) / (2 * 58^2)) * drought_penalty +
    rnorm(n, 0, 0.30)
)

# ── Growth: same allocation logic for all three, scaled by GPP ───────────────
make_output <- function(gpp_vec) {
  set.seed(1)
  diam   <- pmax(0, 0.012 * exp(-((doy - 175)^2) / (2 * 30^2)) + rnorm(n, 0, 5e-4))
  needle <- pmax(0, 0.018 * exp(-((doy - 165)^2) / (2 * 25^2)) + rnorm(n, 0, 6e-4))
  root   <- pmax(0, 0.010 * exp(-((doy - 185)^2) / (2 * 40^2)) + rnorm(n, 0, 4e-4))
  height <- pmax(0, 0.008 * exp(-((doy - 160)^2) / (2 * 20^2)) + rnorm(n, 0, 3e-4))

  scale  <- gpp_vec / mean(gpp_vec, na.rm = TRUE)
  diam   <- diam * scale; needle <- needle * scale
  root   <- root * scale; height <- height * scale

  rw <- numeric(n); running <- 0; prev_yr <- year[1]
  for (i in seq_len(n)) {
    if (year[i] != prev_yr) { running <- 0; prev_yr <- year[i] }
    running <- running + diam[i] * 0.12
    rw[i]  <- running
  }

  list(
    Preles = data.frame(
      GPP       = gpp_vec,
      ET        = pmax(0, 2.5 * gpp_vec / 9 + rnorm(n, 0, 0.1)),
      SoilWater = 0.25 + 0.05 * sin(2 * pi * doy / 365) + rnorm(n, 0, 0.01),
      fAPAR     = rep(0.7, n)
    ),
    Growth = data.frame(
      year                    = year, day = doy,
      g                       = pmax(0, 1 / (1 + exp(-0.185 * (doy - 285)))),
      bud_growth_potential    = diam * 0.3,
      wall_growth_potential   = diam * 1.2,
      needle_growth_potential = needle * 1.2,
      roots_growth_potential  = root   * 1.1,
      height_growth_potential = height * 1.1,
      bud_growth              = diam * 0.3,
      diameter_growth         = diam,  needle_growth = needle,
      root_growth             = root,  height_growth = height,
      respiration_growth      = pmax(0, 0.15 * gpp_vec + rnorm(n, 0, 0.03)),
      respiration_maintenance = pmax(0, 0.20 * gpp_vec + rnorm(n, 0, 0.02)),
      ring_width              = rw,    mm_potenital  = rw * 1.2,
      leaf_mass               = rep(3.8, n),
      leaf_area               = rep(7.5, n)
    )
  )
}

CASSIA_original <- make_output(gpp_original)
CASSIA_preles   <- make_output(gpp_preles)
CASSIA_phydro   <- make_output(gpp_phydro)
```

---

## Comparison

### 1. GPP time series

All three model GPP series overlaid with SMEAR eddy covariance observations
(grey points) as a shape reference. Original (SPP) is the model baseline;
p-hydro diverges during the 2016 mid-summer drought. The SMEAR values are
systematically higher because they represent ecosystem-level flux rather than
a single pine tree.

```{r gpp-timeseries, fig.height=4}
gpp_df <- data.frame(
  day      = seq_len(n),
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)

gpp_long <- reshape(gpp_df,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
gpp_long$model <- factor(gpp_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot() +
  geom_point(data = smear,
             aes(x = as.integer(Date - as.Date("2014-12-31")),
                 y = GPP_mean),
             colour = "grey70", size = 0.5, alpha = 0.6) +
  geom_line(data = gpp_long,
            aes(x = day, y = GPP, colour = model, linewidth = model),
            alpha = 0.85) +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  scale_x_continuous(
    breaks = c(1, 366, 731),
    labels = c("2015", "2016", "2017")
  ) +
  annotate("text", x = 50, y = max(smear$GPP_mean) * 0.9,
           label = "SMEAR EC (obs.)", colour = "grey50", size = 3, hjust = 0) +
  labs(title = "Daily GPP: Original (SPP) vs PRELES vs p-hydro (2015–2017)",
       subtitle = "Grey points = SMEAR eddy covariance (ecosystem GPP); p-hydro shows drought suppression in 2016",
       x = NULL, y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 2. GPP difference from Original

How much does each model deviate from the SPP baseline? Positive = model
predicts more GPP than Original; negative = less.

```{r gpp-difference, fig.height=5}
diff_df <- data.frame(
  day    = seq_len(n),
  PRELES = CASSIA_preles$Preles$GPP   - CASSIA_original$Preles$GPP,
  phydro = CASSIA_phydro$Preles$GPP   - CASSIA_original$Preles$GPP
)

diff_long <- reshape(diff_df,
  varying = c("PRELES", "phydro"), v.names = "diff",
  timevar = "model", times = c("PRELES", "phydro"), direction = "long"
)
diff_long$model <- factor(diff_long$model, levels = c("PRELES", "phydro"))

ggplot(diff_long, aes(x = day, y = diff, fill = diff > 0)) +
  geom_col(width = 1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#56B4E9", "FALSE" = "#D55E00")) +
  scale_x_continuous(
    breaks = c(1, 366, 731),
    labels = c("2015", "2016", "2017")
  ) +
  facet_wrap(~model, ncol = 1) +
  labs(title = "GPP deviation from Original (SPP baseline)",
       subtitle = "Blue = model higher than Original; orange = model lower than Original",
       x = NULL, y = "Δ GPP (g C m⁻² day⁻¹)") +
  theme_bw()
```

### 3. Seasonal GPP pattern

DOY-averaged GPP removes inter-annual variability and shows whether the
models agree on the shape and timing of the growing season. A mismatch in
peak timing or width would indicate a structural difference between models.

```{r seasonal-gpp, fig.height=4}
doy_gpp <- data.frame(
  doy      = doy,
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)
doy_agg <- aggregate(cbind(Original, PRELES, phydro) ~ doy,
                     data = doy_gpp, FUN = mean)

doy_long <- reshape(doy_agg,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
doy_long$model <- factor(doy_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot(doy_long, aes(x = doy, y = GPP, colour = model, linewidth = model)) +
  geom_line() +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  labs(title = "Mean GPP by day of year (averaged 2015–2017)",
       x = "Day of year", y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 4. Annual carbon budget

Annual totals of GPP, growth, and respiration for each model. If the carbon
allocation logic works correctly, differences in GPP between models should
scale through to proportional differences in growth.

```{r annual-budget, fig.height=5}
budget_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(
    model        = model_name,
    year         = g$year,
    GPP          = model_out$Preles$GPP,
    growth_total = g$diameter_growth + g$needle_growth +
                   g$root_growth + g$height_growth,
    resp_growth  = g$respiration_growth,
    resp_maint   = g$respiration_maintenance
  )
}

budget <- rbind(
  budget_model(CASSIA_original, "Original"),
  budget_model(CASSIA_preles,   "PRELES"),
  budget_model(CASSIA_phydro,   "phydro")
)

budget_annual <- aggregate(
  cbind(GPP, growth_total, resp_growth, resp_maint) ~ model + year,
  data = budget, FUN = sum, na.rm = TRUE
)
budget_annual$model <- factor(budget_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

budget_long <- reshape(budget_annual,
  varying   = c("GPP", "growth_total", "resp_growth", "resp_maint"),
  v.names   = "value",
  timevar   = "flux",
  times     = c("GPP", "Total growth", "Growth resp.", "Maint. resp."),
  direction = "long"
)

ggplot(budget_long,
       aes(x = factor(year), y = value, fill = flux)) +
  geom_col(position = "dodge") +
  facet_wrap(~model) +
  scale_fill_manual(values = c(
    "GPP"          = "#E69F00",
    "Total growth" = "#009E73",
    "Growth resp." = "#CC79A7",
    "Maint. resp." = "#D55E00"
  )) +
  labs(title = "Annual carbon budget by photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = NULL) +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 5. Growth components by year

Organ-level breakdown of annual growth carbon. If allocation is functioning
correctly across all three runs, the proportions (diameter : needle : root :
height) should be similar even when total growth differs.

```{r growth-components, fig.height=5}
growth_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(
    model    = model_name,
    year     = g$year,
    diameter = g$diameter_growth,
    needle   = g$needle_growth,
    root     = g$root_growth,
    height   = g$height_growth
  )
}

growth <- rbind(
  growth_model(CASSIA_original, "Original"),
  growth_model(CASSIA_preles,   "PRELES"),
  growth_model(CASSIA_phydro,   "phydro")
)

growth_annual <- aggregate(
  cbind(diameter, needle, root, height) ~ model + year,
  data = growth, FUN = sum, na.rm = TRUE
)
growth_annual$model <- factor(growth_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

growth_long <- reshape(growth_annual,
  varying   = c("diameter", "needle", "root", "height"),
  v.names   = "value",
  timevar   = "organ",
  times     = c("diameter", "needle", "root", "height"),
  direction = "long"
)

ggplot(growth_long,
       aes(x = factor(year), y = value, fill = organ)) +
  geom_col() +
  facet_wrap(~model) +
  scale_fill_manual(values = organ_cols) +
  labs(title = "Annual growth by organ and photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = "Organ") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 6. Cumulative ring width

Ring width is the most directly observable output that can be checked against
dendrometer or core data. Do the models agree on the final annual ring width,
or does the GPP difference translate into a measurable divergence?

```{r ring-width, fig.height=5}
rw_df <- rbind(
  data.frame(model = "Original", year = year, doy = doy,
             rw = CASSIA_original$Growth$ring_width),
  data.frame(model = "PRELES",   year = year, doy = doy,
             rw = CASSIA_preles$Growth$ring_width),
  data.frame(model = "phydro",   year = year, doy = doy,
             rw = CASSIA_phydro$Growth$ring_width)
)
rw_df$model <- factor(rw_df$model, levels = c("Original", "PRELES", "phydro"))

ggplot(rw_df, aes(x = doy, y = rw,
                  colour = factor(year), linetype = model,
                  group  = interaction(model, year))) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = c("2015" = "#1b7837",
                                 "2016" = "#762a83",
                                 "2017" = "#e08214")) +
  scale_linetype_manual(values = c(Original = "solid",
                                   PRELES   = "dashed",
                                   phydro   = "dotted")) +
  labs(title = "Cumulative ring width by year and photosynthesis source",
       subtitle = "Solid = Original (SPP), dashed = PRELES, dotted = p-hydro",
       x = "Day of year", y = "Ring width (mm)",
       colour = "Year", linetype = "Model") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 7. Summary table

```{r summary-table}
summary_model <- function(model_out, model_name) {
  g   <- model_out$Growth
  gpp <- model_out$Preles$GPP

  total_growth <- sum(g$diameter_growth + g$needle_growth +
                      g$root_growth + g$height_growth, na.rm = TRUE)

  annual_rw <- tapply(g$ring_width, g$year, max, na.rm = TRUE)

  data.frame(
    Model                        = model_name,
    `Mean GPP (g C m-2 day-1)`  = round(mean(gpp, na.rm = TRUE), 2),
    `Total growth (g C m-2)`    = round(total_growth, 1),
    `Growth / GPP (%)`          = round(total_growth / sum(gpp, na.rm = TRUE) * 100, 1),
    `Mean annual ring width (mm)` = round(mean(annual_rw), 3),
    check.names = FALSE
  )
}

knitr::kable(
  rbind(
    summary_model(CASSIA_original, "Original (SPP)"),
    summary_model(CASSIA_preles,   "PRELES"),
    summary_model(CASSIA_phydro,   "p-hydro")
  ),
  caption = "Summary statistics across 2015–2017"
)
```
