---
title: "Photosynthesis Sub-model Comparison: Original vs PRELES vs p-hydro (Boreal)"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 5)
```

## Overview

This notebook compares CASSIA output when driven by three different sources
of photosynthesis on the same boreal (ERA5) weather data:

| Sub-model | Toggle | Approach |
|-----------|--------|----------|
| **Photosynthesis (Original)** | `photosynthesis_as_input = TRUE` | GPP taken directly from the `P` column of the weather data, which was pre-computed by the SPP (Scots Pine Process) model. This is the standard baseline used throughout CASSIA development. |
| **PRELES** | `preles = TRUE` | Empirical light-use efficiency; driven by PAR, VPD, CO₂, fAPAR |
| **p-hydro** | `phydro = TRUE` | Hydraulic optimality; driven by PPFD, VPD, SWP, atmospheric pressure |

The key questions are:

1. How well do PRELES and p-hydro reproduce the SPP-generated baseline GPP?
2. Does any GPP difference between models propagate into growth and ring width?
3. Where do PRELES and p-hydro diverge from the original — early season,
   late season, drought events?

The comparison plots use synthetic output that reflects the expected character
of each model. Replace with real runs once the ERA5 input file is available.

## Setup

```{r setup-params}
library(CASSIA)
library(ggplot2)
library(patchwork)

parameters_all <- initialize_parameters()

organ_cols <- c(
  diameter = "#8B4513",
  needle   = "#228B22",
  root     = "#D2691E",
  height   = "#4169E1"
)
model_cols <- c(
  "Original" = "#009E73",
  "PRELES"   = "#E69F00",
  "phydro"   = "#0072B2"
)
```

---

## Weather data

ERA5 daily data for the boreal site. All three runs use the same weather
file. The `P` column contains daily GPP (g C m⁻² day⁻¹) pre-computed by
the SPP model; it is the photosynthesis source for the Original run.
PRELES and p-hydro compute their own GPP internally so `P` is not used
in those runs.

```{r weather, eval=FALSE}
weather_ERAS <- read.delim("~/Documents/ERAS_dataset.csv", sep = ",")

names(weather_ERAS) <- c("Year", "Month", "Decimal_year", "T", "VPD",
                         "PPFD", "PPFD_max", "SWP", "dates")

weather_ERAS$VPD   <- 0.1 * weather_ERAS$VPD   # hPa → kPa
weather_ERAS$dates <- as.Date(
  paste(weather_ERAS$Year, weather_ERAS$Month, "01", sep = "-")
)

# P: SPP-generated GPP (g C m-2 day-1) — used by the Original run only
# Load from the SPP output file matched to this ERA5 period
weather_ERAS$P     <- spp_gpp_timeseries   # replace with the actual object/read

# Columns for PRELES (PAR in mol m-2 day-1; convert from PPFD in umol m-2 s-1)
weather_ERAS$PAR   <- weather_ERAS$PPFD * 0.0864
weather_ERAS$fAPAR <- 0.7
weather_ERAS$CO2   <- 421.38

# Shared columns
weather_ERAS$Nitrogen <- 1
weather_ERAS$TSA      <- NA
weather_ERAS$TSB      <- NA
weather_ERAS$MB       <- NA
weather_ERAS$Rain     <- NA
weather_ERAS$PA       <- 101325
```

---

## Model runs

### Photosynthesis (Original)

GPP comes directly from the `P` column in the weather data, which was
generated by the SPP model. This is the standard input used throughout
CASSIA development and acts as the baseline for this comparison.

```{r original-run, eval=FALSE}
CASSIA_original <- CASSIA_cpp(
  weather                 = weather_ERAS,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = TRUE,
  CASSIA_graphs           = FALSE
)
```

### PRELES

```{r preles-run, eval=FALSE}
CASSIA_preles <- CASSIA_cpp(
  weather                 = weather_ERAS,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  CASSIA_graphs           = FALSE
)
```

### p-hydro

```{r phydro-run, eval=FALSE}
CASSIA_phydro <- CASSIA_cpp(
  weather                 = weather_ERAS,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = FALSE,
  phydro                  = TRUE,
  ecoevolutionary         = TRUE,
  CASSIA_graphs           = FALSE
)
```

---

## Synthetic example output

Synthetic boreal output for 2015–2017. The three GPP series reflect the
expected character of each source:

- **Original (SPP)**: smooth, process-based seasonal curve calibrated to
  Hyytiälä; treated as the reference.
- **PRELES**: similar shape to Original but independently driven by radiation
  and LUE; slight offset due to different calibration.
- **p-hydro**: responds additionally to VPD and soil water potential, causing
  a 35% mid-summer depression in the dry year 2016.

Replace `CASSIA_original`, `CASSIA_preles`, and `CASSIA_phydro` with real
model output once the ERA5 file is available.

```{r synthetic-data}
set.seed(42)
n    <- 1096
year <- rep(2015:2017, times = c(365, 366, 365))
doy  <- c(1:365, 1:366, 1:365)

# ── Original (SPP): process-based reference ───────────────────────────────────
gpp_original <- pmax(0,
  10.2 * exp(-((doy - 192)^2) / (2 * 62^2)) + rnorm(n, 0, 0.20)
)

# ── PRELES: light-use efficiency, similar shape with its own calibration ──────
gpp_preles <- pmax(0,
  9.5 * exp(-((doy - 190)^2) / (2 * 60^2)) + rnorm(n, 0, 0.25)
)

# ── p-hydro: hydraulic optimality, drought-sensitive ─────────────────────────
drought_penalty <- ifelse(
  year == 2016 & doy >= 170 & doy <= 220,
  0.65,   # 35% suppression during mid-summer dry spell in 2016
  1.0
)
gpp_phydro <- pmax(0,
  8.8 * exp(-((doy - 188)^2) / (2 * 58^2)) * drought_penalty +
    rnorm(n, 0, 0.30)
)

# ── Growth: same allocation logic for all three, scaled by GPP ───────────────
make_output <- function(gpp_vec) {
  set.seed(1)
  diam   <- pmax(0, 0.012 * exp(-((doy - 175)^2) / (2 * 30^2)) + rnorm(n, 0, 5e-4))
  needle <- pmax(0, 0.018 * exp(-((doy - 165)^2) / (2 * 25^2)) + rnorm(n, 0, 6e-4))
  root   <- pmax(0, 0.010 * exp(-((doy - 185)^2) / (2 * 40^2)) + rnorm(n, 0, 4e-4))
  height <- pmax(0, 0.008 * exp(-((doy - 160)^2) / (2 * 20^2)) + rnorm(n, 0, 3e-4))

  scale  <- gpp_vec / mean(gpp_vec, na.rm = TRUE)
  diam   <- diam * scale; needle <- needle * scale
  root   <- root * scale; height <- height * scale

  rw <- numeric(n); running <- 0; prev_yr <- year[1]
  for (i in seq_len(n)) {
    if (year[i] != prev_yr) { running <- 0; prev_yr <- year[i] }
    running <- running + diam[i] * 0.12
    rw[i]  <- running
  }

  list(
    Preles = data.frame(
      GPP       = gpp_vec,
      ET        = pmax(0, 2.5 * gpp_vec / 9 + rnorm(n, 0, 0.1)),
      SoilWater = 0.25 + 0.05 * sin(2 * pi * doy / 365) + rnorm(n, 0, 0.01),
      fAPAR     = rep(0.7, n)
    ),
    Growth = data.frame(
      year                    = year, day = doy,
      g                       = pmax(0, 1 / (1 + exp(-0.185 * (doy - 285)))),
      bud_growth_potential    = diam * 0.3,
      wall_growth_potential   = diam * 1.2,
      needle_growth_potential = needle * 1.2,
      roots_growth_potential  = root   * 1.1,
      height_growth_potential = height * 1.1,
      bud_growth              = diam * 0.3,
      diameter_growth         = diam,  needle_growth = needle,
      root_growth             = root,  height_growth = height,
      respiration_growth      = pmax(0, 0.15 * gpp_vec + rnorm(n, 0, 0.03)),
      respiration_maintenance = pmax(0, 0.20 * gpp_vec + rnorm(n, 0, 0.02)),
      ring_width              = rw,    mm_potenital  = rw * 1.2,
      leaf_mass               = rep(3.8, n),
      leaf_area               = rep(7.5, n)
    )
  )
}

CASSIA_original <- make_output(gpp_original)
CASSIA_preles   <- make_output(gpp_preles)
CASSIA_phydro   <- make_output(gpp_phydro)
```

---

## Comparison

### 1. GPP time series

All three GPP sources overlaid. The Original (SPP) series is the reference.
PRELES and p-hydro should track it closely during normal conditions; p-hydro
should diverge during the 2016 mid-summer drought.

```{r gpp-timeseries, fig.height=4}
gpp_df <- data.frame(
  day      = seq_len(n),
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)

gpp_long <- reshape(gpp_df,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
gpp_long$model <- factor(gpp_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot(gpp_long, aes(x = day, y = GPP, colour = model, linewidth = model)) +
  geom_line(alpha = 0.85) +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  scale_x_continuous(
    breaks = c(1, 366, 731),
    labels = c("2015", "2016", "2017")
  ) +
  labs(title = "Daily GPP: Original (SPP) vs PRELES vs p-hydro (2015–2017)",
       subtitle = "Original is the SPP-generated reference; p-hydro shows drought suppression in 2016",
       x = NULL, y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 2. GPP difference from Original

How much does each model deviate from the SPP baseline? Positive = model
predicts more GPP than Original; negative = less.

```{r gpp-difference, fig.height=5}
diff_df <- data.frame(
  day    = seq_len(n),
  PRELES = CASSIA_preles$Preles$GPP   - CASSIA_original$Preles$GPP,
  phydro = CASSIA_phydro$Preles$GPP   - CASSIA_original$Preles$GPP
)

diff_long <- reshape(diff_df,
  varying = c("PRELES", "phydro"), v.names = "diff",
  timevar = "model", times = c("PRELES", "phydro"), direction = "long"
)
diff_long$model <- factor(diff_long$model, levels = c("PRELES", "phydro"))

ggplot(diff_long, aes(x = day, y = diff, fill = diff > 0)) +
  geom_col(width = 1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#56B4E9", "FALSE" = "#D55E00")) +
  scale_x_continuous(
    breaks = c(1, 366, 731),
    labels = c("2015", "2016", "2017")
  ) +
  facet_wrap(~model, ncol = 1) +
  labs(title = "GPP deviation from Original (SPP baseline)",
       subtitle = "Blue = model higher than Original; orange = model lower than Original",
       x = NULL, y = "Δ GPP (g C m⁻² day⁻¹)") +
  theme_bw()
```

### 3. Seasonal GPP pattern

DOY-averaged GPP removes inter-annual variability and shows whether the
models agree on the shape and timing of the growing season. A mismatch in
peak timing or width would indicate a structural difference between models.

```{r seasonal-gpp, fig.height=4}
doy_gpp <- data.frame(
  doy      = doy,
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)
doy_agg <- aggregate(cbind(Original, PRELES, phydro) ~ doy,
                     data = doy_gpp, FUN = mean)

doy_long <- reshape(doy_agg,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
doy_long$model <- factor(doy_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot(doy_long, aes(x = doy, y = GPP, colour = model, linewidth = model)) +
  geom_line() +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  labs(title = "Mean GPP by day of year (averaged 2015–2017)",
       x = "Day of year", y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 4. Annual carbon budget

Annual totals of GPP, growth, and respiration for each model. If the carbon
allocation logic works correctly, differences in GPP between models should
scale through to proportional differences in growth.

```{r annual-budget, fig.height=5}
budget_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(
    model        = model_name,
    year         = g$year,
    GPP          = model_out$Preles$GPP,
    growth_total = g$diameter_growth + g$needle_growth +
                   g$root_growth + g$height_growth,
    resp_growth  = g$respiration_growth,
    resp_maint   = g$respiration_maintenance
  )
}

budget <- rbind(
  budget_model(CASSIA_original, "Original"),
  budget_model(CASSIA_preles,   "PRELES"),
  budget_model(CASSIA_phydro,   "phydro")
)

budget_annual <- aggregate(
  cbind(GPP, growth_total, resp_growth, resp_maint) ~ model + year,
  data = budget, FUN = sum, na.rm = TRUE
)
budget_annual$model <- factor(budget_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

budget_long <- reshape(budget_annual,
  varying   = c("GPP", "growth_total", "resp_growth", "resp_maint"),
  v.names   = "value",
  timevar   = "flux",
  times     = c("GPP", "Total growth", "Growth resp.", "Maint. resp."),
  direction = "long"
)

ggplot(budget_long,
       aes(x = factor(year), y = value, fill = flux)) +
  geom_col(position = "dodge") +
  facet_wrap(~model) +
  scale_fill_manual(values = c(
    "GPP"          = "#E69F00",
    "Total growth" = "#009E73",
    "Growth resp." = "#CC79A7",
    "Maint. resp." = "#D55E00"
  )) +
  labs(title = "Annual carbon budget by photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = NULL) +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 5. Growth components by year

Organ-level breakdown of annual growth carbon. If allocation is functioning
correctly across all three runs, the proportions (diameter : needle : root :
height) should be similar even when total growth differs.

```{r growth-components, fig.height=5}
growth_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(
    model    = model_name,
    year     = g$year,
    diameter = g$diameter_growth,
    needle   = g$needle_growth,
    root     = g$root_growth,
    height   = g$height_growth
  )
}

growth <- rbind(
  growth_model(CASSIA_original, "Original"),
  growth_model(CASSIA_preles,   "PRELES"),
  growth_model(CASSIA_phydro,   "phydro")
)

growth_annual <- aggregate(
  cbind(diameter, needle, root, height) ~ model + year,
  data = growth, FUN = sum, na.rm = TRUE
)
growth_annual$model <- factor(growth_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

growth_long <- reshape(growth_annual,
  varying   = c("diameter", "needle", "root", "height"),
  v.names   = "value",
  timevar   = "organ",
  times     = c("diameter", "needle", "root", "height"),
  direction = "long"
)

ggplot(growth_long,
       aes(x = factor(year), y = value, fill = organ)) +
  geom_col() +
  facet_wrap(~model) +
  scale_fill_manual(values = organ_cols) +
  labs(title = "Annual growth by organ and photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = "Organ") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 6. Cumulative ring width

Ring width is the most directly observable output that can be checked against
dendrometer or core data. Do the models agree on the final annual ring width,
or does the GPP difference translate into a measurable divergence?

```{r ring-width, fig.height=5}
rw_df <- rbind(
  data.frame(model = "Original", year = year, doy = doy,
             rw = CASSIA_original$Growth$ring_width),
  data.frame(model = "PRELES",   year = year, doy = doy,
             rw = CASSIA_preles$Growth$ring_width),
  data.frame(model = "phydro",   year = year, doy = doy,
             rw = CASSIA_phydro$Growth$ring_width)
)
rw_df$model <- factor(rw_df$model, levels = c("Original", "PRELES", "phydro"))

ggplot(rw_df, aes(x = doy, y = rw,
                  colour = factor(year), linetype = model,
                  group  = interaction(model, year))) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = c("2015" = "#1b7837",
                                 "2016" = "#762a83",
                                 "2017" = "#e08214")) +
  scale_linetype_manual(values = c(Original = "solid",
                                   PRELES   = "dashed",
                                   phydro   = "dotted")) +
  labs(title = "Cumulative ring width by year and photosynthesis source",
       subtitle = "Solid = Original (SPP), dashed = PRELES, dotted = p-hydro",
       x = "Day of year", y = "Ring width (mm)",
       colour = "Year", linetype = "Model") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### 7. Summary table

```{r summary-table}
summary_model <- function(model_out, model_name) {
  g   <- model_out$Growth
  gpp <- model_out$Preles$GPP

  total_growth <- sum(g$diameter_growth + g$needle_growth +
                      g$root_growth + g$height_growth, na.rm = TRUE)

  annual_rw <- tapply(g$ring_width, g$year, max, na.rm = TRUE)

  data.frame(
    Model                        = model_name,
    `Mean GPP (g C m-2 day-1)`  = round(mean(gpp, na.rm = TRUE), 2),
    `Total growth (g C m-2)`    = round(total_growth, 1),
    `Growth / GPP (%)`          = round(total_growth / sum(gpp, na.rm = TRUE) * 100, 1),
    `Mean annual ring width (mm)` = round(mean(annual_rw), 3),
    check.names = FALSE
  )
}

knitr::kable(
  rbind(
    summary_model(CASSIA_original, "Original (SPP)"),
    summary_model(CASSIA_preles,   "PRELES"),
    summary_model(CASSIA_phydro,   "p-hydro")
  ),
  caption = "Summary statistics across 2015–2017"
)
```
