---
title: "Photosynthesis sub-models: SPP, PRELES, and p-hydro"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 5)
```

## Overview

CASSIA can derive gross primary production (GPP) from three photosynthesis
sub-models. Each makes different trade-offs between data requirements,
mechanistic complexity, and ability to respond to climate variability:

| Sub-model | Toggle | Driver | Key feature |
|-----------|--------|--------|-------------|
| **SPP (Original)** | `photosynthesis_as_input = TRUE` | Pre-computed GPP in weather `P` column | Calibrated to Hyytiälä; the standard development baseline |
| **PRELES** | `preles = TRUE` | PAR, VPD, CO₂, fAPAR | Empirical light-use efficiency; `fAPAR_Tian = TRUE` improves seasonal shape |
| **p-hydro** | `phydro = TRUE` | PPFD, VPD, SWP, atmospheric pressure | Hydraulic optimality; drought-sensitive |

The key questions:

1. How well do PRELES and p-hydro reproduce the SPP baseline GPP?
2. Does any GPP difference propagate into growth and ring width?
3. Where do PRELES and p-hydro diverge from the original — spring onset,
   late season, drought events?

---

## Setup

```{r setup-params}
devtools::load_all(quiet = TRUE)
library(ggplot2)
library(patchwork)

parameters_all <- initialize_parameters()

organ_cols <- c(
  diameter = "#8B4513",
  needle   = "#228B22",
  root     = "#D2691E",
  height   = "#4169E1"
)
model_cols <- c(
  "Original" = "#009E73",
  "PRELES"   = "#E69F00",
  "phydro"   = "#0072B2"
)

# All weather-loading and model-run chunks are gated on this flag.
# Set TRUE only when the ERA5 data files are present on this machine.
DATA_AVAILABLE <- file.exists("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")
```

---

## Weather data

SPP uses the built-in `weather_original` (2015–2017). PRELES and p-hydro each
require different ERA5 columns, so separate files are loaded. SMEAR eddy
covariance GPP is used as an observational reference in the comparison plots.

### Original (SPP) — built-in package data

```{r weather-original, eval=DATA_AVAILABLE}
data("weather_original")
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))
```

### ERA5 — PRELES run

Columns: `date`, `T`, `VPD`, `PAR`, `Rain`, `CO2`, `fAPAR`, `TSA`, `TSB`, `MB`.

```{r weather-preles, eval=DATA_AVAILABLE}
weather_preles <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")
names(weather_preles)[names(weather_preles) == "date"] <- "dates"
weather_preles$dates <- as.Date(weather_preles$dates)
# Restrict to the same 2015-2017 window as weather_original
weather_preles <- weather_preles[format(weather_preles$dates, "%Y") %in% 2015:2017, ]
```

### ERA5 — p-hydro run

Additional columns beyond PRELES: `PPFD`, `PPFD_max`, `SWP`. `Nitrogen` and
`PA` are added here as placeholders; fit from deposition data if available.

```{r weather-phydro, eval=DATA_AVAILABLE}
weather_phydro <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_phydro.csv")
names(weather_phydro)[names(weather_phydro) == "date"] <- "dates"
weather_phydro$dates    <- as.Date(weather_phydro$dates)
weather_phydro$Nitrogen <- 1        # placeholder
weather_phydro$PA       <- 101325   # sea-level atmospheric pressure (Pa)
# Restrict to the same 2015-2017 window as weather_original
weather_phydro <- weather_phydro[format(weather_phydro$dates, "%Y") %in% 2015:2017, ]
```

### SMEAR eddy covariance observations

Ecosystem GPP from the SMEAR II tower. Values are higher than the single-tree
model output because eddy covariance integrates all species and stand-level
leaf area; used here as a shape and magnitude reference.

```{r smear-obs, eval=DATA_AVAILABLE}
smear <- data.table::fread("~/Documents/CASSIA/data/daily_dataframe.csv")
smear$Date <- as.Date(smear$Date)
```

### Weather comparison: ERA5 vs SMEAR station

ERA5 reanalysis is used for long-term PRELES and p-hydro runs but can carry
systematic biases. The four panels below show how ERA5 compares to the observed
SMEAR meteorological record for 2015–2018.

```{r weather-comparison, eval=DATA_AVAILABLE}
smear_meteo <- data.table::as.data.table(
  readxl::read_xlsx(
    "~/Documents/CASSIA_Calibration/Raw_Data/hyy_all_2008_2019.xlsx",
    sheet = 1, skip = 1
  )
)
names(smear_meteo) <- names(
  readxl::read_xlsx(
    "~/Documents/CASSIA_Calibration/Raw_Data/hyy_all_2008_2019.xlsx",
    sheet = 1
  )
)
smear_meteo$Date <- as.Date(smear_meteo$Date)

era5_2015_2018 <- weather_preles[
  as.integer(format(weather_preles$dates, "%Y")) %in% 2015:2018, ]

p_par  <- ggplot() +
  geom_line(aes(x = era5_2015_2018$dates, y = era5_2015_2018$PAR, colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$PAR, colour = "SMEAR"),
             size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "PAR", colour = NULL) + theme_bw()

p_t    <- ggplot() +
  geom_line(aes(x = era5_2015_2018$dates, y = era5_2015_2018$T, colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$TA_F, colour = "SMEAR"),
             size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "Temperature (°C)", colour = NULL) + theme_bw()

p_rain <- ggplot() +
  geom_line(aes(x = era5_2015_2018$dates, y = era5_2015_2018$Rain, colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$P_F, colour = "SMEAR"),
             size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "Rain (mm day⁻¹)", colour = NULL) + theme_bw()

p_vpd  <- ggplot() +
  geom_line(aes(x = era5_2015_2018$dates, y = era5_2015_2018$VPD, colour = "ERA5")) +
  geom_point(aes(x = smear_meteo$Date, y = smear_meteo$`VPD_F...8`, colour = "SMEAR"),
             size = 0.5) +
  scale_colour_manual(values = c(ERA5 = "#0072B2", SMEAR = "#009E73")) +
  labs(x = NULL, y = "VPD (kPa)", colour = NULL) + theme_bw()

(p_par | p_t) / (p_rain | p_vpd) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title    = "ERA5 vs SMEAR station weather (2015–2018)",
    subtitle = "ERA5 is used for model runs; SMEAR is the observed reference"
  )
```

---

## Model runs

### Original (SPP)

GPP is taken directly from the `P` column of `weather_original`, which was
pre-computed by the SPP (Scots Pine Process) model. This is the calibrated
baseline used throughout CASSIA development.

```{r original-run, eval=DATA_AVAILABLE}
CASSIA_original <- CASSIA_cpp(
  weather                 = weather_original,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = TRUE,
  CASSIA_graphs           = FALSE
)
```

### PRELES

`fAPAR_Tian = TRUE` applies a dynamic fAPAR derived from the Tian et al. LAI
product. This improves the seasonality of GPP relative to a fixed fAPAR value
by capturing the spring increase and autumn senescence of the canopy.

```{r preles-run, eval=DATA_AVAILABLE}
CASSIA_preles <- CASSIA_cpp(
  weather                 = weather_preles,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  CASSIA_graphs           = FALSE
)
```

### p-hydro

The calibrated `phydro_parameters_in` vector specifies the hydraulic parameters.
A separate ERA5 file is used because p-hydro requires `PPFD`, `PPFD_max`, `SWP`,
and `PA` columns that the PRELES file does not contain.
Note: `ecoevolutionary = TRUE` is not currently compiled into the package
(`CASSIA_eeo` export is disabled in `src/ecoevolutionary.cpp`).

> **Note:** The p-hydro photosynthesis call (`calc_plant_assimilation_rate` in
> `src/PlantFateLogic.cpp`) is not yet wired into the main `CASSIA_yearly`
> loop in `src/CASSIA.cpp`. The photosynthesis dispatch currently only handles
> SPP and PRELES; when `phydro = TRUE` is set, GPP returns zero. The underlying
> phydro library and canopy integration code compile correctly — this is purely
> a wiring issue. If you need p-hydro output, get in touch before running this
> section.

```{r phydro-run, eval=FALSE}
phydro_parameters_in <- c(
  0.1008, 0.180496537959982, 5, 0.026263945805926, 0.011, 50,
  0.5, -0.857817410110663, 4.1311874912949e17, 1, 2.45e-2, 2.0, 1.1, 0.1,
  15, 10, 5, 0, 1,
  exp(-0.5 * 1.8), exp(-0.5 * 3.5), exp(-0.5 * 5.5)
)

CASSIA_phydro <- CASSIA_cpp(
  weather                 = weather_phydro,
  site                    = "Hyde",
  pPREL                   = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters              = parameters_all$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_all$sperling_test,
  needle_mass_in          = parameters_all$needle_mass_in,
  Throughfall             = parameters_all$Throughfall,
  phydro_param            = phydro_parameters_in,
  photosynthesis_as_input = FALSE,
  phydro                  = TRUE,
  CASSIA_graphs           = FALSE
)
```

---

## Placeholder output (runs not yet available)

The comparison plots below use synthetic data that reflects the expected
character of each model. Replace the synthetic chunk with real output once
the ERA5 files are present and the model runs above have completed.

```{r synthetic-data, eval=!DATA_AVAILABLE}
set.seed(42)
n    <- 1096
year <- rep(2015:2017, times = c(365, 366, 365))
doy  <- c(1:365, 1:366, 1:365)

gpp_smear <- pmax(0, 12.5 * exp(-((doy - 195)^2) / (2 * 65^2)) + rnorm(n, 0, 0.6))
smear <- data.frame(
  Date     = as.Date(doy - 1, origin = paste0(year, "-01-01")),
  GPP_mean = gpp_smear
)

make_output <- function(gpp_vec) {
  set.seed(1)
  diam   <- pmax(0, 0.012 * exp(-((doy - 175)^2) / (2 * 30^2)) + rnorm(n, 0, 5e-4))
  needle <- pmax(0, 0.018 * exp(-((doy - 165)^2) / (2 * 25^2)) + rnorm(n, 0, 6e-4))
  root   <- pmax(0, 0.010 * exp(-((doy - 185)^2) / (2 * 40^2)) + rnorm(n, 0, 4e-4))
  height <- pmax(0, 0.008 * exp(-((doy - 160)^2) / (2 * 20^2)) + rnorm(n, 0, 3e-4))
  scale  <- gpp_vec / mean(gpp_vec, na.rm = TRUE)
  diam <- diam * scale; needle <- needle * scale
  root <- root * scale; height <- height * scale
  rw <- numeric(n); running <- 0; prev_yr <- year[1]
  for (i in seq_len(n)) {
    if (year[i] != prev_yr) { running <- 0; prev_yr <- year[i] }
    running <- running + diam[i] * 0.12; rw[i] <- running
  }
  list(
    Preles = data.frame(GPP = gpp_vec),
    Growth = data.frame(
      year = year, day = doy,
      diameter_growth         = diam,   needle_growth = needle,
      root_growth             = root,   height_growth = height,
      respiration_growth      = pmax(0, 0.15 * gpp_vec + rnorm(n, 0, 0.03)),
      respiration_maintenance = pmax(0, 0.20 * gpp_vec + rnorm(n, 0, 0.02)),
      ring_width              = rw
    )
  )
}

CASSIA_original <- make_output(
  pmax(0, 10.2 * exp(-((doy - 192)^2) / (2 * 62^2)) + rnorm(n, 0, 0.2))
)
CASSIA_preles <- make_output(
  pmax(0,  9.5 * exp(-((doy - 190)^2) / (2 * 60^2)) + rnorm(n, 0, 0.25))
)
drought_penalty <- ifelse(year == 2016 & doy >= 170 & doy <= 220, 0.65, 1.0)
CASSIA_phydro <- make_output(
  pmax(0, 8.8 * exp(-((doy - 188)^2) / (2 * 58^2)) * drought_penalty + rnorm(n, 0, 0.3))
)
```

---

## GPP comparison

### 1. Time series

All three model GPP series overlaid with SMEAR eddy covariance observations
(grey points). SMEAR values are systematically higher because they represent
ecosystem-level flux; they are used here as a shape and seasonal-timing
reference, not a calibration target.

```{r gpp-timeseries, fig.height=4}
year  <- CASSIA_original$Growth$year
doy   <- CASSIA_original$Growth$day
n     <- length(year)
dates <- as.Date(doy - 1, origin = paste0(year, "-01-01"))

gpp_df <- data.frame(
  date     = dates,
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)
gpp_long <- reshape(gpp_df,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
gpp_long$model <- factor(gpp_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot() +
  geom_point(data = smear,
             aes(x = Date, y = GPP_mean),
             colour = "grey70", size = 0.5, alpha = 0.6) +
  geom_line(data = gpp_long,
            aes(x = date, y = GPP, colour = model, linewidth = model),
            alpha = 0.85) +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  annotate("text", x = min(dates) + 30, y = max(smear$GPP_mean, na.rm = TRUE) * 0.9,
           label = "SMEAR EC (obs.)", colour = "grey50", size = 3, hjust = 0) +
  labs(title = "Daily GPP: Original (SPP) vs PRELES vs p-hydro (2015–2017)",
       subtitle = "Grey = SMEAR eddy covariance (ecosystem); model lines = single-tree",
       x = NULL, y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() + theme(legend.position = "bottom")
```

### 2. Deviation from Original (SPP baseline)

Positive = model predicts more GPP than the SPP baseline; negative = less.
Any sustained bias in one season or year would indicate a structural difference
between models.

```{r gpp-difference, fig.height=4}
diff_df <- data.frame(
  date   = dates,
  PRELES = CASSIA_preles$Preles$GPP - CASSIA_original$Preles$GPP,
  phydro = CASSIA_phydro$Preles$GPP  - CASSIA_original$Preles$GPP
)
diff_long <- reshape(diff_df,
  varying = c("PRELES", "phydro"), v.names = "diff",
  timevar = "model", times = c("PRELES", "phydro"), direction = "long"
)
diff_long$model <- factor(diff_long$model, levels = c("PRELES", "phydro"))

ggplot(diff_long, aes(x = date, y = diff, fill = diff > 0)) +
  geom_col(width = 1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#56B4E9", "FALSE" = "#D55E00")) +
  facet_wrap(~model, ncol = 1) +
  labs(title = "GPP deviation from Original (SPP baseline)",
       subtitle = "Blue = model higher than baseline; orange = model lower",
       x = NULL, y = "Δ GPP (g C m⁻² day⁻¹)") +
  theme_bw()
```

### 3. Seasonal pattern (DOY average)

DOY-averaged GPP removes inter-annual variability and shows whether the models
agree on the shape and timing of the growing season.

```{r seasonal-gpp, fig.height=4}
doy_gpp <- data.frame(
  doy      = doy,
  Original = CASSIA_original$Preles$GPP,
  PRELES   = CASSIA_preles$Preles$GPP,
  phydro   = CASSIA_phydro$Preles$GPP
)
doy_agg  <- aggregate(cbind(Original, PRELES, phydro) ~ doy,
                      data = doy_gpp, FUN = mean)
doy_long <- reshape(doy_agg,
  varying = c("Original", "PRELES", "phydro"), v.names = "GPP",
  timevar = "model", times = c("Original", "PRELES", "phydro"),
  direction = "long"
)
doy_long$model <- factor(doy_long$model, levels = c("Original", "PRELES", "phydro"))

ggplot(doy_long, aes(x = doy, y = GPP, colour = model, linewidth = model)) +
  geom_line() +
  scale_colour_manual(values = model_cols) +
  scale_linewidth_manual(values = c(Original = 1.1, PRELES = 0.7, phydro = 0.7)) +
  labs(title = "Mean GPP by day of year (averaged 2015–2017)",
       x = "Day of year", y = "GPP (g C m⁻² day⁻¹)", colour = "Model") +
  guides(linewidth = "none") +
  theme_bw() + theme(legend.position = "bottom")
```

---

## Growth comparison

### 4. Annual carbon budget

If carbon allocation is functioning correctly, differences in GPP between
models should scale through to proportional differences in growth.

```{r annual-budget, fig.height=5}
budget_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(
    model        = model_name,
    year         = g$year,
    GPP          = model_out$Preles$GPP,
    growth_total = g$diameter_growth + g$needle_growth +
                   g$root_growth     + g$height_growth,
    resp_growth  = g$respiration_growth,
    resp_maint   = g$respiration_maintenance
  )
}

budget <- rbind(
  budget_model(CASSIA_original, "Original"),
  budget_model(CASSIA_preles,   "PRELES"),
  budget_model(CASSIA_phydro,   "phydro")
)
budget_annual <- aggregate(
  cbind(GPP, growth_total, resp_growth, resp_maint) ~ model + year,
  data = budget, FUN = sum, na.rm = TRUE
)
budget_annual$model <- factor(budget_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

budget_long <- reshape(budget_annual,
  varying = c("GPP", "growth_total", "resp_growth", "resp_maint"),
  v.names = "value", timevar = "flux",
  times   = c("GPP", "Total growth", "Growth resp.", "Maint. resp."),
  direction = "long"
)

ggplot(budget_long, aes(x = factor(year), y = value, fill = flux)) +
  geom_col(position = "dodge") +
  facet_wrap(~model) +
  scale_fill_manual(values = c(
    "GPP"          = "#E69F00",
    "Total growth" = "#009E73",
    "Growth resp." = "#CC79A7",
    "Maint. resp." = "#D55E00"
  )) +
  labs(title = "Annual carbon budget by photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = NULL) +
  theme_bw() + theme(legend.position = "bottom")
```

### 5. Growth by organ

Organ-level breakdown of annual growth carbon. If allocation is working
correctly across all three runs, the proportions (diameter : needle : root :
height) should be similar even if total growth differs.

```{r growth-components, fig.height=5}
growth_model <- function(model_out, model_name) {
  g <- model_out$Growth
  data.frame(model = model_name, year = g$year,
             diameter = g$diameter_growth, needle = g$needle_growth,
             root     = g$root_growth,     height = g$height_growth)
}

growth <- rbind(
  growth_model(CASSIA_original, "Original"),
  growth_model(CASSIA_preles,   "PRELES"),
  growth_model(CASSIA_phydro,   "phydro")
)
growth_annual <- aggregate(
  cbind(diameter, needle, root, height) ~ model + year,
  data = growth, FUN = sum, na.rm = TRUE
)
growth_annual$model <- factor(growth_annual$model,
                               levels = c("Original", "PRELES", "phydro"))

growth_long <- reshape(growth_annual,
  varying = c("diameter", "needle", "root", "height"),
  v.names = "value", timevar = "organ",
  times   = c("diameter", "needle", "root", "height"),
  direction = "long"
)

ggplot(growth_long, aes(x = factor(year), y = value, fill = organ)) +
  geom_col() +
  facet_wrap(~model) +
  scale_fill_manual(values = organ_cols) +
  labs(title = "Annual growth by organ and photosynthesis source",
       x = "Year", y = "g C m⁻² year⁻¹", fill = "Organ") +
  theme_bw() + theme(legend.position = "bottom")
```

### 6. Cumulative ring width

Ring width is the most directly observable output that can be compared against
dendrometer or core data. Inter-model differences in final ring width indicate
whether the GPP divergence translates into a detectable structural outcome.

```{r ring-width, fig.height=5}
rw_df <- rbind(
  data.frame(model = "Original", year = year, doy = doy,
             rw = CASSIA_original$Growth$ring_width),
  data.frame(model = "PRELES",   year = year, doy = doy,
             rw = CASSIA_preles$Growth$ring_width),
  data.frame(model = "phydro",   year = year, doy = doy,
             rw = CASSIA_phydro$Growth$ring_width)
)
rw_df$model <- factor(rw_df$model, levels = c("Original", "PRELES", "phydro"))

ggplot(rw_df, aes(x = doy, y = rw,
                  colour = factor(year), linetype = model,
                  group  = interaction(model, year))) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = c("2015" = "#1b7837",
                                 "2016" = "#762a83",
                                 "2017" = "#e08214")) +
  scale_linetype_manual(values = c(Original = "solid",
                                   PRELES   = "dashed",
                                   phydro   = "dotted")) +
  labs(title = "Cumulative ring width by year and photosynthesis source",
       subtitle = "Solid = Original (SPP), dashed = PRELES, dotted = p-hydro",
       x = "Day of year", y = "Ring width (mm)",
       colour = "Year", linetype = "Model") +
  theme_bw() + theme(legend.position = "bottom")
```

---

## Summary

```{r summary-table}
summary_model <- function(model_out, model_name) {
  g   <- model_out$Growth
  gpp <- model_out$Preles$GPP
  total_growth <- sum(g$diameter_growth + g$needle_growth +
                      g$root_growth     + g$height_growth, na.rm = TRUE)
  annual_rw <- tapply(g$ring_width, g$year, max, na.rm = TRUE)
  data.frame(
    Model                          = model_name,
    `Mean GPP (g C m-2 day-1)`    = round(mean(gpp, na.rm = TRUE), 2),
    `Total growth (g C m-2)`      = round(total_growth, 1),
    `Growth / GPP (%)`            = round(total_growth / sum(gpp, na.rm = TRUE) * 100, 1),
    `Mean annual ring width (mm)` = round(mean(annual_rw), 3),
    check.names = FALSE
  )
}

knitr::kable(
  rbind(
    summary_model(CASSIA_original, "Original (SPP)"),
    summary_model(CASSIA_preles,   "PRELES"),
    summary_model(CASSIA_phydro,   "p-hydro")
  ),
  caption = "Summary statistics across 2015–2017"
)
```
