---
title: "Sugar model calibration: Sperling parameters"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 5)
```

## What this document does

This notebook calibrates the enzyme-kinetics sugar model (Sperling et al. 2019)
against observed NSC concentrations from Hyytiälä, then runs a set of scenarios to
explore the effect of surplus carbon and nitrogen availability on growth and
mycorrhizal allocation.

The calibration target is autumn NSC measurements from Schiestl-Aalto et al. (2019)
(`original_data`).  The optimizer is DEoptim run in parallel across all available
cores.  Parameters fitted are the enzyme kinetics (`Ad0`, `lamda`, `delta`) for each
of the five organ pools and the fractional storage capacity of each organ.

All external data are read from `~/Documents/CASSIA_Calibration/Processed_Data/`.

```{r libraries}
library(CASSIA)
library(ggplot2)
library(patchwork)
library(dplyr)
library(knitr)
library(data.table)
library(parallel)
library(DEoptim)
```

---

## 1  Setup

### 1.1  Observed data and initial parameters

```{r setup-data}
data_directory <- "~/Documents/CASSIA_Calibration/Processed_Data/"
loaded_data    <- load_data(data_directory)

new_parameters  <- rep(0.5, 27)
parameters_2015 <- initialize_parameters(calibration = FALSE, new_parameters)

parameters_2015$parameters_test[c("h0"),          "Hyde"] <- 12
parameters_2015$parameters_test[c("D0"),          "Hyde"] <- 0.1416
parameters_2015$parameters_test[c("GPP_mean"),    "Hyde"] <- 463.8833
parameters_2015$parameters_test[c("GPP_initial"), "Hyde"] <- 460.0
```

The lower bounds for each organ are set to the observed minimum so that modelled pools
cannot fall below what has been measured in the field.

```{r setup-lower-bounds}
parameters_2015$sperling_test[c("Wala.needles"),  "Hyde"] <-
  min(loaded_data$original_data$needles_sugar,  na.rm = TRUE)
parameters_2015$sperling_test[c("Wala.phloem"),   "Hyde"] <-
  min(loaded_data$original_data$phloem_sugar,   na.rm = TRUE)
parameters_2015$sperling_test[c("Wala.roots"),    "Hyde"] <-
  min(loaded_data$original_data$roots_sugar,    na.rm = TRUE)
parameters_2015$sperling_test[c("Wala.xylem_sh"), "Hyde"] <-
  min(loaded_data$original_data$xylem_sh_sugar, na.rm = TRUE)
parameters_2015$sperling_test[c("Wala.xylem_st"), "Hyde"] <-
  min(loaded_data$original_data$xylem_st_sugar, na.rm = TRUE)
```

Initial conditions are pinned to the first measurement date (2015-03-17).

```{r setup-initial-conditions}
obs_date <- as.Date("2015-03-17")
obs_row  <- rownames(loaded_data$original_data) %in% obs_date

set_ic <- function(param, col) {
  parameters_2015$sperling_test[param, "Hyde"] <<-
    loaded_data$original_data[[col]][obs_row]
}

set_ic("starch.needles0",  "needles_starch")
set_ic("starch.phloem0",   "phloem_starch")
set_ic("starch.roots0",    "roots_starch")
set_ic("starch.xylem.sh0", "xylem_sh_starch")
set_ic("starch.xylem.st0", "xylem_st_starch")
set_ic("sugar.needles0",   "needles_sugar")
set_ic("sugar.phloem0",    "phloem_sugar")
set_ic("sugar.roots0",     "roots_sugar")
set_ic("sugar.xylem.sh0",  "xylem_sh_sugar")
set_ic("sugar.xylem.st0",  "xylem_st_sugar")
```

### 1.2  Calibration target

The calibration uses the autumn measurements (first 13 observations) for all five
organ pools, sugar and starch.

```{r setup-calibration-target}
time_of_calibration <- 1:13

validate_sugar <- c(
  loaded_data$original_data$needles_sugar[time_of_calibration],
  loaded_data$original_data$needles_starch[time_of_calibration],
  loaded_data$original_data$phloem_sugar[time_of_calibration],
  loaded_data$original_data$phloem_starch[time_of_calibration],
  loaded_data$original_data$roots_sugar[time_of_calibration],
  loaded_data$original_data$roots_starch[time_of_calibration],
  loaded_data$original_data$xylem_sh_sugar[time_of_calibration],
  loaded_data$original_data$xylem_sh_starch[time_of_calibration],
  loaded_data$original_data$xylem_st_sugar[time_of_calibration],
  loaded_data$original_data$xylem_st_starch[time_of_calibration]
)
```

### 1.3  Weather data

```{r setup-weather}
weather_preles    <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")

weather_2015 <- weather_preles[
  weather_preles$date >= as.Date("2015-01-01") &
  weather_preles$date <= as.Date("2015-12-31"), ]
weather_2015$dates <- as.Date(weather_2015$date)

weather_1998_2022 <- weather_preles[
  weather_preles$date >= as.Date("1998-01-01") &
  weather_preles$date <= as.Date("2022-12-31"), ]
weather_1998_2022$dates <- as.Date(weather_1998_2022$date)

dates_2015       <- seq(as.Date("2015-01-01"), as.Date("2015-12-31"), by = "day")
dates_1998_2022  <- weather_1998_2022$dates

autumn           <- as.Date(rownames(loaded_data$original_data))[time_of_calibration]
autumn_indices   <- which(dates_2015 %in% autumn)
```

---

## 2  Initial sugar sensitivity (single-pool model)

Before calibrating the multi-pool model, it is useful to check how sensitive the
single-pool model is to the initial total sugar level.  Three values are tested:
low (0.1), medium (0.42), and high (0.75) kg C stem⁻¹.

```{r sensitivity-runs}
run_single_pool <- function(sugar00) {
  sp <- parameters_2015$sperling_test
  sp["sugar00", "Hyde"] <- sugar00
  CASSIA_cpp(
    weather                 = weather_1998_2022,
    site                    = "Hyde",
    pPREL                   = c(parameters_2015$pPREL, parameters_2015$N_parameters),
    parameters              = parameters_2015$parameters_test,
    common                  = common_p,
    ratios                  = ratios_p,
    sperling                = sp,
    needle_mass_in          = parameters_2015$needle_mass_in,
    Throughfall             = parameters_2015$Throughfall,
    photosynthesis_as_input = FALSE,
    preles                  = TRUE,
    fAPAR_Tian              = TRUE
  )
}

out_sugar_low    <- run_single_pool(0.10)
out_sugar_medium <- run_single_pool(0.42)
out_sugar_high   <- run_single_pool(0.75)
```

```{r sensitivity-plot, fig.height = 8}
scenario_colours <- c(Low = "#e41a1c", Medium = "#ff7f00", High = "#377eb8")

make_sensitivity_long <- function(col, label) {
  bind_rows(
    data.frame(dates = dates_1998_2022,
               value = out_sugar_low$Sugar[[col]],    Scenario = "Low"),
    data.frame(dates = dates_1998_2022,
               value = out_sugar_medium$Sugar[[col]], Scenario = "Medium"),
    data.frame(dates = dates_1998_2022,
               value = out_sugar_high$Sugar[[col]],   Scenario = "High")
  )
}

p_sens_su <- ggplot(make_sensitivity_long("sugar", "Sugar"),
                    aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = scenario_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Total sugar — initial value sensitivity",
       x = NULL, y = "kg C stem⁻¹", colour = "sugar₀") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_sens_st <- ggplot(make_sensitivity_long("starch", "Starch"),
                    aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = scenario_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Total starch — initial value sensitivity",
       x = NULL, y = "kg C stem⁻¹", colour = "sugar₀") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

(p_sens_su / p_sens_st) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 3  Calibration helpers

The parameter vector and DEoptim boilerplate are defined once and reused for all
calibration runs.

```{r calibration-helpers}
par_names <- c(
  "Ad0.needles",  "lamda.needles",  "delta.needles",
  "Ad0.phloem",   "lamda.phloem",   "delta.phloem",
  "Ad0.roots",    "lamda.roots",    "delta.roots",
  "Ad0.xylem.sh", "lamda.xylem.sh", "delta.xylem.sh",
  "Ad0.xylem.st", "lamda.xylem.st", "delta.xylem.st",
  "percentage_needle_storage",  "percentage_phloem_storage",
  "percentage_xylem_sh_storage","percentage_xylem_st_storage",
  "percentage_roots_storage",   "myco.thresh"
)

initial_guess <- c(
  0.438024, 0.002, 0.05,   # needles
  0.608733, 0.001, 0.01,   # phloem
  0.790206, 0.002, 0.05,   # roots
  0.7712,   0.002, 0.05,   # xylem shoot
  0.768777, 0.002, 0.05,   # xylem stem
  0.11, 0.11, 0.00001, 0.5, 0.5,
  0.9
)

lower_bounds <- rep(0, length(initial_guess))
upper_bounds <- c(rep(1, length(initial_guess) - 1), 1)

# Build parallel cluster once; reused for all DEoptim calls
ncores <- detectCores() - 1
cl     <- makeCluster(ncores)
clusterEvalQ(cl, library(CASSIA))
clusterExport(cl, varlist = c(
  "CASSIA_cpp", "parameters_2015", "weather_2015",
  "validate_sugar", "common_p", "ratios_p", "autumn_indices",
  "par_names"
))

make_objective <- function(surplus_c = FALSE) {
  force(surplus_c)
  function(par) {
    names(par) <- par_names
    sp_local   <- parameters_2015$sperling_test
    sp_local[par_names, "Hyde"] <- par

    tryCatch({
      out <- CASSIA_cpp(
        weather                 = weather_2015,
        site                    = "Hyde",
        pPREL                   = c(parameters_2015$pPREL,
                                    parameters_2015$N_parameters),
        parameters              = parameters_2015$parameters_test,
        common                  = common_p,
        ratios                  = ratios_p,
        sperling                = sp_local,
        needle_mass_in          = parameters_2015$needle_mass_in,
        Throughfall             = parameters_2015$Throughfall,
        photosynthesis_as_input = FALSE,
        preles                  = TRUE,
        fAPAR_Tian              = TRUE,
        sperling_model          = TRUE,
        surplus_c               = surplus_c
      )

      model_values <- c(
        out$Sugar$sugar_needles[autumn_indices],
        out$Sugar$starch_needles[autumn_indices],
        out$Sugar$sugar_phloem[autumn_indices],
        out$Sugar$starch_phloem[autumn_indices],
        out$Sugar$sugar_roots[autumn_indices],
        out$Sugar$starch_roots[autumn_indices],
        out$Sugar$sugar_xylem_sh[autumn_indices],
        out$Sugar$starch_xylem_sh[autumn_indices],
        out$Sugar$sugar_xylem_st[autumn_indices],
        out$Sugar$starch_xylem_st[autumn_indices]
      )

      if (any(!is.finite(model_values)) ||
          sum(!out$Culm_Growth$tree_alive) > 0) return(1e10)

      sqrt(mean((model_values - validate_sugar)^2, na.rm = TRUE))
    }, error = function(e) 1e10)
  }
}

run_deoptim <- function(objective_fn, itermax = 200) {
  DEoptim(
    fn      = objective_fn,
    lower   = lower_bounds,
    upper   = upper_bounds,
    control = DEoptim.control(
      itermax      = itermax,
      parallelType = 1,
      cluster      = cl,
      trace        = TRUE
    )
  )
}

apply_fit <- function(fit) {
  best        <- fit$optim$bestmem
  names(best) <- par_names
  parameters_2015$sperling_test[par_names, "Hyde"] <<- best
  invisible(best)
}
```

---

## 4  Calibration: standard model

The standard calibration minimises RMSE between modelled and observed autumn NSC pools
using DEoptim (200 iterations).  Running on `r ncores` cores.

```{r calibrate-standard, cache=TRUE}
fit_standard <- run_deoptim(make_objective(surplus_c = FALSE))

cat("Best RMSE:", fit_standard$optim$bestval, "\n")
kable(data.frame(Parameter = par_names,
                 Value = round(fit_standard$optim$bestmem, 6)),
      caption = "Best parameters — standard calibration")

apply_fit(fit_standard)
```

### 4.1  Single-year diagnostic (2015)

```{r run-2015-standard}
CASSIA_2015 <- CASSIA_cpp(
  weather                 = weather_2015,
  site                    = "Hyde",
  pPREL                   = c(parameters_2015$pPREL, parameters_2015$N_parameters),
  parameters              = parameters_2015$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_2015$sperling_test,
  needle_mass_in          = 0,
  Throughfall             = parameters_2015$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  sperling_model          = TRUE,
  needle_mass_grows       = TRUE
)

su_2015        <- CASSIA_2015$Sugar
su_2015$dates  <- dates_2015
obs_dates      <- as.Date(rownames(loaded_data$original_data))
```

```{r plot-2015-standard, fig.height = 10}
organ_colours <- c(
  "Needles"       = "#4daf4a",
  "Phloem"        = "#8b4513",
  "Roots"         = "#000000",
  "Xylem (shoot)" = "#e41a1c",
  "Xylem (stem)"  = "#ff7f00"
)

make_organ_plot_2015 <- function(su_col, obs_col, pool, colour) {
  obs <- data.frame(dates = obs_dates,
                    value = loaded_data$original_data[[obs_col]])
  ggplot(data.frame(dates = dates_2015, value = su_2015[[su_col]]),
         aes(x = dates, y = value)) +
    geom_line(colour = colour, linewidth = 0.9) +
    geom_point(data = obs, aes(x = dates, y = value),
               colour = colour, size = 2.5) +
    scale_x_date(date_breaks = "2 months", date_labels = "%b") +
    labs(title = pool, x = NULL, y = "kg C stem⁻¹") +
    theme_bw(base_size = 10) +
    theme(plot.title = element_text(face = "bold", size = 10))
}

pn_su  <- make_organ_plot_2015("sugar_needles",  "needles_sugar",  "Needles — sugar",       "#4daf4a")
pph_su <- make_organ_plot_2015("sugar_phloem",   "phloem_sugar",   "Phloem — sugar",        "#8b4513")
pr_su  <- make_organ_plot_2015("sugar_roots",    "roots_sugar",    "Roots — sugar",         "#000000")
pxs_su <- make_organ_plot_2015("sugar_xylem_sh", "xylem_sh_sugar", "Xylem (shoot) — sugar", "#e41a1c")
pxt_su <- make_organ_plot_2015("sugar_xylem_st", "xylem_st_sugar", "Xylem (stem) — sugar",  "#ff7f00")

pn_st  <- make_organ_plot_2015("starch_needles",  "needles_starch",  "Needles — starch",       "#4daf4a")
pph_st <- make_organ_plot_2015("starch_phloem",   "phloem_starch",   "Phloem — starch",        "#8b4513")
pr_st  <- make_organ_plot_2015("starch_roots",    "roots_starch",    "Roots — starch",         "#000000")
pxs_st <- make_organ_plot_2015("starch_xylem_sh", "xylem_sh_starch", "Xylem (shoot) — starch", "#e41a1c")
pxt_st <- make_organ_plot_2015("starch_xylem_st", "xylem_st_starch", "Xylem (stem) — starch",  "#ff7f00")

(pn_su + pph_su + pr_su) / (pxs_su + pxt_su + plot_spacer()) +
  plot_annotation(title = "2015 — sugar pools vs observations",
                  theme = theme(plot.title = element_text(face = "bold")))
```

```{r plot-2015-starch, fig.height = 10}
(pn_st + pph_st + pr_st) / (pxs_st + pxt_st + plot_spacer()) +
  plot_annotation(title = "2015 — starch pools vs observations",
                  theme = theme(plot.title = element_text(face = "bold")))
```

### 4.2  Long-run diagnostic (1998–2022)

```{r run-1998-2022-standard}
CASSIA_long <- CASSIA_cpp(
  weather                 = weather_1998_2022,
  site                    = "Hyde",
  pPREL                   = c(parameters_2015$pPREL, parameters_2015$N_parameters),
  parameters              = parameters_2015$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_2015$sperling_test,
  needle_mass_in          = 0,
  Throughfall             = parameters_2015$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  sperling_model          = TRUE,
  needle_mass_grows       = TRUE
)

su_long       <- CASSIA_long$Sugar
su_long$dates <- dates_1998_2022
```

```{r plot-1998-2022-standard, fig.height = 8}
organ_long_su <- bind_rows(
  data.frame(dates = dates_1998_2022, value = su_long$sugar_needles,  Organ = "Needles"),
  data.frame(dates = dates_1998_2022, value = su_long$sugar_phloem,   Organ = "Phloem"),
  data.frame(dates = dates_1998_2022, value = su_long$sugar_roots,    Organ = "Roots"),
  data.frame(dates = dates_1998_2022, value = su_long$sugar_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = dates_1998_2022, value = su_long$sugar_xylem_st, Organ = "Xylem (stem)")
)

organ_long_st <- bind_rows(
  data.frame(dates = dates_1998_2022, value = su_long$starch_needles,  Organ = "Needles"),
  data.frame(dates = dates_1998_2022, value = su_long$starch_phloem,   Organ = "Phloem"),
  data.frame(dates = dates_1998_2022, value = su_long$starch_roots,    Organ = "Roots"),
  data.frame(dates = dates_1998_2022, value = su_long$starch_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = dates_1998_2022, value = su_long$starch_xylem_st, Organ = "Xylem (stem)")
)

p_long_su <- ggplot(organ_long_su, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Sugar pools, 1998–2022", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_long_st <- ggplot(organ_long_st, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Starch pools, 1998–2022", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

(p_long_su / p_long_st) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 5  Biomass data

Stand-level measurements from SMEAR II are loaded here for use in scenario comparisons
below.

```{r load-biomass}
filter_smear <- function(vars) {
  loaded_data$smearII_data[loaded_data$smearII_data$variable %in% vars, ]
}

height_df     <- filter_smear(c("dominant_height",
                                "pine height BA weighted mean",
                                "pine height arithmetic mean",
                                "tree_height_baseal_area_weighted_mean"))
height_cl_df  <- filter_smear(c("pine CrownL arithmetic mean",
                                "BA weighted mean CrownL pine",
                                "pine CrownL BA weighted mean"))
diameter_df   <- filter_smear(c("BA weighted mean DBH pine",
                                "Arithmetic mean DBH pine",
                                "pine DBH arithmetic mean",
                                "pine diameter BA weighted mean"))
LAI_df        <- filter_smear(c("LAI_pine_allsided", "LAI_pine_ICOS"))
foliage_df    <- filter_smear(c("pine_foliage_biomass", "pine_foliage_biomass_ICOS",
                                "Needles / Leaves before thinning",
                                "Needles / Leaves after thinning"))

# Keep only the primary height measurement for cleaner scenario plots
height_df <- height_df[height_df$variable == "tree_height_baseal_area_weighted_mean", ]
```

---

## 6  Scenario: surplus carbon

Running with `surplus_c = TRUE` forces the tree into a conservative carbohydrate
strategy where NSC levels above optimal trigger carbon export rather than additional
growth.  A separate DEoptim calibration is run for this scenario.

```{r calibrate-surplus-c, cache=TRUE}
fit_surplus_c <- run_deoptim(make_objective(surplus_c = TRUE))

cat("Best RMSE (surplus C):", fit_surplus_c$optim$bestval, "\n")
kable(data.frame(Parameter = par_names,
                 Value = round(fit_surplus_c$optim$bestmem, 6)),
      caption = "Best parameters — surplus C calibration")

apply_fit(fit_surplus_c)
```

```{r run-surplus-c}
CASSIA_surplus_c <- CASSIA_cpp(
  weather                 = weather_1998_2022,
  site                    = "Hyde",
  pPREL                   = c(parameters_2015$pPREL, parameters_2015$N_parameters),
  parameters              = parameters_2015$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_2015$sperling_test,
  needle_mass_in          = parameters_2015$needle_mass_in,
  Throughfall             = parameters_2015$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  sperling_model          = TRUE,
  surplus_c               = TRUE
)
```

```{r plot-surplus-c, fig.height = 8}
scenario_cols_c <- c("Control" = "#4daf4a", "Surplus C" = "#e41a1c")

organs <- c("needles", "phloem", "roots", "xylem_sh", "xylem_st")

total_pool <- function(su, prefix)
  rowSums(sapply(paste0(prefix, "_", organs), function(col) su[[col]]))

sc_su_long <- bind_rows(
  data.frame(dates = dates_1998_2022,
             value = total_pool(CASSIA_long$Sugar, "sugar"),    Scenario = "Control"),
  data.frame(dates = dates_1998_2022,
             value = total_pool(CASSIA_surplus_c$Sugar, "sugar"), Scenario = "Surplus C")
)

sc_st_long <- bind_rows(
  data.frame(dates = dates_1998_2022,
             value = total_pool(CASSIA_long$Sugar, "starch"),    Scenario = "Control"),
  data.frame(dates = dates_1998_2022,
             value = total_pool(CASSIA_surplus_c$Sugar, "starch"), Scenario = "Surplus C")
)

p_sc_su <- ggplot(sc_su_long, aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = scenario_cols_c) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Total sugar (sum of organs)", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_sc_st <- ggplot(sc_st_long, aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = scenario_cols_c) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Total starch (sum of organs)", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

(p_sc_su + p_sc_st) + plot_layout(ncol = 2, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 7  Nitrogen scenarios

Three constant nitrogen levels (low: 30 %, medium: 50 %, high: 100 % of the baseline
25 kg N ha⁻¹ yr⁻¹) are run using the standard calibrated parameters.

```{r run-nitrogen}
run_nitrogen <- function(nitrogen_balance, nitrogen_change = FALSE,
                         nitrogen_contrast = FALSE) {
  CASSIA_cpp(
    weather                 = weather_1998_2022,
    site                    = "Hyde",
    pPREL                   = c(parameters_2015$pPREL, parameters_2015$N_parameters),
    parameters              = parameters_2015$parameters_test,
    common                  = common_p,
    ratios                  = ratios_p,
    sperling                = parameters_2015$sperling_test,
    needle_mass_in          = parameters_2015$needle_mass_in,
    Throughfall             = parameters_2015$Throughfall,
    photosynthesis_as_input = FALSE,
    nitrogen_balance        = nitrogen_balance,
    nitrogen_change         = nitrogen_change,
    nitrogen_contrast       = nitrogen_contrast,
    preles                  = TRUE,
    fAPAR_Tian              = TRUE,
    sperling_model          = TRUE
  )
}

CASSIA_N_high   <- run_nitrogen(25)
CASSIA_N_medium <- run_nitrogen(25 * 0.5)
CASSIA_N_low    <- run_nitrogen(25 * 0.3)
```

```{r plot-nitrogen-height, fig.height = 5}
nitrogen_cols <- c(Control = "#4daf4a", High = "#377eb8",
                   Medium  = "#ff7f00", Low  = "#e41a1c")

height_long <- bind_rows(
  data.frame(dates = dates_1998_2022,
             value = CASSIA_long$Culm_Growth$culm_growth_height,  Scenario = "Control"),
  data.frame(dates = dates_1998_2022,
             value = CASSIA_N_high$Culm_Growth$culm_growth_height, Scenario = "High"),
  data.frame(dates = dates_1998_2022,
             value = CASSIA_N_medium$Culm_Growth$culm_growth_height, Scenario = "Medium"),
  data.frame(dates = dates_1998_2022,
             value = CASSIA_N_low$Culm_Growth$culm_growth_height, Scenario = "Low")
)

obs_height <- data.frame(
  dates  = as.Date(paste0(height_df$date, "-12-31")),
  value  = height_df$amount
)

ggplot(height_long, aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  geom_point(data = obs_height, aes(x = dates, y = value),
             colour = "#4daf4a", shape = 17, size = 2.5, inherit.aes = FALSE) +
  scale_colour_manual(values = nitrogen_cols) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  coord_cartesian(ylim = c(12, NA)) +
  labs(title = "Cumulative height — nitrogen scenarios vs observations",
       x = NULL, y = "Height (m)", colour = "N level") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

```{r plot-nitrogen-myco, fig.height = 5}
myco_long <- bind_rows(
  data.frame(dates = dates_1998_2022,
             value = cumsum(CASSIA_long$Sugar$sugar_to_mycorrhiza),   Scenario = "Control"),
  data.frame(dates = dates_1998_2022,
             value = cumsum(CASSIA_N_high$Sugar$sugar_to_mycorrhiza), Scenario = "High"),
  data.frame(dates = dates_1998_2022,
             value = cumsum(CASSIA_N_medium$Sugar$sugar_to_mycorrhiza), Scenario = "Medium"),
  data.frame(dates = dates_1998_2022,
             value = cumsum(CASSIA_N_low$Sugar$sugar_to_mycorrhiza),  Scenario = "Low")
)

ggplot(myco_long, aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = nitrogen_cols) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Cumulative sugar to mycorrhiza — nitrogen scenarios",
       x = NULL, y = "kg C stem⁻¹", colour = "N level") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

---

## 8  Nitrogen with variable supply (nitrogen_change = TRUE)

A further scenario allows the nitrogen supply to vary dynamically via
`nitrogen_change = TRUE`.

```{r run-nitrogen-varies}
CASSIA_N_varies <- run_nitrogen(25 * 0.3, nitrogen_change = TRUE)
```

```{r plot-nitrogen-varies, fig.height = 5}
height_var <- bind_rows(
  data.frame(dates = dates_1998_2022,
             value = CASSIA_long$Culm_Growth$culm_growth_height,    Scenario = "Control"),
  data.frame(dates = dates_1998_2022,
             value = CASSIA_N_low$Culm_Growth$culm_growth_height,   Scenario = "Low (constant)"),
  data.frame(dates = dates_1998_2022,
             value = CASSIA_N_varies$Culm_Growth$culm_growth_height, Scenario = "Low (variable)")
)

ggplot(height_var, aes(x = dates, y = value, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  geom_point(data = obs_height, aes(x = dates, y = value),
             colour = "#4daf4a", shape = 17, size = 2.5, inherit.aes = FALSE) +
  scale_colour_manual(values = c(Control       = "#4daf4a",
                                 "Low (constant)"  = "#e41a1c",
                                 "Low (variable)"  = "#984ea3")) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  coord_cartesian(ylim = c(12, NA)) +
  labs(title = "Height — constant vs variable low nitrogen",
       x = NULL, y = "Height (m)", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

```{r cleanup-cluster}
stopCluster(cl)
```
