---
title: "Sugar and starch pools: Sperling model validation"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 5)
```

## What this document shows

This notebook validates the multi-pool sugar model (Sperling et al. 2019) by running
CASSIA with `sperling_model = TRUE` and comparing simulated organ-level sugar and starch
pools against observed concentrations from Hyytiälä.

Two simulation periods are shown:

1. **2015–2017** using the standard weather forcing (`weather_original`).
2. **1998–2022** using ERA5-derived climate data with PRELES computing GPP internally,
   and parameters calibrated for the longer period.

Observed data come from two sources: Schiestl-Aalto et al. (2019) (`original_data`) and
Yu et al. (unpublished, `yu_data`).  Both are read from an external calibration directory;
sections that depend on them are skipped automatically if that directory is absent.

Numerical agreement with the base (non-sugar) model is covered separately in
`tests_c++_against_R.Rmd`.

```{r load-libraries}
devtools::load_all(quiet = TRUE)
library(ggplot2)
library(patchwork)
library(dplyr)
library(knitr)
library(data.table)
```

---

## 1  Setup: 2015–2017 run

```{r setup-weather}
data("weather_original")

# weather_original covers 2015-2017; 2016 is a leap year: 365 + 366 + 365 = 1096 rows
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))
```

`initialize_parameters()` adjusts a small set of bounds and initial conditions relative
to the defaults in `parameters_p`; see `R/all_tests.R` for the full list.

```{r setup-parameters}
new_parameters <- rep(0.5, 27)
parameters_all <- initialize_parameters(calibration = FALSE, new_parameters)
```

```{r run-base}
out_base <- CASSIA_cpp(
  weather        = weather_original,
  site           = "Hyde",
  pPREL          = c(parameters_all$pPREL, parameters_all$N_parameters),
  parameters     = parameters_all$parameters_test,
  common         = common_p,
  ratios         = ratios_p,
  sperling       = sperling_p,
  Throughfall    = 1,
  sperling_model = TRUE,
  mycorrhiza     = FALSE
)

su <- out_base$Sugar
su$dates <- weather_original$dates
```

---

## 2  Sugar pools, 2015–2017

Each line is one organ pool.  The pink dashed line marks the bloom-threshold parameter
(`SCb`) from `sperling_p`; when phloem sugar drops below this the model can trigger
budburst (if `phloem_trigger = TRUE`).

```{r sugar-by-organ, fig.height = 8}
organ_colours <- c(
  "Needles"      = "#4daf4a",
  "Phloem"       = "#8b4513",
  "Roots"        = "#000000",
  "Xylem (shoot)" = "#e41a1c",
  "Xylem (stem)"  = "#ff7f00"
)

su_organ_long <- bind_rows(
  data.frame(dates = su$dates, value = su$sugar_needles,  Organ = "Needles"),
  data.frame(dates = su$dates, value = su$sugar_phloem,   Organ = "Phloem"),
  data.frame(dates = su$dates, value = su$sugar_roots,    Organ = "Roots"),
  data.frame(dates = su$dates, value = su$sugar_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = su$dates, value = su$sugar_xylem_st, Organ = "Xylem (stem)")
)

su_total <- data.frame(
  dates = su$dates,
  value = su$sugar_needles + su$sugar_phloem + su$sugar_roots +
          su$sugar_xylem_sh + su$sugar_xylem_st
)

SCb <- sperling_p["SCb", "Hyde"]

p_su_organs <- ggplot(su_organ_long, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = SCb, linetype = "dashed", colour = "pink", linewidth = 0.8) +
  annotate("text", x = min(su$dates), y = SCb + 0.015,
           label = "SCb (bloom threshold)", colour = "pink3", hjust = 0, size = 3) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Sugar pools by organ", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_su_total <- ggplot(su_total, aes(x = dates, y = value)) +
  geom_line(linewidth = 1.1, colour = "steelblue") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Total sugar (all organs)", x = NULL, y = "kg C stem⁻¹") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

(p_su_organs / p_su_total) + plot_layout(ncol = 1)
```

---

## 3  Starch pools, 2015–2017

```{r starch-by-organ, fig.height = 8}
st_organ_long <- bind_rows(
  data.frame(dates = su$dates, value = su$starch_needles,  Organ = "Needles"),
  data.frame(dates = su$dates, value = su$starch_phloem,   Organ = "Phloem"),
  data.frame(dates = su$dates, value = su$starch_roots,    Organ = "Roots"),
  data.frame(dates = su$dates, value = su$starch_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = su$dates, value = su$starch_xylem_st, Organ = "Xylem (stem)")
)

st_total <- data.frame(
  dates = su$dates,
  value = su$starch_needles + su$starch_phloem + su$starch_roots +
          su$starch_xylem_sh + su$starch_xylem_st
)

p_st_organs <- ggplot(st_organ_long, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Starch pools by organ", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_st_total <- ggplot(st_total, aes(x = dates, y = value)) +
  geom_line(linewidth = 1.1, colour = "darkorange") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Total starch (all organs)", x = NULL, y = "kg C stem⁻¹") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

(p_st_organs / p_st_total) + plot_layout(ncol = 1)
```

---

## 4  Comparison with observed data

Observed NSC concentrations from Schiestl-Aalto et al. (2019) (crosses) and Yu et al.
(circles) are overlaid on the simulated organ pools.  This section is skipped
automatically if the external calibration data directory is not present.

```{r load-obs}
data_directory <- "~/Documents/CASSIA_Calibration/Processed_Data/"
has_obs <- dir.exists(data_directory)
if (has_obs) {
  loaded_data   <- load_data(data_directory)
  original_data <- loaded_data$original_data
  yu_data       <- loaded_data$yu_data

  obs_sa <- data.frame(
    dates      = as.Date(rownames(original_data)),
    su_needles = original_data$needles_sugar,
    su_phloem  = original_data$phloem_sugar,
    su_roots   = original_data$roots_sugar,
    su_xsh     = original_data$xylem_sh_sugar,
    su_xst     = original_data$xylem_st_sugar,
    st_needles = original_data$needles_starch,
    st_phloem  = original_data$phloem_starch,
    st_roots   = original_data$roots_starch,
    st_xsh     = original_data$xylem_sh_starch,
    st_xst     = original_data$xylem_st_starch
  )

  obs_yu <- data.frame(
    dates      = as.Date(yu_data$Date),
    su_needles = yu_data$sugar.needles,
    su_phloem  = yu_data$sugar.phloem,
    su_roots   = yu_data$sugar.root,
    su_xsh     = yu_data$sugar.xylem.sh,
    su_xst     = yu_data$sugar.xylem.st,
    st_needles = yu_data$starch.needles,
    st_phloem  = yu_data$starch.phloem,
    st_roots   = yu_data$starch.root,
    st_xsh     = yu_data$starch.xylem.sh,
    st_xst     = yu_data$starch.xylem.st
  )
}
```

```{r obs-sugar, eval = has_obs, fig.height = 10}
make_organ_plot <- function(sim_col, sa_col, yu_col,
                            line_colour, point_colour, title) {
  ggplot(su, aes(x = dates, y = .data[[sim_col]])) +
    geom_line(colour = line_colour, linewidth = 0.9) +
    geom_point(data = obs_sa, aes(x = dates, y = .data[[sa_col]]),
               shape = 4, colour = point_colour, size = 2.5) +
    geom_point(data = obs_yu, aes(x = dates, y = .data[[yu_col]]),
               shape = 1, colour = point_colour, size = 2.5) +
    scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
    labs(title = title, x = NULL, y = "kg C stem⁻¹") +
    theme_bw(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 10))
}

pn  <- make_organ_plot("sugar_needles",  "su_needles", "su_needles",
                       "steelblue", "#4daf4a", "Needles")
pph <- make_organ_plot("sugar_phloem",   "su_phloem",  "su_phloem",
                       "steelblue", "#8b4513", "Phloem")
pr  <- make_organ_plot("sugar_roots",    "su_roots",   "su_roots",
                       "steelblue", "#000000", "Roots")
pxs <- make_organ_plot("sugar_xylem_sh", "su_xsh",     "su_xsh",
                       "steelblue", "#e41a1c", "Xylem (shoot)")
pxt <- make_organ_plot("sugar_xylem_st", "su_xst",     "su_xst",
                       "steelblue", "#ff7f00", "Xylem (stem)")

(pn + pph) / (pr + pxs + pxt) +
  plot_annotation(
    title = "Sugar pools vs observations  (× = Schiestl-Aalto et al. 2019, ○ = Yu et al.)",
    theme = theme(plot.title = element_text(face = "bold"))
  )
```

```{r obs-starch, eval = has_obs, fig.height = 10}
qn  <- make_organ_plot("starch_needles",  "st_needles", "st_needles",
                       "darkorange", "#4daf4a", "Needles")
qph <- make_organ_plot("starch_phloem",   "st_phloem",  "st_phloem",
                       "darkorange", "#8b4513", "Phloem")
qr  <- make_organ_plot("starch_roots",    "st_roots",   "st_roots",
                       "darkorange", "#000000", "Roots")
qxs <- make_organ_plot("starch_xylem_sh", "st_xsh",     "st_xsh",
                       "darkorange", "#e41a1c", "Xylem (shoot)")
qxt <- make_organ_plot("starch_xylem_st", "st_xst",     "st_xst",
                       "darkorange", "#ff7f00", "Xylem (stem)")

(qn + qph) / (qr + qxs + qxt) +
  plot_annotation(
    title = "Starch pools vs observations  (× = Schiestl-Aalto et al. 2019, ○ = Yu et al.)",
    theme = theme(plot.title = element_text(face = "bold"))
  )
```

---

## 5  Long run: 1998–2022 with ERA5 / PRELES

This section uses ERA5-derived climate data and runs PRELES internally to generate GPP,
with parameters calibrated for the 1998–2022 period.

```{r setup-preles-weather}
weather_preles <- data.table::fread("~/Documents/CASSIA/data/ERA5_CASSIA_preles.csv")
weather_preles_1998_2022 <- weather_preles[
  substring(weather_preles$date, 1, 4) %in% 1998:2022, ]
weather_preles_1998_2022$dates <- as.Date(weather_preles_1998_2022$date)
```

```{r setup-preles-parameters}
parameters_long <- initialize_parameters(calibration = FALSE, new_parameters)

parameters_long$parameters_test[c("h0"),          "Hyde"] <- 12
parameters_long$parameters_test[c("D0"),          "Hyde"] <- 0.1416
parameters_long$parameters_test[c("GPP_mean"),    "Hyde"] <- 463.8833
parameters_long$parameters_test[c("GPP_initial"), "Hyde"] <- 460.0

parameters_long$sperling_test[
  c("Ad0.needles", "Ad0.phloem", "Ad0.roots", "Ad0.xylem.sh", "Ad0.xylem.st"),
  "Hyde"] <- c(0.1, 0.1, 0.1, 0.1, 0.1)

parameters_long$sperling_test[
  c("lamda.needles", "lamda.phloem", "lamda.roots", "lamda.xylem.sh", "lamda.xylem.st"),
  "Hyde"] <- c(1e-4, 1e-4, 1e-4, 1e-4, 1e-4)

parameters_long$sperling_test[
  c("delta.needles", "delta.phloem", "delta.roots", "delta.xylem.sh", "delta.xylem.st"),
  "Hyde"] <- c(25e-6, 25e-6, 25e-6, 25e-6, 25e-6)

parameters_long$sperling_test["myco.thresh", "Hyde"] <- 0.3
```

```{r run-preles}
out_long <- CASSIA_cpp(
  weather                 = weather_preles_1998_2022,
  site                    = "Hyde",
  pPREL                   = c(parameters_long$pPREL, parameters_long$N_parameters),
  parameters              = parameters_long$parameters_test,
  common                  = common_p,
  ratios                  = ratios_p,
  sperling                = parameters_long$sperling_test,
  needle_mass_in          = parameters_long$needle_mass_in,
  Throughfall             = parameters_long$Throughfall,
  photosynthesis_as_input = FALSE,
  preles                  = TRUE,
  fAPAR_Tian              = TRUE,
  sperling_model          = TRUE
)

su_long <- out_long$Sugar
su_long$dates <- weather_preles_1998_2022$dates
```

```{r preles-organ-plots, fig.height = 8}
su_long_organ <- bind_rows(
  data.frame(dates = su_long$dates, value = su_long$sugar_needles,  Organ = "Needles"),
  data.frame(dates = su_long$dates, value = su_long$sugar_phloem,   Organ = "Phloem"),
  data.frame(dates = su_long$dates, value = su_long$sugar_roots,    Organ = "Roots"),
  data.frame(dates = su_long$dates, value = su_long$sugar_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = su_long$dates, value = su_long$sugar_xylem_st, Organ = "Xylem (stem)")
)

st_long_organ <- bind_rows(
  data.frame(dates = su_long$dates, value = su_long$starch_needles,  Organ = "Needles"),
  data.frame(dates = su_long$dates, value = su_long$starch_phloem,   Organ = "Phloem"),
  data.frame(dates = su_long$dates, value = su_long$starch_roots,    Organ = "Roots"),
  data.frame(dates = su_long$dates, value = su_long$starch_xylem_sh, Organ = "Xylem (shoot)"),
  data.frame(dates = su_long$dates, value = su_long$starch_xylem_st, Organ = "Xylem (stem)")
)

p_long_su <- ggplot(su_long_organ, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Sugar pools, 1998–2022 (ERA5 / PRELES)",
       x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

p_long_st <- ggplot(st_long_organ, aes(x = dates, y = value, colour = Organ)) +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = organ_colours) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Starch pools, 1998–2022 (ERA5 / PRELES)",
       x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

(p_long_su / p_long_st) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```

```{r preles-obs-needle, eval = has_obs}
ggplot(data.frame(dates = su_long$dates, value = su_long$sugar_needles),
       aes(x = dates, y = value)) +
  geom_line(colour = "steelblue", linewidth = 0.8) +
  geom_point(data = obs_sa, aes(x = dates, y = su_needles),
             shape = 4, colour = "#4daf4a", size = 2.5) +
  geom_point(data = obs_yu, aes(x = dates, y = su_needles),
             shape = 1, colour = "#4daf4a", size = 2.5) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "Needle sugar, 1998–2022 vs observations  (× = Schiestl-Aalto, ○ = Yu)",
       x = NULL, y = "kg C stem⁻¹") +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))
```
