---
title: "Sugar model: individual functions (R reference)"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 4)
library(ggplot2)

load("../data/weather_original.RData")
# 2015 is the first 365 rows
weather_2015 <- weather_original[1:365, ]
doy          <- weather_2015$X
```

## Overview

This notebook contains R implementations of the individual functions that make
up the CASSIA sugar model (`src/sugar_model.cpp`). Each function is explored in
isolation using the built-in `weather_original` dataset (Hyytiälä SMEAR II,
2015). The C++ function name is given in each section header so the two versions
can be compared directly.

---

## 1  Sugar limitation factor

**C++ equivalent:** `storage_update_organs()` in `sugar_model.cpp`

Given a sugar pool $s$, an organ-specific lower reserve limit $l$, and the
potential growth + maintenance demand for that day $g$, the storage term is the
fraction of growth that can be satisfied:

$$
k(s,\, l,\, g) = \begin{cases}
0 & s \leq l \text{ or } g \leq 0 \\
\min\!\left(1,\; \dfrac{s - l}{g}\right) & s > l
\end{cases}
$$

```{r sugar-limitation}
sugar_limitation <- function(s, l, g) {
  if (g <= 0) return(0)
  available <- s - l
  if (available <= 0) return(0)
  min(1, available / g)
}

# Vectorised wrapper for plotting
sugar_lim_vec <- function(s, l, g) pmax(0, pmin(1, (s - l) / pmax(g, 1e-10))) * (s > l) * (g > 0)

sugar <- seq(0, 1, length.out = 200)

par(mfrow = c(1, 2))
plot(sugar, sugar_lim_vec(sugar, 0.1, 0.5), type = "l", col = "blue",
     ylim = c(0, 1), xlab = "Sugar pool (kg C)", ylab = "Storage term k",
     main = "Effect of reserve limit l\n(demand g = 0.5 fixed)")
lines(sugar, sugar_lim_vec(sugar, 0.2, 0.5))
lines(sugar, sugar_lim_vec(sugar, 0.3, 0.5), col = "red")
legend("bottomright", legend = c("l = 0.1", "l = 0.2", "l = 0.3"),
       col = c("blue", "black", "red"), lty = 1, bty = "n", cex = 0.8)

plot(sugar, sugar_lim_vec(sugar, 0.2, 0.25), type = "l",
     ylim = c(0, 1), xlab = "Sugar pool (kg C)", ylab = "Storage term k",
     main = "Effect of growth demand g\n(reserve l = 0.2 fixed)")
lines(sugar, sugar_lim_vec(sugar, 0.2, 0.50), col = "green")
lines(sugar, sugar_lim_vec(sugar, 0.2, 0.75), col = "purple")
legend("bottomright", legend = c("g = 0.25", "g = 0.50", "g = 0.75"),
       col = c("black", "green", "purple"), lty = 1, bty = "n", cex = 0.8)
par(mfrow = c(1, 1))
```

---

## 2  Winter reserve scaling

**C++ equivalent:** `winter_factor()` in `sugar_model.cpp`

The winter reserve $w$ scales the fraction of each organ's sugar pool that must
be retained as antifreeze protection. Rather than responding instantaneously,
$w$ tracks a physiological target $w^*$ with a first-order lag (time constant
$\tau = 7$ days). After each update $w$ is clamped to $[0, 1]$:

$$
w^* = 1 - \frac{1}{1+e^{-(T-T_c)/k_T}} \cdot \frac{1}{1+e^{-(R-R_c)/k_R}}, \qquad
w_t = \text{clamp}\!\left(w_{t-1} + \frac{w^*-w_{t-1}}{\tau},\; 0, 1\right)
$$

Default parameters: $T_c = 5\,^\circ\text{C}$, $R_c = 30\;\mu\text{mol m}^{-2}
\text{s}^{-1}$, $k_T = 2$, $k_R = 10$, $\tau = 7$.

```{r winter-factor}
w_target <- function(T_air, PAR, T_crit = 5, PAR_crit = 30, kT = 2, kPAR = 10) {
  1 - (1 / (1 + exp(-(T_air - T_crit) / kT))) * (1 / (1 + exp(-(PAR - PAR_crit) / kPAR)))
}

update_w <- function(w_prev, T_air, PAR, tau = 7) {
  w_star <- w_target(T_air, PAR)
  pmin(pmax(w_prev + (w_star - w_prev) / tau, 0), 1)   # clamp to [0, 1] as in C++
}

w_star_daily <- w_target(weather_2015$T, weather_2015$PAR)

w <- numeric(365)
w[1] <- w_star_daily[1]
for (i in 2:365) {
  w[i] <- update_w(w[i - 1], weather_2015$T[i], weather_2015$PAR[i])
}

plot(doy, w_star_daily, type = "l", col = "grey60", lwd = 1,
     ylim = c(0, 1),
     xlab = "Day of year (Hyytiälä 2015)",
     ylab = "Winter reserve scaling w",
     main = "Winter reserve scaling (τ = 7 d lag, clamped to [0, 1])")
lines(doy, w, col = "blue", lwd = 2)
legend("bottomright",
       legend = c("Instantaneous target w*", "Delayed w (τ = 7 d)"),
       col = c("grey60", "blue"), lwd = c(1, 2), bty = "n", cex = 0.8)
```

---

## 3  Sugar–starch flux: current C++ formulation

**C++ equivalent:** `safe_to_starch_transfer()` in `sugar_model.cpp`

The daily net flux from sugar to starch is computed by two simple linear terms:

- **Synthesis**: a fixed 30% of the sugar surplus above the winter reserve is
  converted to starch each day.
- **Remobilisation**: if sugar falls below the winter reserve, starch is drawn
  down at a rate of 30% of the shortfall, capped by the available starch.

After both terms, three physical limits are applied: the flux cannot exceed the
sugar surplus, it cannot push starch above its capacity, and it cannot drain
more starch than exists.

$$
S = \max(s - w,\; 0), \qquad
F_\text{syn} = 0.3 \cdot S
$$

$$
F_\text{rem} = \begin{cases}
\min\!\left(\theta,\; 0.3(w - s)\right) & s < w \\
0 & \text{otherwise}
\end{cases}
$$

$$
F_\text{net} = \text{clamp}\!\left(F_\text{syn} - F_\text{rem},\; -\theta,\; \min(S,\; \Theta - \theta)\right)
$$

where $s$ = sugar, $w$ = winter reserve, $\theta$ = current starch, $\Theta$ =
starch capacity.

```{r safe-to-starch}
safe_to_starch_transfer <- function(sugar, winter, starch, max_capacity) {
  S             <- max(sugar - winter, 0)
  capacity_free <- max(max_capacity - starch, 0)

  synthesis     <- 0.3 * S

  remobilization <- 0
  if (sugar < winter) {
    remobilization <- min(starch, 0.3 * (winter - sugar))
  }

  flux <- synthesis - remobilization
  flux <- min(flux, S)             # cannot store more sugar than is available
  flux <- min(flux, capacity_free) # cannot exceed starch capacity
  flux <- max(flux, -starch)       # cannot drain more starch than exists
  flux
}

# Vectorised for plotting
stt_vec <- function(sugar, winter, starch, max_capacity) {
  mapply(safe_to_starch_transfer, sugar, winter, starch, max_capacity)
}

sugar_vals <- seq(0, 1, length.out = 200)
winter_fixed  <- 0.2
max_cap       <- 1.0

par(mfrow = c(1, 2))
plot(sugar_vals, stt_vec(sugar_vals, winter_fixed, 0.25, max_cap),
     type = "l", col = "yellow3", ylim = c(-0.2, 0.3),
     xlab = "Sugar (kg C)", ylab = "Net flux to starch (kg C)",
     main = "Net flux vs sugar\n(max_capacity = 1.0, winter = 0.2)")
lines(sugar_vals, stt_vec(sugar_vals, winter_fixed, 0.50, max_cap), col = "orange")
lines(sugar_vals, stt_vec(sugar_vals, winter_fixed, 0.75, max_cap), col = "red")
abline(v = winter_fixed, lty = 2, col = "grey50"); abline(h = 0, lty = 2, col = "grey50")
legend("topleft", legend = c("Starch = 0.25", "Starch = 0.50", "Starch = 0.75"),
       col = c("yellow3", "orange", "red"), lty = 1, bty = "n", cex = 0.8)

starch_fixed <- 0.5
plot(sugar_vals, stt_vec(sugar_vals, 0.1, starch_fixed, max_cap),
     type = "l", col = "blue", ylim = c(-0.2, 0.3),
     xlab = "Sugar (kg C)", ylab = "Net flux to starch (kg C)",
     main = "Effect of winter reserve w\n(starch = 0.5, max_capacity = 1.0)")
lines(sugar_vals, stt_vec(sugar_vals, 0.2, starch_fixed, max_cap))
lines(sugar_vals, stt_vec(sugar_vals, 0.4, starch_fixed, max_cap), col = "red")
abline(h = 0, lty = 2, col = "grey50")
legend("topleft", legend = c("w = 0.1", "w = 0.2", "w = 0.4"),
       col = c("blue", "black", "red"), lty = 1, bty = "n", cex = 0.8)
par(mfrow = c(1, 1))
```

### Seasonal behaviour using 2015 weather

The daily winter reserve from Section 2 is used to drive the flux, with sugar
and starch pools held constant to isolate the seasonal pattern:

```{r starch-seasonal}
flux_series <- stt_vec(0.4, w, 0.3, 1.0)

plot(doy, flux_series, type = "l", col = "steelblue", lwd = 2,
     xlab = "Day of year (Hyytiälä 2015)", ylab = "Net flux to starch (kg C)",
     main = "Daily sugar→starch flux driven by real winter reserve\n(sugar = 0.4, starch = 0.3, capacity = 1.0)")
abline(h = 0, lty = 2, col = "grey50")
lines(doy, w * 0.3, col = "grey70", lty = 3)
legend("topright", legend = c("Net flux to starch", "Winter reserve w (×0.3, for scale)"),
       col = c("steelblue", "grey70"), lty = c(1, 3), bty = "n", cex = 0.8)
```

---

## 4  Sugar–starch flux: prototype Michaelis–Menten formulation

The function below was an earlier prototype using Michaelis–Menten scalers for
both the availability of surplus sugar and the remaining storage capacity. It is
retained here for comparison and exploration; the C++ uses the simpler linear
formulation in Section 3.

```{r F-function-def}
F_function <- function(Ss, St, V, Theta, A0, delta, eps = 1e-6) {
  s           <- pmax(0, Ss - V)
  K_s         <- 0.25 * Theta
  sigma_s     <- s / (s + K_s)
  Theta_free  <- pmax(0, Theta - St)
  sigma_Theta <- Theta_free / (Theta_free + eps)
  F_syn <- A0    * sigma_s       * sigma_Theta
  F_rem <- delta * St * (1 - sigma_s)
  F_syn - F_rem
}
```

```{r F-function-plots, fig.height=4}
V <- 0.1; Theta <- 1.0; A0 <- 0.9; delta <- 0.9
starch <- seq(0, 1, length.out = 200)

par(mfrow = c(1, 2))
plot(sugar_vals, F_function(sugar_vals, 0.25, V, Theta, A0, delta),
     type = "l", col = "yellow3", ylim = c(-1, 1),
     xlab = "Sugar (kg C)", ylab = "Net flux to starch",
     main = "Prototype: flux vs sugar\n(Michaelis–Menten scalers)")
lines(sugar_vals, F_function(sugar_vals, 0.50, V, Theta, A0, delta), col = "orange")
lines(sugar_vals, F_function(sugar_vals, 0.75, V, Theta, A0, delta), col = "red")
abline(h = 0, lty = 2, col = "grey50")
legend("topleft", legend = c("St = 0.25", "St = 0.50", "St = 0.75"),
       col = c("yellow3", "orange", "red"), lty = 1, bty = "n", cex = 0.8)

plot(starch, F_function(0.25, starch, V, Theta, A0, delta),
     type = "l", col = "yellow3", ylim = c(-1, 1),
     xlab = "Starch (kg C)", ylab = "Net flux to starch",
     main = "Prototype: flux vs starch")
lines(starch, F_function(0.50, starch, V, Theta, A0, delta), col = "orange")
lines(starch, F_function(0.75, starch, V, Theta, A0, delta), col = "red")
abline(h = 0, lty = 2, col = "grey50")
legend("topright", legend = c("Ss = 0.25", "Ss = 0.50", "Ss = 0.75"),
       col = c("yellow3", "orange", "red"), lty = 1, bty = "n", cex = 0.8)
par(mfrow = c(1, 1))
```

---

## 5  Mycorrhizal nitrogen transfer

**C++ equivalent:** `nitrogen_transfer_from_mycorrhiza()` in `sugar_model.cpp`

The fraction of mycorrhizal nitrogen transferred to the tree is a Michaelis–
Menten saturating function of the sugar exported to the fungi:

$$
f_m(S_m) = \frac{S_m}{S_m + \beta_m}
$$

The C++ implementation hardcodes $\beta_m = 5 \times 10^{-5}$ (`sugar_half_saturation`).

```{r f-mycorrhiza}
f_m <- function(S_m, beta_m = 0.00005) S_m / (S_m + beta_m)

S_m_vals <- seq(0, 0.001, length.out = 200)

beta_vals <- c(0.00005, 0.0002, 0.001)
cols      <- c("blue", "orange", "red")

plot(S_m_vals, f_m(S_m_vals, beta_vals[1]), type = "l", col = cols[1], lwd = 2,
     ylim = c(0, 1),
     xlab = "Sugar transferred to mycorrhiza (kg C)",
     ylab = "Fraction of mycorrhizal N transferred",
     main = sprintf("Mycorrhizal N transfer (C++ default β_m = %.5f)", 0.00005))
for (i in 2:length(beta_vals)) {
  lines(S_m_vals, f_m(S_m_vals, beta_vals[i]), col = cols[i], lwd = 2)
}
legend("bottomright",
       legend = sapply(beta_vals, function(b) bquote(beta[m] == .(b))),
       col = cols, lwd = 2, bty = "n")
```

---

## 6  Mycorrhizal and starch seasonality

**C++ equivalent:** inline in `sugar_model()` in `sugar_model.cpp`

The C++ model uses two hardcoded logistic functions to scale mycorrhizal
allocation and starch demand through the growing season. These are not separate
C++ functions — they are computed inline inside `sugar_model()`.

**Mycorrhizal allocation seasonality** (sugar_model.cpp lines 416–418):

```cpp
double spring = 1.0 / (1.0 + exp(-0.075 * (day - 130)));
double autumn = 1.0 / (1.0 + exp( 0.075 * (day - 220)));
double allocation = spring * autumn;
```

**Starch demand seasonality** (line 606) uses a slightly earlier autumn cutoff
(day 200 instead of 220):

```cpp
double star = 1.0 / (1.0 + exp(-0.075 * (day - 130)))
            * 1.0 / (1.0 + exp( 0.075 * (day - 200)));
```

```{r myco-seasonality}
spring    <- 1 / (1 + exp(-0.075 * (doy - 130)))
autumn    <- 1 / (1 + exp( 0.075 * (doy - 220)))
myco_seas <- spring * autumn

star <- 1 / (1 + exp(-0.075 * (doy - 130))) *
        1 / (1 + exp( 0.075 * (doy - 200)))

plot(doy, myco_seas, type = "l", col = "blue", lwd = 2,
     ylim = c(0, 1), xlab = "Day of year (Hyytiälä 2015)",
     ylab = "Seasonality scaling (0–1)",
     main = "C++ inline seasonality scalers")
lines(doy, star, col = "red", lwd = 2, lty = 2)
legend("topright",
       legend = c("Mycorrhizal allocation (autumn cutoff day 220)",
                  "Starch demand (autumn cutoff day 200)"),
       col = c("blue", "red"), lwd = 2, lty = c(1, 2), bty = "n", cex = 0.9)
```

### Seasonality with Hyytiälä 2015 temperature

```{r myco-with-weather, fig.height=4}
par(mar = c(5, 4, 4, 4) + 0.1)
plot(doy, weather_2015$T, type = "l", col = "grey60",
     xlab = "Day of year (2015)", ylab = "Temperature (°C)",
     main = "Mycorrhizal seasonality vs 2015 temperature")
par(new = TRUE)
plot(doy, myco_seas, type = "l", col = "blue", lwd = 2,
     axes = FALSE, xlab = "", ylab = "")
axis(side = 4, col.axis = "blue")
mtext("Seasonality scaling", side = 4, line = 3, col = "blue")
legend("topleft", legend = c("Temperature (°C)", "Myco seasonality"),
       col = c("grey60", "blue"), lty = 1, lwd = c(1, 2), bty = "n")
```
