---
title: "Below-ground sugar allocation: nitrogen-feedback optimisation"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 4)
library(Rsolnp)
library(ggplot2)
library(reshape2)
library(viridis)
library(gridExtra)
```

## Background

CASSIA includes an organ-level sugar model (Sperling et al. 2019) that tracks
non-structural carbohydrate pools in needles, phloem, xylem, and roots. The
full transfer equations are given in the CASSIA Instruction Booklet
(§ Sugar and starch transfer). In brief, each organ exchanges sugar with the
phloem via a concentration-gradient diffusion term, and growth and respiration
draw down the pool each day.

One open question in that framework is how much sugar the tree should allocate
**below ground** — to the root nitrogen-uptake pathway versus to mycorrhizal
fungi — and whether that allocation should be driven by a fixed proportion of
the phloem pool or by a target nitrogen demand. The Instruction Booklet
describes two strategies (§ Below-ground allocation):

1. **Empirical fixed fraction**: approximately 30% of phloem sugar is
   exported below ground, matching measured carbohydrate budgets. This is the
   current default in the model.

2. **Nitrogen-feedback**: the below-ground allocation is chosen to meet a
   nitrogen uptake target, minimising total carbon spent. If the target cannot
   be met within the available carbon budget, the tree maximises nitrogen
   uptake instead.

This notebook explores the second strategy. It formulates and solves the
nitrogen-feedback allocation as a constrained optimisation, then shows how
the solution changes across the space of root and mycorrhizal nitrogen
uptake capacities.

## Model formulation

Let $c_s$ (g C) be the carbon allocated to the root nitrogen-uptake pathway
and $c_m$ (g C) be the carbon allocated to mycorrhizal fungi. These appear
as $c_s^{Ss}$ and $c_m^{Ss}$ respectively in the phloem sugar equation
(Instruction Booklet, Eq. for $dc_{\text{phl}}^{Ss}/dt$).

Total nitrogen uptake is

$$
N(c_s,\, c_m) = \frac{c_s}{c_s + c_m + \varepsilon}\, N_r \;+\; \frac{c_m}{c_m + \beta_m}\, N_m
$$

where:

| Symbol | Meaning | Units |
|--------|---------|-------|
| $N_r$ | Maximum nitrogen uptake rate via roots | g N |
| $N_m$ | Maximum nitrogen uptake rate via mycorrhizal fungi | g N |
| $\beta_m$ | Michaelis–Menten half-saturation constant for mycorrhizal uptake | g C |
| $\varepsilon$ | Small regularisation constant preventing division by zero | — |

The first term describes root uptake: the fraction of below-ground carbon going
to the root pathway scales the available root nitrogen supply. The second term
describes mycorrhizal uptake using Michaelis–Menten kinetics, reflecting that
fungal nitrogen acquisition saturates at high carbon investment.

The optimisation is:

$$
\min_{c_s,\, c_m} \; (c_s + c_m)
\quad \text{subject to} \quad
N(c_s, c_m) = N_T, \quad c_s + c_m \leq c_I, \quad c_s, c_m \geq 0
$$

where $N_T$ is the nitrogen demand and $c_I$ is the total carbon available for
below-ground allocation. If no feasible point exists (i.e. $N_T$ cannot be
reached within the carbon budget), the problem switches to maximising $N$
subject only to the budget constraint.

---

## Single-scenario optimisation

The example below uses fixed parameter values to illustrate how the optimiser
finds the minimum-carbon allocation that meets a nitrogen target.

```{r single-opt}
N_r    <- 10     # root N uptake capacity (g N)
N_m    <- 100    # mycorrhizal N uptake capacity (g N)
epsilon <- 1e-6  # numerical regularisation
beta_m  <- 0.1   # Michaelis-Menten half-saturation (g C)
N_T    <- 8      # nitrogen demand (g N)
c_I    <- 5      # carbon available for below-ground allocation (g C)

N_fun <- function(c) {
  c_s <- c[1]; c_m <- c[2]
  (c_s / (c_s + c_m + epsilon)) * N_r +
    (c_m / (c_m + beta_m)) * N_m
}

# Step 1: minimise carbon spend while meeting nitrogen target
res1 <- try(
  solnp(
    pars   = c(0.1, 0.1),
    fun    = function(c) sum(c),
    eqfun  = function(c) N_fun(c) - N_T,
    eqB    = 0,
    ineqfun = function(c) c_I - sum(c),
    ineqLB = 0,
    ineqUB = Inf,
    LB = c(0, 0),
    UB = c(c_I, c_I)
  ),
  silent = TRUE
)

if (!inherits(res1, "try-error") && abs(N_fun(res1$pars) - N_T) < 1e-4) {
  c_opt <- res1$pars
  mode  <- "Target nitrogen met — minimum carbon solution."
} else {
  # Step 2: target infeasible; maximise achievable nitrogen
  res2 <- solnp(
    pars    = c(0.1, 0.1),
    fun     = function(c) -N_fun(c),
    ineqfun = function(c) c_I - sum(c),
    ineqLB  = 0,
    ineqUB  = Inf,
    LB = c(0, 0),
    UB = c(c_I, c_I)
  )
  c_opt <- res2$pars
  mode  <- "Target nitrogen NOT achievable within budget — nitrogen maximised."
}

cat(mode, "\n\n")
cat(sprintf("  Root pathway carbon  c_s = %.4f g C\n", c_opt[1]))
cat(sprintf("  Mycorrhizal carbon   c_m = %.4f g C\n", c_opt[2]))
cat(sprintf("  Total carbon spent       = %.4f g C\n", sum(c_opt)))
cat(sprintf("  Nitrogen achieved        = %.4f g N\n", N_fun(c_opt)))
```

The solution shows that, with these parameter values, the nitrogen target can
be met by preferentially investing in the mycorrhizal pathway (large $c_m$),
which has a higher maximum uptake capacity ($N_m = 100$) but saturates. A
small root-pathway investment ($c_s$) is still needed because the root term
provides nitrogen proportional to the carbon fraction it receives.

---

## Sensitivity to uptake capacities $N_r$ and $N_m$

In practice, $N_r$ and $N_m$ vary with soil nitrogen availability, root
architecture, and fungal community composition. The heatmaps below show how
the optimal allocation ($c_s^*$, $c_m^*$) changes across the two-dimensional
space of these uptake capacities, with all other parameters held at the
single-scenario values.

### Method

For each $(N_r, N_m)$ combination the constraint set is scanned over a fine
grid of $c_m$ values; for each $c_m$ the equality $N(c_s, c_m) = N_T$ is
solved analytically for $c_s$. If any feasible point exists the minimum-carbon
pair is selected. If no feasible point exists the maximum-nitrogen allocation
is found by brute force over the budget boundary.

```{r sensitivity-grid, cache=TRUE}
N_r_vals <- seq(0, 10, length.out = 50)
N_m_vals <- seq(0, 10, length.out = 50)
grid     <- expand.grid(N_r = N_r_vals, N_m = N_m_vals)

grid$c_s_opt   <- NA_real_
grid$c_m_opt   <- NA_real_
grid$N_achieved <- NA_real_

c_m_seq <- seq(0, c_I, length.out = 500)

for (i in seq_len(nrow(grid))) {
  N_r_i <- grid$N_r[i]
  N_m_i <- grid$N_m[i]

  feasible <- data.frame(c_s = numeric(0), c_m = numeric(0))

  for (c_m in c_m_seq) {
    term2 <- (c_m / (c_m + beta_m)) * N_m_i
    if (N_r_i == 0) {
      if (abs(term2 - N_T) < 1e-6) c_s <- 0 else next
    } else {
      c_s <- (N_T - term2) * (c_m + epsilon) / (N_r_i - (N_T - term2))
    }
    if (is.na(c_s) || c_s < 0 || c_s + c_m > c_I) next
    feasible <- rbind(feasible, data.frame(c_s = c_s, c_m = c_m))
  }

  if (nrow(feasible) > 0) {
    best_idx         <- which.min(feasible$c_s + feasible$c_m)
    grid$c_s_opt[i]  <- feasible$c_s[best_idx]
    grid$c_m_opt[i]  <- feasible$c_m[best_idx]
    grid$N_achieved[i] <- N_T
  } else {
    # Budget insufficient for N_T — find the maximum achievable N
    best_N <- -Inf; best_cs <- NA; best_cm <- NA
    for (c_m in c_m_seq) {
      max_cs <- c_I - c_m
      if (max_cs < 0) next
      N_try <- (max_cs / (max_cs + c_m + epsilon)) * N_r_i +
               (c_m / (c_m + beta_m)) * N_m_i
      if (N_try > best_N) { best_N <- N_try; best_cs <- max_cs; best_cm <- c_m }
    }
    grid$c_s_opt[i]   <- best_cs
    grid$c_m_opt[i]   <- best_cm
    grid$N_achieved[i] <- best_N
  }
}
```

### Optimal carbon allocation heatmaps

White cells indicate parameter combinations for which the nitrogen target
$N_T = `r N_T`$ g N cannot be met even at the full carbon budget $c_I =
`r c_I`$ g C.

```{r sensitivity-plots, fig.height=4}
c_s_melt <- melt(grid, id.vars = c("N_r", "N_m"), measure.vars = "c_s_opt")
c_m_melt <- melt(grid, id.vars = c("N_r", "N_m"), measure.vars = "c_m_opt")

p1 <- ggplot(c_s_melt, aes(x = N_r, y = N_m, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", na.value = "white") +
  labs(title = "Optimal root-pathway carbon (c_s)",
       subtitle = "White = N target infeasible",
       x = "Root N capacity (N_r)", y = "Mycorrhizal N capacity (N_m)",
       fill = "c_s (g C)") +
  theme_minimal()

p2 <- ggplot(c_m_melt, aes(x = N_r, y = N_m, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", na.value = "white") +
  labs(title = "Optimal mycorrhizal carbon (c_m)",
       subtitle = "White = N target infeasible",
       x = "Root N capacity (N_r)", y = "Mycorrhizal N capacity (N_m)",
       fill = "c_m (g C)") +
  theme_minimal()

grid.arrange(p1, p2, ncol = 2)
```

### Interpretation

- **High $N_r$, low $N_m$** (top-left region): the root pathway is efficient;
  the optimiser invests more in $c_s$ and keeps $c_m$ small.
- **Low $N_r$, high $N_m$** (bottom-right region): mycorrhizal uptake is the
  dominant pathway; $c_m$ is large and $c_s$ is minimal.
- **Both low** (bottom-left): neither pathway can reach $N_T$ within the
  carbon budget (white); the tree would need to either accept a nitrogen
  shortfall or increase $c_I$.
- **Both high** (top-right): the target is reachable at low total carbon cost,
  leaving more sugar available for growth and respiration in the phloem pool.

These patterns directly inform how the nitrogen-feedback strategy would
modulate phloem sugar export ($c_s^{Ss}$ and $c_m^{Ss}$ in the Instruction
Booklet equations) relative to the empirical fixed-fraction default.
