---
title: "Tests: C++ against original R model"
author: "Joanna"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 5)
```

## What this document tests

The CASSIA model was originally written in R and then translated into C++.
This notebook checks that the C++ version produces the same outputs as the
original R model, using the stored reference output `Hyde_daily_original`
(Hyytiälä site, 2015–2018).

The weather data used here (`weather_original`) covers 2015–2017.  A `dates`
column is constructed from the day-of-year (`X`) column before the model is
run.

```{r load-and-run}
library(CASSIA)
library(ggplot2)
library(patchwork)
library(dplyr)
library(knitr)

data("weather_original")
data("Hyde_daily_original")

# Build the dates column (2016 is a leap year: 365 + 366 + 365 = 1096 rows)
weather_original$dates <- as.Date(
  strptime(paste(rep(2015:2017, times = c(365, 366, 365)), weather_original$X),
           format = "%Y %j"))

# Run with tests = TRUE so LH/LN are fixed to the calibrated values used in
# the original R model
out <- CASSIA_cpp(weather = weather_original, site = "Hyde", tests = TRUE)

# Trim reference to matching years
orig <- Hyde_daily_original[Hyde_daily_original$year %in% 2015:2017, ]
orig$dates <- as.Date(strptime(paste(orig$year, orig$day), format = "%Y %j"))

new_g  <- out$Growth
new_g$dates  <- weather_original$dates
new_su <- out$Sugar
new_su$dates <- weather_original$dates
```

---

## 1  Summary metrics

The table below compares the **annual means** of each output variable.
`ratio` = C++ mean / original mean; values near 1.0 indicate a good match.

The C++ model distinguishes *potential* growth (driven by temperature and
phenology only) from *actual* growth (potential × carbon-storage term).
The original R model only computed potential growth, so the meaningful
comparison is against the `*_potential` columns.

```{r metrics-table}
make_row <- function(new_col, orig_col, label) {
  nm <- mean(aligned[[new_col]], na.rm = TRUE)
  om <- mean(aligned[[orig_col]],  na.rm = TRUE)
  data.frame(Variable = label,
             `C++ mean`  = round(nm, 6),
             `Orig mean` = round(om, 6),
             Ratio       = round(nm / om, 4),
             check.names = FALSE)
}

# Align on date so the leap-year row difference doesn't cause length errors
aligned <- merge(
  new_g,
  orig[, c("dates", "bud", "wall_daily", "needle_daily", "root_daily", "height_daily", "P",
           "sugar", "starch", "storage_term", "Rg", "Rm")],
  by = "dates"
)

metrics <- rbind(
  make_row("bud_growth",    "bud",          "Bud growth (actual)"),
  make_row("diameter_growth","wall_daily",  "Diameter growth (actual)"),
  make_row("needle_growth", "needle_daily", "Needle growth (actual)"),
  make_row("root_growth",   "root_daily",   "Root growth (actual)"),
  make_row("height_growth", "height_daily", "Height growth (actual)"),
  make_row("bud_growth_potential",    "bud",          "Bud growth (potential)"),
  make_row("wall_growth_potential",   "wall_daily",   "Diameter growth (potential)"),
  make_row("needle_growth_potential", "needle_daily", "Needle growth (potential)"),
  make_row("roots_growth_potential",  "root_daily",   "Root growth (potential)"),
  make_row("height_growth_potential", "height_daily", "Height growth (potential)")
)

# GPP: weather P is in g C m-2 day-1; Hyde_daily stores per-stem value
# (P / no_trees * 10000/1000)
no_trees <- 1010
metrics <- rbind(
  metrics,
  data.frame(Variable   = "GPP (per stem, scaled)",
             `C++ mean` = round(mean(out$Preles$GPP / no_trees * 10, na.rm = TRUE), 6),
             `Orig mean` = round(mean(orig$P, na.rm = TRUE), 6),
             Ratio       = round(mean(out$Preles$GPP / no_trees * 10, na.rm = TRUE) /
                                 mean(orig$P, na.rm = TRUE), 4),
             check.names = FALSE)
)

kable(metrics, align = c("l", "r", "r", "r"),
      caption = "Mean values over 2015–2017; ratio = C++ / original")
```

---

## 2  Growth variables: time series

Each panel shows the C++ output (red) against the original R model (black).
The shaded period in January–February of each year illustrates where growth
is zero for both models (cold dormancy).

```{r growth-plots, fig.height = 14}
# 'aligned' was already built in the metrics chunk above.
model_colours <- c("Original R" = "black", "C++" = "firebrick")

plot_pair <- function(new_col, orig_col, title, units = "kg C stem⁻¹ day⁻¹") {
  df <- data.frame(
    dates = rep(aligned$dates, 2),
    value = c(aligned[[orig_col]], aligned[[new_col]]),
    Model = rep(c("Original R", "C++"), each = nrow(aligned))
  )
  ggplot(df, aes(x = dates, y = value, colour = Model)) +
    geom_line(linewidth = 1.1) +
    scale_colour_manual(values = model_colours) +
    scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
    labs(title = title, x = NULL, y = units, colour = NULL) +
    theme_bw(base_size = 11) +
    theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
}

p1 <- plot_pair("bud_growth",    "bud",          "Bud growth")
p2 <- plot_pair("diameter_growth","wall_daily",  "Diameter (wall) growth")
p3 <- plot_pair("needle_growth", "needle_daily", "Needle growth")
p4 <- plot_pair("root_growth",   "root_daily",   "Root growth")
p5 <- plot_pair("height_growth", "height_daily", "Height growth", units = "m day⁻¹")

(p1 / p2 / p3 / p4 / p5) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 3  Carbon storage

```{r storage-plots, fig.height = 8}
df_stor <- data.frame(
  dates        = aligned$dates,
  cpp_sugar    = new_su$sugar[match(aligned$dates, weather_original$dates)],
  cpp_starch   = new_su$starch[match(aligned$dates, weather_original$dates)],
  cpp_storage  = new_su$storage[match(aligned$dates, weather_original$dates)],
  orig_sugar   = aligned$sugar,
  orig_starch  = aligned$starch,
  orig_storage = aligned$storage_term
)

stor_colours <- c("Original R" = "black", "C++" = "firebrick")

make_stor_long <- function(cpp_col, orig_col) {
  data.frame(
    dates = rep(df_stor$dates, 2),
    value = c(df_stor[[orig_col]], df_stor[[cpp_col]]),
    Model = rep(c("Original R", "C++"), each = nrow(df_stor))
  )
}

ps1 <- ggplot(make_stor_long("cpp_sugar",   "orig_sugar"),
              aes(x = dates, y = value, colour = Model)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = stor_colours) +
  labs(title = "Sugar pool", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) + theme(plot.title = element_text(face = "bold"),
                                   legend.position = "bottom")

ps2 <- ggplot(make_stor_long("cpp_starch",  "orig_starch"),
              aes(x = dates, y = value, colour = Model)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = stor_colours) +
  labs(title = "Starch pool", x = NULL, y = "kg C stem⁻¹", colour = NULL) +
  theme_bw(base_size = 11) + theme(plot.title = element_text(face = "bold"),
                                   legend.position = "bottom")

ps3 <- ggplot(make_stor_long("cpp_storage", "orig_storage"),
              aes(x = dates, y = value, colour = Model)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = stor_colours) +
  labs(title = "Storage term (carbon availability fraction, 0–1)",
       x = NULL, y = "fraction", colour = NULL) +
  theme_bw(base_size = 11) + theme(plot.title = element_text(face = "bold"),
                                   legend.position = "bottom")

(ps1 / ps2 / ps3) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```

---

## 4  GPP

The `weather_original$P` column (g C m⁻² day⁻¹) is passed through unchanged
and stored in `Preles$GPP`.  The original model stored this divided by
`no_trees × (10000/1000)` to give per-stem GPP.  The comparison below
applies that scaling.

```{r gpp-plot}
df_gpp <- data.frame(
  dates    = aligned$dates,
  cpp_gpp  = out$Preles$GPP[match(aligned$dates, weather_original$dates)] / no_trees * 10,
  orig_gpp = aligned$P
)

df_gpp_long <- data.frame(
  dates = rep(df_gpp$dates, 2),
  value = c(df_gpp$orig_gpp, df_gpp$cpp_gpp),
  Model = rep(c("Original R", "C++ (weather P / no_trees × 10)"), each = nrow(df_gpp))
)

ggplot(df_gpp_long, aes(x = dates, y = value, colour = Model)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = c("Original R" = "black",
                                 "C++ (weather P / no_trees × 10)" = "firebrick")) +
  labs(title = "GPP per stem", x = NULL, y = "kg C stem⁻¹ day⁻¹", colour = NULL) +
  theme_bw(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

---

## 5  Potential vs actual growth

The C++ model adds an *actual growth* step that multiplies potential growth
by a carbon-storage availability term (`storage_term`, range 0–1).  With
correct initialisation the storage term is very close to 1.0 throughout the
growing season, so actual and potential growth track each other closely.
The panels below show where carbon limitation would reduce growth.

```{r actual-vs-potential, fig.height = 8}
pot_act_cols <- c("Potential" = "grey50", "Actual" = "firebrick")

make_pot_act <- function(pot_col, act_col, title) {
  df <- data.frame(
    dates = rep(aligned$dates, 2),
    value = c(aligned[[pot_col]], aligned[[act_col]]),
    Type  = rep(c("Potential", "Actual"), each = nrow(aligned))
  )
  ggplot(df, aes(x = dates, y = value, colour = Type)) +
    geom_line(linewidth = 1.1) +
    scale_colour_manual(values = pot_act_cols) +
    scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
    labs(title = title, x = NULL, y = "kg C stem⁻¹ day⁻¹", colour = NULL) +
    theme_bw(base_size = 11) +
    theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
}

pa1 <- make_pot_act("bud_growth_potential",    "bud_growth",    "Bud growth")
pa2 <- make_pot_act("needle_growth_potential", "needle_growth", "Needle growth")

(pa1 / pa2) + plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom")
```
